### 2021-07-18 paper/agujournaltemplate.tex
Start paper

%% March 2018
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%

%% To submit your paper:
\documentclass[draft]{agujournal2018}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Enter journal name here}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{=enter title here=}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{=list all authors here=}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{=number=}{=Affiliation Address=}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{=name=}{=email address=}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation
% \end{keypoints}

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
% A good abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
enter abstract here

\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

\section{Introduction}

The largest earthquakes on Earth occur in subduction zones, where large segments of the boundary between the subducting tectonic plate and the overlying plate can store energy for centuries and then release it in minutes during an earthquake. To better assess seismic hazard, it is thus necessary to improve our understanding of subduction zone processes. Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks. Slow slip events can last from a few days to several years, and have a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs (Beroza and Ide, 2011 ~\cite{BER_2011}).  Moreover, whereas regular earthquakes occur in the seismogenic / locked zone, slow slip events occur downdip of the locked section of the subduction zone. Interactions between the slow slip zone and the seismogenic zone could potentially trigger large earthquakes. Through future study, this interaction may improve our understanding of the seismic hazard.

Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next. They have been widely used for geophysical applications. However, very few studies have used wavelet methods to analyze recordings of slow slip events. The aim of this project is thus to use \textbf{wavelet methods to analyze GNSS recordings of slow slip events in Cascadia  and New Zealand}. These advanced signal processing methods will enable extracting much more information from the GNSS data currently available and will improve our knowledge of this newly discovered geological phenomenon. Specifically, as a result of the project, I will be able to \emph{detect possible smaller (magnitude 5) slow slip events} in Zealand that may be currently undetected with standard methods and \emph{detect longer (months to years) slow slip events} that are more difficult to detect than slow slip events with a short duration (days to weeks). Additionally, the methods developed for the study of slow slip events can have a broader potential impact for the detection of transient signals in GNSS time series. The codes developed will be publicly released on the student's GitHub repository (url available on the student's CV).

\textbf{2. Introduction}

As with ordinary earthquakes, slow slip events are caused by slip on a fault, such as the plate boundary between a tectonic plate subducting under another tectonic plate. However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes. Moreover, the seismic waves they generate are much weaker than the seismic waves generated by ordinary earthquakes, and may not be detectable.  A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion. Slow slip events have been observed in many subduction zones, such as Cascadia, Nankai (southwest Japan), Alaska, Costa Rica, Mexico, and New Zealand (Beroza and Ide, 2011 ~\cite{BER_2011}; Audet and Kim, 2016 ~\cite{AUD_2016}).

In many places, tectonic tremor are also observed in relation to slow slip. Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs ; Shelly \textit{et al.}, 2007 ~\cite{SHE_2007_nature}), that is small magnitude earthquakes (M $\sim$ 1), for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz). In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations (Obara, 2002 ~\cite{OBA_2002}; Rogers and Dragert, 2003 ~\cite{ROG_2003}). Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS). However, this is not always the case. For instance, in northern New Zealand, tremor are more challenging to detect, and seem to be located downdip of the slow slip on the plate boundary.

New Zealand exhibits a wide range of slow slip behavior, and is therefore an exciting site for research. The tectonics of the North Island of New Zealand are dominated by the westward subduction of the Pacific Plate under the Australian Plate at the Hikurangi Trench. Two types of slow slip events have been observed at the Hikurangi margin. Shallow (10-15 km depth), shorter (1-3 weeks), and usually smaller (Mw 6.3-6.8) slow slip events have been observed every 18-24 months in the northern part of the margin. Deeper (35-60 km depth), longer (12-18 months), and larger (Mw 7.0) slow slip events have been observed every 5 years in the southern part of the margin (Wallace and Beavan, 2010 ~\cite{WAL_2010}; Todd and Schwartz, 2016 ~\cite{TOD_2016}).

It used to be thought that there were no tremor associated with slow slip events in northern Hikurangi. Delahaye \textit{et al.} (2009 ~\cite{DEL_2009}) observed an increase in the rate of microseismicity downdip of the 2004 Gisborne slow slip event. More recently, however, Kim \textit{et al.} (2011 ~\cite{KIM_2011}) detected a low level of tremor activity that increased during the 2010 Gisborne slow slip event. As was the case for the microearthquakes, the source of the tremor was located downdip of the slow slip patch determined from GNSS data. Ide (2012 ~\cite{IDE_2012}) detected tremor downdip of the location of two deep slow slip events observed by Wallace and Eberhart-Phillips (2013 ~\cite{WAL_2013}) in 2006 and 2008. However, contrary to Episodic Tremor and Slip events in Cascadia and Nankai, the tremor activity did not seem to increase during the slow slip events. Todd and Schwartz (2016 ~\cite{TOD_2016}) detected tremor associated with most of the shallow slow slip events between 2010 and 2015, and located downdip of the geodetically inferred slip area. They also detected deeper tremor between 20 and 50 km depth with unclear origin. They hypothesized that these tremor may be related to currently undetected deep long-term slow slip events. In other subduction zones, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data (Aguiar \textit{et al.}, 2009 ~\cite{AGU_2009} ; Frank, 2016 ~\cite{FRA_2016}). However, as there is no clear relationship between tremor and slow slip occurrence in New Zealand, \textbf{these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip.}

Wavelet methods have been widely used for geophysical applications (Kumar and Foufoula-Georgiou, 1997 ~\cite{KUM_1997}). However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events with a signal-to-noise ratio greater than 1. Alba \textit{et al.} (2019 ~\cite{ALB_2019}) analyzed tidal records at four different sites during slow slip events. They applied a Discrete Wavelet Transform (DWT) to each of the sites studied, in order to retrieve the noise signal shared by all sites and remove it from the time series. They could then clearly see a difference in uplift between the two tide gauges at Port Angeles and Port Townsend. Szeliga \textit{et al.} (2008 ~\cite{SZE_2008}) used a continuous wavelet transform of the GPS time series to pick the onset of the ETS events. Finally, instead of using wavelets in the time domain, Ohtani \textit{et al.} (2010 ~\cite{OHT_2010}) used 2D wavelet functions in the spatial domain to detect slow slip events. Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone (Wei \textit{et al.}, 2012 ~\cite{WEI_2012}). While wavelet methods have a great potential for analyzing slow slip, their potential for the detection of small size (magnitude 5) events and long term (several months) events with signal-to-noise ratio lower than 1 has been relatively untapped.

\section{Data}

\section{Method}

The wavelet methods for time series analysis are explained in a more detailed way in Percival \& Walden (2000 ~\cite{PER_2000}). \\

\textbf{Discrete Wavelet Transform}

The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, ... , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , ... , N - 1 \right)$. If we denote $J$ the level of the wavelet decomposition, and we have $N = n* 2^J$, where n is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ... , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$. \\

Each wavelet vector $W_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $V_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. \\

We can also define for $j = 1 , ... , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to scale $\tau_j = dt 2^{j - 1}$. Similarly, we can define for $j = 1 , ... , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher. Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:

\begin{equation}
X = \sum_{j = 1}^{J} D_j + S_J
\end{equation}

One main advantage of the DWT is that it is an orthonormal transform, and thus we can write the analysis of variance (ANOVA):

\begin{equation}
\left\Vert X \right\Vert ^2 = \left\Vert W \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert W_j \right\Vert ^2 + \left\Vert V_J \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert D_j \right\Vert ^2 + \left\Vert S_J \right\Vert ^2
\end{equation}

Moreover, the DWT can be computed using $O \left( N \right)$ multiplications. \\

However, the DWT present several disadvantages:

\begin{itemize}
	\item The length of the time series must be a multiple of $2^J$ where $J$ is the level of the DWT decomposition.
	\item The time step of the wavelet vector $W_j$ is $dt 2^j$, which may not correspond to the time when some interesting phenomenon is visible on the original time series.
	\item When we circularly shift the time series, the corresponding wavelet coefficients, details and smooths are not a circularly shifted version of the wavelet coefficients, details and smooths of the original time series. Thus, the values of the wavelet coefficients, details and smooths are strongly dependent on the time when we start experimentally gathering the data.
	\item When we filter the time series to obtain the details and smooths, we introduce a phase shift, which makes difficult to line up meaningfully the features of the MRA with the original time series.
\end{itemize}

\textbf{Maximum Overlap Discrete Wavelet Transform}

To get rid of these problems, we introduce the Maximum Overlap Discrete Wavelet Transform (MODWT). The MODWT transforms the time series $X_t \left( t = 0, ... , N - 1 \right)$ into J wavelet vectors $\widetilde{W}_j \left( j = 1 , ... , J \right)$ of length $N$ and a scaling vector $\widetilde{V}_J$ of length $N$. As is the case for the DWT, each wavelet vector $\widetilde{W}_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $\widetilde{V}_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. \\

As is the case for the DWT, we can write the MRA:

\begin{equation}
X = \sum_{j = 1}^{J} \widetilde{D}_j + \widetilde{S}_J
\end{equation}

and the ANOVA:

\begin{equation}
\left\Vert X \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert \widetilde{W}_j \right\Vert ^2 + \left\Vert \widetilde{V}_J \right\Vert ^2
\end{equation}

Now, we have the following properties:

\begin{itemize}
	\item The MODWT of a time series can be defined for any length $N$.
	\item The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series.
	\item When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount.
	\item The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series.
\end{itemize}

However, the MODWT has some disadvantages over the DWT:

\begin{itemize}
	\item The MODWT can only be computed using $O \left( N \log_2 N \right)$ multiplications.
	\item We can no longer write the ANOVA for the details and smooths:

\begin{equation}
\left\Vert X \right\Vert ^2 \neq \sum_{j = 1}^{J} \left\Vert D_j \right\Vert ^2 + \left\Vert S_J \right\Vert ^2 \text{ and } \left\Vert \widetilde{W}_j \right\Vert ^2 \neq \left\Vert \widetilde{D}_j \right\Vert ^2
\end{equation}

\end{itemize}

Example: Synthetics

\section{Results}

Explanation vespagram

Correlation

\section{Discussion}

\section{Conclusion}

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}[h]
% \centering
% when using pdflatex, use pdf file:
% \includegraphics[natwidth=800px,natheight=600px]{figsamp.pdf}
%
% when using dvips, use .eps file:
% \includegraphics[natwidth=800px,natheight=600px]{figsamp.eps}
%
% \caption{Short caption}
% \label{figone}
%  \end{figure}
%
% We recommend that you provide the native width and height (natwidth, natheight) of your figures.
% Specifying native dimensions ensures that your figures are properly scaled
%
%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
Enter acknowledgments, including your data availability statement, here.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BibTeX is preferred:
%
% \bibliography{<name of your .bib file>}
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
%% Example \citet and \citep:
%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
%  \citet{Ghel00}, and \citet{Leit74}.

%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
%  \citep{Ghel00, Leit74}.

%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.

\end{document}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2021-07-22 src/compare_tremor_GPS.py
Plot cc

@@ -1,0 +1,447 @@
+"""
+Script to cross-correlate GPS and tremor MODWT
+"""
+
+import datetime
+import matplotlib.cm as cm
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import pickle
+
+from datetime import datetime
+from math import cos, floor, pi, sin, sqrt
+from matplotlib.colors import Normalize
+
+import date
+import DWT
+from MODWT import get_DS, get_scaling, pyramid
+
+def correlate(X, Y, ncor):
+    """
+    """
+    N = len(X)
+    cc = np.zeros(2 * ncor + 1)
+    for i in range(0, 2 * ncor + 1):
+        if i < ncor:
+            input1 = X[(ncor - i) : N]
+            input2 = Y[0 : (N - ncor + i)]
+        else:
+            input1 = X[0 : (N - i + ncor)]
+            input2 = Y[(i - ncor) : N]
+        cc[i] = np.mean((input1 - np.mean(input1)) * (input2 - np.mean(input2))) / (np.std(input1) * np.std(input2))
+    return cc
+
+def compare_tremor_GPS(station_file, lats, lons, dataset, direction, \
+    radius_GPS, radius_tremor, slowness, wavelet, J, J0):
+    """
+    """
+    # Read station file
+    stations = pd.read_csv(station_file, sep=r'\s{1,}', header=None, engine='python')
+    stations.columns = ['name', 'longitude', 'latitude']
+
+    # Read tremor files (A. Wech)
+    data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
+    data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2010 = pd.read_csv('../data/tremor/tremor_events-2010-01-01T00 00 00-2010-12-31T23 59 59.csv')
+    data_2010['time '] = pd.to_datetime(data_2010['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2011 = pd.read_csv('../data/tremor/tremor_events-2011-01-01T00 00 00-2011-12-31T23 59 59.csv')
+    data_2011['time '] = pd.to_datetime(data_2011['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2012 = pd.read_csv('../data/tremor/tremor_events-2012-01-01T00 00 00-2012-12-31T23 59 59.csv')
[…]

### 2021-07-22 src/vespagram.py

@@ -100,114 +100,6 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         filename = 'tmp/' + dataset + '_' + station + '_' + direction + '.pkl'
         pickle.dump([time, disp, W, V, D, S], open(filename, 'wb'))
 
-def compute_wavelets_tremor(lats, lons, radius_tremor, wavelet, J):
-    """
-    """
-    # Read tremor files (A. Wech)
-    data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
-    data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2010 = pd.read_csv('../data/tremor/tremor_events-2010-01-01T00 00 00-2010-12-31T23 59 59.csv')
-    data_2010['time '] = pd.to_datetime(data_2010['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2011 = pd.read_csv('../data/tremor/tremor_events-2011-01-01T00 00 00-2011-12-31T23 59 59.csv')
-    data_2011['time '] = pd.to_datetime(data_2011['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2012 = pd.read_csv('../data/tremor/tremor_events-2012-01-01T00 00 00-2012-12-31T23 59 59.csv')
-    data_2012['time '] = pd.to_datetime(data_2012['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2013 = pd.read_csv('../data/tremor/tremor_events-2013-01-01T00 00 00-2013-12-31T23 59 59.csv')
-    data_2013['time '] = pd.to_datetime(data_2013['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2014 = pd.read_csv('../data/tremor/tremor_events-2014-01-01T00 00 00-2014-12-31T23 59 59.csv')
-    data_2014['time '] = pd.to_datetime(data_2014['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2015 = pd.read_csv('../data/tremor/tremor_events-2015-01-01T00 00 00-2015-12-31T23 59 59.csv')
-    data_2015['time '] = pd.to_datetime(data_2015['time '], format='%Y-%m-%d %H:%M:%S')
[…]

### 2021-07-25 paper/agujournaltemplate.tex
Write synthetics part

@@
 
-Example: Synthetics
+To illustrate the wavelet transform method, we first apply the MODWT to synthetics data.
+As slow slip events occur in Cascadia on a regular basis, every twelve to eighteen months, we create a synthetic signal of period $T = 500$ days.
+To reproduce the ground displacement observed on the longitudinal component of GPS stations in Cascadia, we divide each period into two parts: In the first part of duration $T - N$, the displacement is linearly increasing and corresponds to the secular plate motion in the eastern direction; in the second part of duration $N$, the displacement is linearly decreasing and corresponds to a slow slip event on a reverse fault at depth triggering a ground displacement in the western direction.
+To see the effect of the magnitude of the slow slip event, we use different values for $N = 2, 5, 10, 20$ days.
+Figure X shows the synthetics, the details of the wavelet decomposition for levels 1 to 8, and the smooth for the four durations of a slow slip event.
+\\
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={5cm 6cm 5cm 9cm},clip]{figures/500_DS.pdf}
+\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
+\label{pngfiguresample}
+\end{figure}
+
+The ramp-like signal is transformed through the wavelet filtering into a waveform with first a positive peak and then a negative peak.
+The width of the waveform increases with the scale level.
+For the 8th level of the wavelet decomposition, the width of the waveform is nearly as large as the time between two events.
+We do not show details at larger scales as the corresponding waveforms would start to merge two contiguous events together, and make the wavelet decomposition less interpretable.
+For an event of duration 2 days, the wavelet details at levels higher than 2 have a larger amplitude than the wavelet detail at level 1.
+For an event of duration 5 days, the wavelet details at levels higher than 3 have a larger amplitude than the wavelet details at lower scales.
+For an event of duration 10 days, the wavelet details at levels higher than 5 have a larger amplitude than the wavelet details at lower scales.
+For an event of duration 20 days, the wavelet details at levels higher than 6 have a larger amplitude than the wavelet details at lower scales.
+Thus, the scale levels at which an event is being seen in the wavelet details give us an indication about the duration (and the magnitude) of the slow slip event.
+We expect the big slow slip events of magnitude 6-7 that lasts about 10 days to start being visible at the level 5 of the wavelet decomposition, but to not be noticeable at lower time scales.
+\\
 

### 2021-07-25 src/synthetics.py

@@ -27,8 +27,15 @@ np.random.seed(0)
 
 # Loop on time between events
 for timestep in timesteps:
+
+    # Start figure
+    params = {'xtick.labelsize':20,
+              'ytick.labelsize':20}
+    pylab.rcParams.update(params)   
+    fig = plt.figure(1, figsize=(5 * len(durations), 3 * (J + 3)))
+
     # Loop on duration
-    for duration in durations:
+    for i, duration in enumerate(durations):
         # Create displacements vectors
         part1 = np.arange(0, timestep - duration) / (timestep - duration)
         part2 = (timestep - np.arange(timestep - duration, timestep)) / duration
@@ -39,29 +46,36 @@ for timestep in timesteps:
         (D, S) = MODWT.get_DS(disp, W, name, J)
         maxD = max([np.max(Dj) for Dj in D])
         minD = min([np.min(Dj) for Dj in D])
-        # Plot MODWT
-        params = {'xtick.labelsize':24,
-                  'ytick.labelsize':24}
-        pylab.rcParams.update(params)   
-        fig = plt.figure(1, figsize=(15, 5 * (J + 3)))
         # Plot data
-        plt.subplot2grid((J + 2, 1), (J + 1, 0))
+        ax = plt.subplot2grid((J + 2, len(durations)), (J + 1, i))
         plt.plot(time, disp, 'k', label='Data')
+        plt.xlabel('Time (days)', fontsize=20)
         plt.ylim(-2.0, 2.0)
-        plt.legend(loc=3, fontsize=14)
+        plt.legend(loc=3, fontsize=20)
+        if i != 0:
+            ax.axes.yaxis.set_ticks([])
         # Plot details
         for j in range(0, J):
-            plt.subplot2grid((J + 2, 1), (J - j, 0))
+            ax = plt.subplot2grid((J + 2, len(durations)), (J - j, i))
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
             plt.ylim(minD, maxD)
-            plt.legend(loc=3, fontsize=14)
+            plt.legend(loc=3, fontsize=20)
+            ax.axes.xaxis.set_ticks([])
+            if i != 0:
+                ax.axes.yaxis.set_ticks([])
         # Plot smooth
-        plt.subplot2grid((J + 2, 1), (0, 0))
+        ax = plt.subplot2grid((J + 2, len(durations)), (0, i))
[…]

### 2021-07-27 paper/agujournaltemplate.tex
Work on introduction

@@
 
-Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next.
  They have been widely used for geophysical applications.
@@
 They hypothesized that these tremor may be related to currently undetected deep long-term slow slip events.
-In other subduction zones, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data (Aguiar \textit{et al.}, 2009 ~\cite{AGU_2009} ; Frank, 2016 ~\cite{FRA_2016}).
-However, as there is no clear relationship between tremor and slow slip occurrence in New Zealand, \textbf{these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip.}
 
-Wavelet methods have been widely used for geophysical applications (Kumar and Foufoula-Georgiou, 1997 ~\cite{KUM_1997}).
-However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events with a signal-to-noise ratio greater than 1.
-Alba \textit{et al.} (2019 ~\cite{ALB_2019}) analyzed tidal records at four different sites during slow slip events.
-They applied a Discrete Wavelet Transform (DWT) to each of the sites studied, in order to retrieve the noise signal shared by all sites and remove it from the time series.
-They could then clearly see a difference in uplift between the two tide gauges at Port Angeles and Port Townsend.
-Szeliga \textit{et al.} (2008 ~\cite{SZE_2008}) used a continuous wavelet transform of the GPS time series to pick the onset of the ETS events.
-Finally, instead of using wavelets in the time domain, Ohtani \textit{et al.} (2010 ~\cite{OHT_2010}) used 2D wavelet functions in the spatial domain to detect slow slip events.
+In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data.
+For instance, ~\cite{AGU_2009} computed the GPS-estimated moment release for 23 ETS events in Cascadia between 1997 and 2008.
+Simultaneously, they computed the cumulative number of hours of tectonic tremor recorded for each event.
+They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8.
+For all these events, at least 50 hours of tectonic tremor where observed simultaneously with the GPS deformation.
+However, many smaller bursts of tremor of duration 1 to 50 hours were also observed in between the big ETS events.
+Based on the relationship between slow slip moment and number of hours of tremor, ~\cite{AGU_2009} suggested that smaller slow slip events of magnitude 5-6 may occur simultaneously with the tremor bursts without being detectable in the GPS data.
+\\
+
+~\cite{FRA_2016} transformed the GPS time series into daily increments of surface motion by computing the first order differentiation of the time series.
+He then discarded the daily increments observed during known big slow slip events, and focused on the inter-events period.
+He divided the daily increments into two groups: the first group contains days when slow seismicity (tectonic tremor and LFEs) is detected, the second group contains days when the numbers of LFEs (for Guerrero) or tremor (for Cascadia) is lower than a threshold.
+He then stacked separately the two groups of daily increments and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
+He also observed a cumulative displacement in the southern direction (for Guerrero) and the western direction (for Cascadia) corresponding to the release period when tremor and LFEs are observed.
+This reverse displacement corresponds to smaller slow slip events not directly observable in the GPS data.
+\\
+
+However, in other subduction zones such as New Zealand, there is no clear relationship between tremor and slow slip occurrence and these methods cannot be applied.
+We thus need other methods to be able to better detect and quantify slow slip.
+
+Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next.
+Wavelet methods have been widely used for geophysical applications ~\cite{KUM_1997}.
+However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events.
+
+~\cite{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011.
+The noise in the tide gauges data is associated with tides, and ocean and atmospheric noise on multiple timescales (a few days for storms to decades for oscillations between ocean basins), and is assumed to be coherent between each of the four tidal gauges studied.
+On the contrary, the uplift due to ETS events should be different at each tidal gauge.
+They first removed the tides using NOAA hourly harmonic tidal predictions.
+They then removed the residual noise using a method based on the Discrete Wavelet Transform (DWT).
+More precisely, the authors applied a DWT to each of the four sites studied, and to the average of the four sites.
+Then, for each level of the DWT decomposition, they carried out a linear regression between the detail for one site and the detail for the average of the four sites.
+This process gives a coefficient for each level and for each site.
+They then constructed a noise signal for each site by multiplying the coefficient from the linear regression by the detail of the average over the four sites, and summing for all levels.
+The noise signal thus obtained was then removed from the time series.
+They then stacked multiple events to obtain an average event uplift rate, aligning the 12 ETS events using exact timing from GPS data.
+A difference in uplift between the two tidal gauges at Port Angeles and Port Townsend was then clearly seen in the stacked time series.
+Finally, the authors removed the long-term uplift rate and the long-term sea level rise to obtain an average inter-event uplift rate.
+They found that the inter-event deformation at a site is equal and opposite to the deformation during an ETS event, suggesting that ETS events are, on average, releasing the strain accumulated between ETS events.
+\\
+
+~\cite{SZE_2008}) determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005.
+They stabilized the GPS time series using a reference set of stations from stable North America.
+They then modelled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, and Heaviside step functions corresponding to earthquakes and hardware upgrades.
+The linear system was then solved using a weighted QR decomposition.
+Finally, they applied a Gaussian wavelet transform to the residual time series to get the exact timing of the slow slip at each GPS station.
+The succeeding wavelet basis functions are increasingly sensitive to temporal localization of a given signal, and the onset of faulting appears on the wavelet spectrum as an amplitude spike present over all frequencies.
+The offset for each slow slip event was then used to invert for the slow slip at depth by assuming a thrust fault slip at each subfault of the plate boundary.
+An equivalent moment magnitude was thus obtained.
+\\
+
+Finally, instead of using wavelets in the time domain, ~\cite{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events.
+They designed the Network Stain Filter (NSF) to detect transient deformation signals from large-scale geodetic arrays.
+Contrary to their previous work on the Network Inversion Filter (NIF), there is no need to specify potential sources of deformation.
+They modeled the position of the GPS station by the sum of the secular velocity, a spatially coherent field, site-specific noise, reference frame errors, and observation errors.
+The spatial displacement field is modeled by the sum of basis wavelets (the Deslauriers-Dubuc wavelet of degree 3) with time-varying weights.
+The transient is considered to be nearly steady-sate, so that it has spatial weights for the displacement and the velocity, but the acceleration is modeled by a random walk with a time-varying variance.
+All the time varying coefficients are estimated using Kalman filtering, and the optimization problem is regularized with the spatial sum of the transient strain rate field.
 Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone (Wei \textit{et al.}, 2012 ~\cite{WEI_2012}).
+\\
+
 While wavelet methods have a great potential for analyzing slow slip, their potential for the detection of small size (magnitude 5) events and long term (several months) events with signal-to-noise ratio lower than 1 has been relatively untapped.
@@
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={5cm 6cm 5cm 9cm},clip]{figures/500_DS.pdf}
-\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
-\label{pngfiguresample}
-\end{figure}
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={5cm 6cm 5cm 9cm},clip]{figures/500_DS.pdf}
+%\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
+%\label{pngfiguresample}
+%\end{figure}
 

### 2021-07-29 paper/agujournaltemplate.tex
Wrote introduction

@@
 
-The largest earthquakes on Earth occur in subduction zones, where large segments of the boundary between the subducting tectonic plate and the overlying plate can store energy for centuries and then release it in minutes during an earthquake.
-To better assess seismic hazard, it is thus necessary to improve our understanding of subduction zone processes.
 Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks.
-Slow slip events can last from a few days to several years, and have a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs (Beroza and Ide, 2011 ~\cite{BER_2011}).
-Moreover, whereas regular earthquakes occur in the seismogenic / locked zone, slow slip events occur downdip of the locked section of the subduction zone.
-Interactions between the slow slip zone and the seismogenic zone could potentially trigger large earthquakes.
-Through future study, this interaction may improve our understanding of the seismic hazard.
-
- They have been widely used for geophysical applications.
-However, very few studies have used wavelet methods to analyze recordings of slow slip events.
-The aim of this project is thus to use \textbf{wavelet methods to analyze GNSS recordings of slow slip events in Cascadia  and New Zealand}.
-These advanced signal processing methods will enable extracting much more information from the GNSS data currently available and will improve our knowledge of this newly discovered geological phenomenon.
-Specifically, as a result of the project, I will be able to \emph{detect possible smaller (magnitude 5) slow slip events} in Zealand that may be currently undetected with standard methods and \emph{detect longer (months to years) slow slip events} that are more difficult to detect than slow slip events with a short duration (days to weeks).
-Additionally, the methods developed for the study of slow slip events can have a broader potential impact for the detection of transient signals in GNSS time series.
-The codes developed will be publicly released on the student's GitHub repository (url available on the student's CV).
-
-\textbf{2. Introduction}
-
 As with ordinary earthquakes, slow slip events are caused by slip on a fault, such as the plate boundary between a tectonic plate subducting under another tectonic plate.
-However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes.
-Moreover, the seismic waves they generate are much weaker than the seismic waves generated by ordinary earthquakes, and may not be detectable.
+However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes, and they have a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}.
 A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion.
-Slow slip events have been observed in many subduction zones, such as Cascadia, Nankai (southwest Japan), Alaska, Costa Rica, Mexico, and New Zealand (Beroza and Ide, 2011 ~\cite{BER_2011}; Audet and Kim, 2016 ~\cite{AUD_2016}).
+Slow slip events have been observed in many subduction zones, such as Cascadia, Nankai (southwest Japan), Alaska, Costa Rica, Mexico, and New Zealand ~\citep{BER_2011,AUD_2016}.
+\\
 
@@
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
-Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs ; Shelly \textit{et al.}, 2007 ~\cite{SHE_2007_nature}), that is small magnitude earthquakes (M $\sim$ 1), for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
-In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations (Obara, 2002 ~\cite{OBA_2002}; Rogers and Dragert, 2003 ~\cite{ROG_2003}).
+Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, that is small magnitude earthquakes (M $\sim$ 1), for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
+In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations~\citep{OBA_2002,ROG_2003}.
 Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS).
@@
 For instance, in northern New Zealand, tremor are more challenging to detect, and seem to be located downdip of the slow slip on the plate boundary.
-
-New Zealand exhibits a wide range of slow slip behavior, and is therefore an exciting site for research.
-The tectonics of the North Island of New Zealand are dominated by the westward subduction of the Pacific Plate under the Australian Plate at the Hikurangi Trench.
-Two types of slow slip events have been observed at the Hikurangi margin.
-Shallow (10-15 km depth), shorter (1-3 weeks), and usually smaller (Mw 6.3-6.8) slow slip events have been observed every 18-24 months in the northern part of the margin.
-Deeper (35-60 km depth), longer (12-18 months), and larger (Mw 7.0) slow slip events have been observed every 5 years in the southern part of the margin (Wallace and Beavan, 2010 ~\cite{WAL_2010}; Todd and Schwartz, 2016 ~\cite{TOD_2016}).
-
-It used to be thought that there were no tremor associated with slow slip events in northern Hikurangi.
-Delahaye \textit{et al.} (2009 ~\cite{DEL_2009}) observed an increase in the rate of microseismicity downdip of the 2004 Gisborne slow slip event.
-More recently, however, Kim \textit{et al.} (2011 ~\cite{KIM_2011}) detected a low level of tremor activity that increased during the 2010 Gisborne slow slip event.
-As was the case for the microearthquakes, the source of the tremor was located downdip of the slow slip patch determined from GNSS data.
-Ide (2012 ~\cite{IDE_2012}) detected tremor downdip of the location of two deep slow slip events observed by Wallace and Eberhart-Phillips (2013 ~\cite{WAL_2013}) in 2006 and 2008.
-However, contrary to Episodic Tremor and Slip events in Cascadia and Nankai, the tremor activity did not seem to increase during the slow slip events.
-Todd and Schwartz (2016 ~\cite{TOD_2016}) detected tremor associated with most of the shallow slow slip events between 2010 and 2015, and located downdip of the geodetically inferred slip area.
-They also detected deeper tremor between 20 and 50 km depth with unclear origin.
-They hypothesized that these tremor may be related to currently undetected deep long-term slow slip events.
+\\
 
 In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data.
-For instance, ~\cite{AGU_2009} computed the GPS-estimated moment release for 23 ETS events in Cascadia between 1997 and 2008.
+For instance, ~\citet{AGU_2009} computed the GPS-estimated moment release for 23 ETS events in Cascadia between 1997 and 2008.
 Simultaneously, they computed the cumulative number of hours of tectonic tremor recorded for each event.
@@
 However, many smaller bursts of tremor of duration 1 to 50 hours were also observed in between the big ETS events.
-Based on the relationship between slow slip moment and number of hours of tremor, ~\cite{AGU_2009} suggested that smaller slow slip events of magnitude 5-6 may occur simultaneously with the tremor bursts without being detectable in the GPS data.
+Based on the relationship between slow slip moment and number of hours of tremor, ~\citet{AGU_2009} suggested that smaller slow slip events of magnitude 5-6 may occur simultaneously with the tremor bursts without being detectable in the GPS data.
 \\
 
-~\cite{FRA_2016} transformed the GPS time series into daily increments of surface motion by computing the first order differentiation of the time series.
+~\citet{FRA_2016} transformed the GPS time series into daily increments of surface motion by computing the first order differentiation of the time series.
 He then discarded the daily increments observed during known big slow slip events, and focused on the inter-events period.
@@
 We thus need other methods to be able to better detect and quantify slow slip.
+\\
 
 Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next.
-Wavelet methods have been widely used for geophysical applications ~\cite{KUM_1997}.
+Wavelet methods have been widely used for geophysical applications ~\citep{KUM_1997}.
 However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events.
+\\
 
-~\cite{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011.
+~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011.
 The noise in the tide gauges data is associated with tides, and ocean and atmospheric noise on multiple timescales (a few days for storms to decades for oscillations between ocean basins), and is assumed to be coherent between each of the four tidal gauges studied.
@@
 They first removed the tides using NOAA hourly harmonic tidal predictions.
-They then removed the residual noise using a method based on the Discrete Wavelet Transform (DWT).
+They then removed the residual noise using a method based on the DWT.
 More precisely, the authors applied a DWT to each of the four sites studied, and to the average of the four sites.
@@
 
-~\cite{SZE_2008}) determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005.
+~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005.
 They stabilized the GPS time series using a reference set of stations from stable North America.
@@
 
-Finally, instead of using wavelets in the time domain, ~\cite{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events.
+Finally, instead of using wavelets in the time domain, ~\citet{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events.
 They designed the Network Stain Filter (NSF) to detect transient deformation signals from large-scale geodetic arrays.
@@
 All the time varying coefficients are estimated using Kalman filtering, and the optimization problem is regularized with the spatial sum of the transient strain rate field.
-Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone (Wei \textit{et al.}, 2012 ~\cite{WEI_2012}).
+Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone ~\citep{WEI_2012}.
 \\
 
-While wavelet methods have a great potential for analyzing slow slip, their potential for the detection of small size (magnitude 5) events and long term (several months) events with signal-to-noise ratio lower than 1 has been relatively untapped.
+In this study, we use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia.
+Our objective is to verify that there is a good correlation between slow slip events detected with only GPS data, and slow slip events detected with only seismic data.
+We thus want to demonstrate that the wavelet-based detection method can be applied to detect possible smaller (magnitude 5) slow slip events that may be currently undetected with standard methods.
 
@@
 %
-% \bibliography{<name of your .bib file>}
+\bibliography{bibliography.bib}
 %

### 2021-08-08 src/synthetics.py
synthetics

@@ -77,5 +77,5 @@ for timestep in timesteps:
     # Save figure
     title = 'Time between events = ' + str(timestep)
     plt.suptitle(title, fontsize=30)
-    plt.savefig('synthetics/' + str(timestep) + '_DS.pdf', format='pdf')
+    plt.savefig('synthetics/' + str(timestep) + '_DS.eps', format='eps')
     plt.close(1)

### 2021-08-08 src/synthetics.py
Remake synthetics for paper

@@ -75,7 +75,8 @@ for timestep in timesteps:
         plt.title(title, fontsize=20)
 
     # Save figure
-    title = 'Time between events = ' + str(timestep)
-    plt.suptitle(title, fontsize=30)
+#    title = 'Time between events = ' + str(timestep)
+#    plt.suptitle(title, fontsize=30)
+    plt.tight_layout()
     plt.savefig('synthetics/' + str(timestep) + '_DS.eps', format='eps')
     plt.close(1)

### 2021-08-08 src/tremor_MODWT.py
MODWT tremor

@@ -2,25 +2,141 @@
 Script to compute MODWT of cumulative tremor count
 """
 
-import datetime
-import matplotlib.cm as cm
+#import datetime
+#import matplotlib.cm as cm
 import matplotlib.pylab as pylab
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import pickle
 
-from datetime import datetime
+#from datetime import datetime
 from math import cos, floor, pi, sin, sqrt
-from matplotlib.colors import Normalize
-from scipy.io import loadmat
-from scipy.signal import butter, detrend, lfilter
+#from matplotlib.colors import Normalize
+#from scipy.io import loadmat
+from scipy.signal import butter, lfilter
 
 import date
 import DWT
 from MODWT import get_DS, get_scaling, pyramid
 
-def compute_wavelets_tremor(lats, lons, radius_tremor, wavelet, J):
+def tremor_polynomial(lats, lons, radius_tremor, wavelet, J, degree):
+    """
+    """
+    # Read tremor files (A. Wech)
+    data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
+    data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2010 = pd.read_csv('../data/tremor/tremor_events-2010-01-01T00 00 00-2010-12-31T23 59 59.csv')
+    data_2010['time '] = pd.to_datetime(data_2010['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2011 = pd.read_csv('../data/tremor/tremor_events-2011-01-01T00 00 00-2011-12-31T23 59 59.csv')
+    data_2011['time '] = pd.to_datetime(data_2011['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2012 = pd.read_csv('../data/tremor/tremor_events-2012-01-01T00 00 00-2012-12-31T23 59 59.csv')
+    data_2012['time '] = pd.to_datetime(data_2012['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2013 = pd.read_csv('../data/tremor/tremor_events-2013-01-01T00 00 00-2013-12-31T23 59 59.csv')
+    data_2013['time '] = pd.to_datetime(data_2013['time '], format='%Y-%m-%d %H:%M:%S')
[…]

### 2021-08-08 paper/agujournaltemplate.tex
Figures for MODWT

@@
 In this study, we use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia.
-Our objective is to verify that there is a good correlation between slow slip events detected with only GPS data, and slow slip events detected with only seismic data.
+Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
 We thus want to demonstrate that the wavelet-based detection method can be applied to detect possible smaller (magnitude 5) slow slip events that may be currently undetected with standard methods.
+\\
 
@@
 
+We focused our study on northwest Washington State.
+For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University.
+These are network solutions in ITRF2008 with phase ambiguities resolved.
+Solutions are computed with JPL/NASA orbits and satellite clocks.
+North, East, and Vertical directions are available.
+However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio.
+The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times eries with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}.
+For the seismic data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}.
+Tremor were detected and located using waveform envelope correlation and clustering and a centroid location is available for every given five-minute time window when tremor was detected.
+As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later.
+\\
+
 \section{Method}
 
-The wavelet methods for time series analysis are explained in a more detailed way in Percival \& Walden (2000 ~\cite{PER_2000}). \\
+\subsection{The Maximal Overlap Discrete Wavelet Transform}
 
-\textbf{Discrete Wavelet Transform}
-
-The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, ...
-, N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , ...
-, N - 1 \right)$.
-If we denote $J$ the level of the wavelet decomposition, and we have $N = n* 2^J$, where n is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
+The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$.
+If we denote $J$ the level of the wavelet decomposition, and we have $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
 , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$.
-\\
-
 Each wavelet vector $W_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$.
 The scaling vector $V_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$.
-\\
-
-We can also define for $j = 1 , ...
-, J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to scale $\tau_j = dt 2^{j - 1}$.
-Similarly, we can define for $j = 1 , ...
-, J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher.
+We can also define for $j = 1 , \cdots , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to scale $\tau_j = dt 2^{j - 1}$.
+Similarly, we can define for $j = 1 , \cdots , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher.
 Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:
 
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 
-One main advantage of the DWT is that it is an orthonormal transform, and thus we can write the analysis of variance (ANOVA):
+The DWT present several disadvantages.
+First, the length of the time series must be a multiple of $2^J$ where $J$ is the level of the DWT decomposition.
+Second, the time step of the wavelet vector $W_j$ is $dt 2^j$, which may not correspond to the time when some interesting phenomenon is visible on the original time series.
+Third, when we circularly shift the time series, the corresponding wavelet coefficients, details and smooths are not a circularly shifted version of the wavelet coefficients, details and smooths of the original time series.
+Thus, the values of the wavelet coefficients, details and smooths are strongly dependent on the time when we start experimentally gathering the data.
+Finally, when we filter the time series to obtain the details and smooths, we introduce a phase shift, which makes difficult to line up meaningfully the features of the MRA with the original time series.
+\\
 
-\begin{equation}
-\left\Vert X \right\Vert ^2 = \left\Vert W \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert W_j \right\Vert ^2 + \left\Vert V_J \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert D_j \right\Vert ^2 + \left\Vert S_J \right\Vert ^2
-\end{equation}
-
-Moreover, the DWT can be computed using $O \left( N \right)$ multiplications. \\
-
-However, the DWT present several disadvantages:
-
-\begin{itemize}
-	\item The length of the time series must be a multiple of $2^J$ where $J$ is the level of the DWT decomposition.
-	\item The time step of the wavelet vector $W_j$ is $dt 2^j$, which may not correspond to the time when some interesting phenomenon is visible on the original time series.
-	\item When we circularly shift the time series, the corresponding wavelet coefficients, details and smooths are not a circularly shifted version of the wavelet coefficients, details and smooths of the original time series.
-Thus, the values of the wavelet coefficients, details and smooths are strongly dependent on the time when we start experimentally gathering the data.
-	\item When we filter the time series to obtain the details and smooths, we introduce a phase shift, which makes difficult to line up meaningfully the features of the MRA with the original time series.
-\end{itemize}
-
-\textbf{Maximum Overlap Discrete Wavelet Transform}
-
-To get rid of these problems, we introduce the Maximum Overlap Discrete Wavelet Transform (MODWT).
+This is why we use instead the Maximal Overlap Discrete Wavelet Transform (MODWT).
 The MODWT transforms the time series $X_t \left( t = 0, ...
-, N - 1 \right)$ into J wavelet vectors $\widetilde{W}_j \left( j = 1 , ...
-, J \right)$ of length $N$ and a scaling vector $\widetilde{V}_J$ of length $N$.
+, N - 1 \right)$ into J wavelet vectors $\widetilde{W}_j \left( j = 1 ,  \cdots , J \right)$ of length $N$ and a scaling vector $\widetilde{V}_J$ of length $N$.
 As is the case for the DWT, each wavelet vector $\widetilde{W}_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$.
 The scaling vector $\widetilde{V}_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$.
-\\
-
 As is the case for the DWT, we can write the MRA:
 
+\begin{linenomath*}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath*}
 
-and the ANOVA:
+The MODWT of a time series can be defined for any length $N$.
+The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series.
+When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount.
+The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series.
+The wavelet methods for time series analysis are explained in a more detailed way in ~\citet{PER_2000}).
+\\
 
-\begin{equation}
-\left\Vert X \right\Vert ^2 = \sum_{j = 1}^{J} \left\Vert \widetilde{W}_j \right\Vert ^2 + \left\Vert \widetilde{V}_J \right\Vert ^2
-\end{equation}
-
-Now, we have the following properties:
-
-\begin{itemize}
-	\item The MODWT of a time series can be defined for any length $N$.
-	\item The time step of the wavelet vectors $\widetilde{W}_j$ and the scaling vector $\widetilde{V}_J$ is equal to the time step of the original time series.
-	\item When we circularly shift the time series, the corresponding wavelet vectors, scaling vector, details and smooths are shifted by the same amount.
-	\item The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series.
-\end{itemize}
-
-However, the MODWT has some disadvantages over the DWT:
-
-\begin{itemize}
-	\item The MODWT can only be computed using $O \left( N \log_2 N \right)$ multiplications.
-	\item We can no longer write the ANOVA for the details and smooths:
-
-\begin{equation}
-\left\Vert X \right\Vert ^2 \neq \sum_{j = 1}^{J} \left\Vert D_j \right\Vert ^2 + \left\Vert S_J \right\Vert ^2 \text{ and } \left\Vert \widetilde{W}_j \right\Vert ^2 \neq \left\Vert \widetilde{D}_j \right\Vert ^2
-\end{equation}
-
-\end{itemize}
+\subsection{Application to synthetic data}
 
@@
 To see the effect of the magnitude of the slow slip event, we use different values for $N = 2, 5, 10, 20$ days.
-Figure X shows the synthetics, the details of the wavelet decomposition for levels 1 to 8, and the smooth for the four durations of a slow slip event.
+Figure 1 shows the synthetics, the details of the wavelet decomposition for levels 1 to 8, and the smooth for the four durations of a slow slip event.
 \\
 
-%\begin{figure}
-%\noindent\includegraphics[width=\textwidth, trim={5cm 6cm 5cm 9cm},clip]{figures/500_DS.pdf}
-%\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
-%\label{pngfiguresample}
-%\end{figure}
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
+\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
+\label{pngfiguresample}
+\end{figure}
 
@@
 \\
+
+\subsection{MODWT of GPS and tremor data}
 

### 2021-08-08 src/tremor_MODWT.py

@@ -2,25 +2,25 @@
 Script to compute MODWT of cumulative tremor count
 """
 
-#import datetime
-#import matplotlib.cm as cm
+import datetime
+import matplotlib.cm as cm
 import matplotlib.pylab as pylab
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import pickle
 
-#from datetime import datetime
+from datetime import datetime
 from math import cos, floor, pi, sin, sqrt
-#from matplotlib.colors import Normalize
-#from scipy.io import loadmat
-from scipy.signal import butter, lfilter
+from matplotlib.colors import Normalize
+from scipy.io import loadmat
+from scipy.signal import butter, detrend, lfilter
 
 import date
 import DWT
 from MODWT import get_DS, get_scaling, pyramid
 
-def tremor_polynomial(lats, lons, radius_tremor, wavelet, J, degree):
+def compute_wavelets_tremor(lats, lons, radius_tremor, wavelet, J):
     """
     """
     # Read tremor files (A. Wech)
@@ -60,7 +60,7 @@ def tremor_polynomial(lats, lons, radius_tremor, wavelet, J, degree):
     e = 0.006694470
 
     # Time vector
-    time = pickle.load(open('tmp/time.pkl', 'rb'))
+    time = pickle.load(open('time.pkl', 'rb'))
 
     # Loop on latitude and longitude
     for index, (lat, lon) in enumerate(zip(lats, lons)):
@@ -91,271 +91,49 @@ def tremor_polynomial(lats, lons, radius_tremor, wavelet, J, degree):
 
         # Interpolate
         tremor = np.interp(time, np.sort(time_tremor), (1.0 / nt) * np.arange(0, len(time_tremor)))
+        tremor_detrend = detrend(tremor)
 
-        # Start figure
-#        params = {'xtick.labelsize':24,
-#                  'ytick.labelsize':24}
-#        pylab.rcParams.update(params)   
-#        fig = plt.figure(1, figsize=(10 * degree, 4 * (J + 3)))
+        # MODWT
+        (W, V) = pyramid(tremor_detrend, wavelet, J)
+        (D, S) = get_DS(tremor_detrend, W, wavelet, J)
 
-        # Loop on degree of polynomial
-        for d in range(1, degree + 1):
-            
[…]

### 2021-08-08 src/tremor_MODWT_transform.py

@@ -1,0 +1,375 @@
+"""
+Script to compute MODWT of cumulative tremor count
+"""
+
+#import datetime
+#import matplotlib.cm as cm
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import pickle
+
+#from datetime import datetime
+from math import cos, floor, pi, sin, sqrt
+#from matplotlib.colors import Normalize
+#from scipy.io import loadmat
+from scipy.signal import butter, lfilter
+
+import date
+import DWT
+from MODWT import get_DS, get_scaling, pyramid
+
+def tremor_polynomial(lats, lons, radius_tremor, wavelet, J, degree):
+    """
+    """
+    # Read tremor files (A. Wech)
+    data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
+    data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2010 = pd.read_csv('../data/tremor/tremor_events-2010-01-01T00 00 00-2010-12-31T23 59 59.csv')
+    data_2010['time '] = pd.to_datetime(data_2010['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2011 = pd.read_csv('../data/tremor/tremor_events-2011-01-01T00 00 00-2011-12-31T23 59 59.csv')
+    data_2011['time '] = pd.to_datetime(data_2011['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2012 = pd.read_csv('../data/tremor/tremor_events-2012-01-01T00 00 00-2012-12-31T23 59 59.csv')
+    data_2012['time '] = pd.to_datetime(data_2012['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2013 = pd.read_csv('../data/tremor/tremor_events-2013-01-01T00 00 00-2013-12-31T23 59 59.csv')
+    data_2013['time '] = pd.to_datetime(data_2013['time '], format='%Y-%m-%d %H:%M:%S')
+    data_2014 = pd.read_csv('../data/tremor/tremor_events-2014-01-01T00 00 00-2014-12-31T23 59 59.csv')
+    data_2014['time '] = pd.to_datetime(data_2014['time '], format='%Y-%m-%d %H:%M:%S')
[…]

### 2021-08-08 src/vespagram.py

@@ -97,9 +97,45 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         (W, V) = pyramid(disp, wavelet, J)
         (D, S) = get_DS(disp, W, wavelet, J)
         # Save wavelets into file
-        filename = 'tmp/' + dataset + '_' + station + '_' + direction + '.pkl'
+        filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
         pickle.dump([time, disp, W, V, D, S], open(filename, 'wb'))
 
+        # Start figure
+        params = {'xtick.labelsize':24,
+                  'ytick.labelsize':24}
+        pylab.rcParams.update(params)   
+        fig = plt.figure(1, figsize=(10, 3 * (J + 3)))
+
+        maxD = max([np.max(Dj) for Dj in D])
+        minD = min([np.min(Dj) for Dj in D])
+
+        # Plot data
+        plt.subplot2grid((J + 2, 1), (J + 1, 0))
+        plt.plot(time, disp, 'k', label='Data')
+        plt.xlabel('Time (years)', fontsize=24)
+        plt.xlim([2009.25, 2021.25])
+        plt.ylim([np.min(disp), np.max(disp)])
+        plt.legend(loc=3, fontsize=20)
+        # Plot details
+        for j in range(0, J):
+            plt.subplot2grid((J + 2, 1), (J - j, 0))
+            plt.plot(time, D[j], 'k', label='D' + str(j + 1))
+            plt.xlim([2009.25, 2021.25])
+            plt.ylim(minD, maxD)
+            plt.legend(loc=3, fontsize=20)
+        # Plot smooth
+        plt.subplot2grid((J + 2, 1), (0, 0))
+        plt.plot(time, S[J], 'k', label='S' + str(J))
+        plt.xlim([2009.25, 2021.25])
+        plt.ylim([np.min(disp), np.max(disp)])
+        plt.legend(loc=3, fontsize=20)
+        
+        # Save figure
+        plt.tight_layout()
+        plt.savefig('MODWT_GPS/' + dataset + '_' + station + '_' + \
+            direction + '.eps', format='eps')
+        plt.close(1)
+        
 def vesp_tremor(station_file, tremor_file, lats, lons, dataset, direction, \
[…]

### 2021-08-10 paper/agujournaltemplate.tex
Write methods

@@
 Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
-We thus want to demonstrate that the wavelet-based detection method can be applied to detect possible smaller (magnitude 5) slow slip events that may be currently undetected with standard methods.
+We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may be currently undetected with standard methods.
 \\
@@
 \subsection{MODWT of GPS and tremor data}
+
+We then applied the wavelet filtering to real GPS data.
+Figure 2 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth.
+In the data, we can see a sharp drop in displacement whenever there is a slow slip event.
+For levels 5 to 8, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data.
+We thus verify that the wavelet method can detect slow slip events.
+\\
+
+\begin{figure}
+\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
+\caption{Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.}
+\label{pngfiguresample}
+\end{figure}
+
+To increase the signal-to-noise ratio and be able to better detect slow slip events, we stack the signal over several GPS stations.
+We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal.
+We choose 16 points located on the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 3).
+Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations.
+We thus have a stacked detail for each level 1 to 8 of the wavelet decomposition.
+\\
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations.eps}
+\caption{GPS stations used in this study (black triangles).
+The black line represents the 40 km depth contour of the plate boundary model by ~\citet{PRE_2003}.
+The red triangles are the locations where we stack the GPS data.
+The small grey dots are all the tremor locations from the PNSN catalog.}
+\label{pngfiguresample}
+\end{figure}
+
+To compare slow slip events detected with GPS data and slow slip events detected with seismic data, we took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3.
+Then we computed the cumulative number of tremor within this circle.
+Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform.
+Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5).
+Contrary to what happens for the GPS data, we see a sharp increase in the data whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
+
+\begin{figure}
+\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
+\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.}
+\label{pngfiguresample}
+\end{figure}
 

### 2021-08-10 src/compare_tremor_GPS.py

@@ -573,10 +573,10 @@ def plot_tremor(lats, J, threshold):
     for index, lat in enumerate(lats):
 
         # Read MODWT of tremor
-        filename = 'tremor/loc' + str(index) + '.pkl'
+        filename = 'MODWT_tremor/tremor_' + str(index) + '.pkl'
         MODWT_tremor = pickle.load(open(filename, 'rb'))
         times_stacked = MODWT_tremor[0]
-        D_tremor = MODWT_tremor[5][J]
+        D_tremor = MODWT_tremor[4][J]
         
         # Figure
         if len(times_stacked) > 0:
@@ -611,6 +611,7 @@ def plot_tremor(lats, J, threshold):
         format(J + 1), fontsize=24)
     plt.savefig('tremor_detail_' + str(J + 1) + '.pdf', format='pdf')
     plt.close(1)
+
 if __name__ == '__main__':
 
     station_file = '../data/PANGA/stations.txt'
@@ -644,7 +645,7 @@ if __name__ == '__main__':
     # For GPS data
     # Level 8: 0.5 - Level 7: 0.4 - Level 6: 0.3 - Level 5: 0.3
     # For tremor data
-    # Level 8: 0.008 - Level 7: 0.006 - Level 6: 0.005
+    # Level 8: 0.008 - Level 7: 0.006 - Level 6: 0.005 - Level 5: 0.005
     threshold = 0.005
 
 #    plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J - 1, threshold)

### 2021-08-21 src/compare_tremor_GPS.py
compare DWT

@@ -612,6 +612,119 @@ def plot_tremor(lats, J, threshold):
     plt.savefig('tremor_detail_' + str(J + 1) + '.pdf', format='pdf')
     plt.close(1)
 
+def find_threshold(station_file, lats, lons, dataset, direction, radius_GPS, \
+    J, thresh_GPS, thresh_tremor):
+    """
+    """
+    # Read station file
+    stations = pd.read_csv(station_file, sep=r'\s{1,}', header=None, engine='python')
+    stations.columns = ['name', 'longitude', 'latitude']
+
+    a = 6378.136
+    e = 0.006694470
+
+    # Loop on latitude and longitude
+    for index, (lat, lon) in enumerate(zip(lats, lons)):
+
+        # Keep only stations in a given radius
+        dx = (pi / 180.0) * a * cos(lat * pi / 180.0) / sqrt(1.0 - e * e * \
+            sin(lat * pi / 180.0) * sin(lat * pi / 180.0))
+        dy = (3.6 * pi / 648.0) * a * (1.0 - e * e) / ((1.0 - e * e * \
+            sin(lat * pi / 180.0) * sin(lat * pi / 180.0)) ** 1.5)
+        x = dx * (stations['longitude'] - lon)
+        y = dy * (stations['latitude'] - lat)
+        stations['distance'] = np.sqrt(np.power(x, 2.0) + np.power(y, 2.0))
+        mask = stations['distance'] <= radius_GPS
+        sub_stations = stations.loc[mask].copy()
+        sub_stations.reset_index(drop=True, inplace=True)
+
+        # Wavelet vectors initialization
+        times = []
+        disps = []
+        Ws = []
+        Vs = []
+        Ds = []
+        Ss = []
+
+        # Read output files from wavelet transform
+        for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
+            filename = 'tmp/' + dataset + '_' + station + '_' + direction + '.pkl'
+            (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
+            if ((np.min(time) <= 2021.25) and (np.max(time) >= 2009.25)):
+                times.append(time)
+                disps.append(disp)
+                Ws.append(W)
+                Vs.append(V)
+                Ds.append(D)
+                Ss.append(S)
+
+        # Divide into time blocks
+        tblocks = []
+        for time in times:
+            tblocks.append(np.min(time))
+            tblocks.append(np.max(time))
[…]

### 2021-08-22 src/compare_tremor_GPS.py
Look for threshold

@@ -622,6 +622,9 @@ def find_threshold(station_file, lats, lons, dataset, direction, radius_GPS, \
 
     a = 6378.136
     e = 0.006694470
+    
+    nonzeros = np.zeros(len(lats))
+    equals = np.zeros(len(lats))
 
     # Loop on latitude and longitude
     for index, (lat, lon) in enumerate(zip(lats, lons)):
@@ -708,30 +711,30 @@ def find_threshold(station_file, lats, lons, dataset, direction, radius_GPS, \
         # Read MODWT of tremor
         filename = 'MODWT_tremor/tremor_' + str(index) + '.pkl'
         MODWT_tremor = pickle.load(open(filename, 'rb'))
-        D_tremor = MODWT_tremor[4][J]
+        time = MODWT_tremor[0]
+        D_tremor = - MODWT_tremor[4][J]
 
+        print(index, len(times_stacked))
         # Correlation
-        if len(times_stacked) > 0:
-            
-            pos_GPS = np.where(disps_stacked >= thresh_GPS)[0]
-            neg_GPS = np.where(disps_stacked <= - thresh_GPS)[0]
-            pos_tremor = np.where(D_tremor >= threshold)[0]
-            neg_tremor = np.where(D_tremor <= - threshold)[0]
-            input1 = np.zeros(len(times_stacked))
-            input1[pos_GPS] = 1
-            input1[neg_GPS] = -1
-            input2 = np.zeros(len(times_stacked))
-            input2[pos_tremor] = 1
-            input2[neg_tremor] = -1
+#        if len(times_stacked) > 0:
 
+#            pos_GPS = np.where(disps_stacked >= thresh_GPS)[0]
+#            neg_GPS = np.where(disps_stacked <= - thresh_GPS)[0]
+#            pos_tremor = np.where(D_tremor >= thresh_tremor)[0]
+#            neg_tremor = np.where(D_tremor <= - thresh_tremor)[0]
+#            input1 = np.zeros(len(times_stacked))
+#            input1[pos_GPS] = 1
+#            input1[neg_GPS] = -1
+#            input2 = np.zeros(len(times_stacked))
+#            input2[pos_tremor] = 1
+#            input2[neg_tremor] = -1
[…]

### 2021-08-22 src/compare_tremor_GPS.py
Found thresholds

@@ -459,7 +459,7 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
 
         # Read output files from wavelet transform
         for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
-            filename = 'tmp/' + dataset + '_' + station + '_' + direction + '.pkl'
+            filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
             if ((np.min(time) <= 2021.25) and (np.max(time) >= 2009.25)):
                 times.append(time)
@@ -651,7 +651,7 @@ def find_threshold(station_file, lats, lons, dataset, direction, radius_GPS, \
 
         # Read output files from wavelet transform
         for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
-            filename = 'tmp/' + dataset + '_' + station + '_' + direction + '.pkl'
+            filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
             if ((np.min(time) <= 2021.25) and (np.max(time) >= 2009.25)):
                 times.append(time)
@@ -714,22 +714,21 @@ def find_threshold(station_file, lats, lons, dataset, direction, radius_GPS, \
         time = MODWT_tremor[0]
         D_tremor = - MODWT_tremor[4][J]
 
-        print(index, len(times_stacked))
         # Correlation
-#        if len(times_stacked) > 0:
+        if len(times_stacked) > 0:
 
-#            pos_GPS = np.where(disps_stacked >= thresh_GPS)[0]
-#            neg_GPS = np.where(disps_stacked <= - thresh_GPS)[0]
-#            pos_tremor = np.where(D_tremor >= thresh_tremor)[0]
-#            neg_tremor = np.where(D_tremor <= - thresh_tremor)[0]
-#            input1 = np.zeros(len(times_stacked))
-#            input1[pos_GPS] = 1
[…]

### 2021-08-22 src/tremor_MODWT.py

@@ -20,9 +20,14 @@ import date
 import DWT
 from MODWT import get_DS, get_scaling, pyramid
 
-def compute_wavelets_tremor(lats, lons, radius_tremor, wavelet, J):
+def compute_wavelets_tremor(station_file, lats, lons, dataset, direction, \
+    radius_GPS, radius_tremor, wavelet, J):
     """
     """
+    # Read station file
+    stations = pd.read_csv(station_file, sep=r'\s{1,}', header=None, engine='python')
+    stations.columns = ['name', 'longitude', 'latitude']
+
     # Read tremor files (A. Wech)
     data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
     data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
@@ -59,11 +64,57 @@ def compute_wavelets_tremor(lats, lons, radius_tremor, wavelet, J):
     a = 6378.136
     e = 0.006694470
 
-    # Time vector
-    time = pickle.load(open('time.pkl', 'rb'))
-
     # Loop on latitude and longitude
     for index, (lat, lon) in enumerate(zip(lats, lons)):
+
+        # Keep only stations in a given radius
+        dx = (pi / 180.0) * a * cos(lat * pi / 180.0) / sqrt(1.0 - e * e * \
+            sin(lat * pi / 180.0) * sin(lat * pi / 180.0))
+        dy = (3.6 * pi / 648.0) * a * (1.0 - e * e) / ((1.0 - e * e * \
+            sin(lat * pi / 180.0) * sin(lat * pi / 180.0)) ** 1.5)
+        x = dx * (stations['longitude'] - lon)
+        y = dy * (stations['latitude'] - lat)
+        stations['distance'] = np.sqrt(np.power(x, 2.0) + np.power(y, 2.0))
+        mask = stations['distance'] <= radius_GPS
+        sub_stations = stations.loc[mask].copy()
+        sub_stations.reset_index(drop=True, inplace=True)
+
+        # Time vectors initialization
+        times = []
+
+        # Read output files from wavelet transform
+        for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
+            filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
[…]

### 2021-08-22 src/compare_tremor_GPS.py
Look for Michel events

@@ -428,7 +428,6 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     e = 0.006694470
 
     # Start figure
-#    fig = plt.figure(1, figsize=(16, 8))
     fig, ax = plt.subplots(figsize=(16, 8))
     params = {'xtick.labelsize':24,
               'ytick.labelsize':24}
@@ -519,24 +518,15 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
         # Figure
         if len(times_stacked) > 0:
             
-#            filename = 'level8/times' + str(index) + '_sup.txt'
             slowslip = np.where(np.abs(disps_stacked) >= threshold)[0]
             times_slowslip = times_stacked[slowslip]
             disps_slowslip = disps_stacked[slowslip]
-#            np.savetxt(filename, times_stacked[slowslip], fmt='%.4f')
-#            filename = 'level8/times' + str(index) + '_diff.txt'
             difference = np.diff(times_slowslip)
-#            np.savetxt(filename, np.stack((times_slowslip[:-1], difference)).T, fmt='%.4f')
-#            filename = 'level8/times' + str(index) + '_times.txt'
             jumps = np.where(difference > 1.5 / 365.25)[0]
             begin_jumps = np.insert(jumps + 1, 0, 0)
             end_jumps = np.append(jumps, len(times_slowslip) - 1)
-#            begin_times = np.insert(times_slowslip[jumps + 1], 0, times_slowslip[0])
-#            end_times = np.insert(times_slowslip[jumps], -1, times_slowslip[-1])
             begin_times = times_slowslip[begin_jumps]
             end_times = times_slowslip[end_jumps]
-#            np.savetxt(filename, np.stack((begin_times, end_times, \
-#                disps_slowslip[begin_jumps], disps_slowslip[end_jumps])).T, fmt='%.4f')
 
             for i in range(0, len(jumps) + 1):
                 x0 = begin_times[i]
@@ -625,6 +615,12 @@ def find_threshold(station_file, lats, lons, dataset, direction, radius_GPS, \
     
     nonzeros = np.zeros(len(lats))
     equals = np.zeros(len(lats))
+
+    # Start figure
+    fig, ax = plt.subplots(figsize=(16, 8))
[…]

### 2021-08-22 src/missing_values.py

@@ -1,0 +1,172 @@
+""" Script to look at MODWT of GPS data
+with defferent methods to fill in missing values """
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from math import log, sqrt
+
+import DWT, MODWT
+from MODWT import get_DS, get_scaling, inv_pyramid, pyramid
+
+# Choose the station
+station = 'PGC5'
+direction = 'lon'
+dataset = 'cleaned'
+filename = '../data/PANGA/' + dataset + '/' + station + '.' + direction
+# Load the data
+data = np.loadtxt(filename, skiprows=26)
+time = data[:, 0]
+disp = data[:, 1]
+error = data[:, 2]
+# Correct for the repeated value
+dt = np.diff(time)
+gap = np.where(dt < 1.0 / 365.0 - 0.001)[0]
+time[gap[0] + 1] = time[gap[0] + 2]
+time[gap[0] + 2] = 0.5 * (time[gap[0] + 2] + time[gap[0] + 3])
+# Look for gaps greater than 1 day
+days = 1
+dt = np.diff(time)
+gap = np.where(dt > days / 365.0 + 0.001)[0]
+# Select a subset of the data without gaps
+ibegin = 2943
+iend = 4333
+time = time[ibegin + 1 : iend + 1]
+disp = disp[ibegin + 1 : iend + 1]
+error = error[ibegin + 1 : iend + 1]
+sigE2 = np.mean(np.square(error))
+N = np.shape(disp)[0]
+
+# Parameters
+name = 'LA8'
+g = MODWT.get_scaling(name)
+L = len(g)
+J = 6
+nmax = 42
+nstep = 3
+filling = 'interpolation'
+
+# Set random seed
+np.random.seed(0)
+
+# Compute MODWT
+[W0, V0] = pyramid(disp, name, J)
+(nuH, nuG) = DWT.get_nu(name, J)
+(D0, S0) = get_DS(disp, W0, name, J)
+W0t = []
+for j in range(1, J + 1):
+    W0j = W0[j - 1]
+    deltaj = sqrt(2.0 * sigE2 * log(N) / (2.0 ** j))
+    W0jt = np.where(np.abs(W0j) >= deltaj, W0j, 0.0)
+    if (j == J):
+        V0t = np.where(np.abs(V0) >= deltaj, V0, 0.0)
+    W0t.append(W0jt)
+dispt = inv_pyramid(W0t, V0t, name, J)
+
+# Loop on missing values
[…]

### 2021-08-24 paper/agujournaltemplate.tex
add missing values to paper

@@
 
+The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings.
+To deal with the gaps in the GNSS recordings, we simply replace the missing values by Gaussian noise with mean zero and standard deviation equal to the standard deviation of the whole time series.
+We verify how the wavelet details may be affected by looking at a GPS time series without missing values and comparing the wavelet details with and without removing some data points.
+Station PGC5 has recorded during 1390 days between 2009 and 2013, without any missing values.
+We first computed the wavelet details without missing values.
+Then, we removed ten neighboring missing values, replaced them by Gaussian noise, and computed the wavelet details with the replaced values.
+Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values.
+We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition.
+However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels the details, and are barely visible for the levels we are mostly interested in (levels 6 to 8).
+We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events.
+\\
+
+\begin{figure}
+\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
+\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by Gaussian noise (red) for two locations of the missing values (left and right).
+Bottom to top: Corresponding eight details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by Gaussian noise (red).}
+\label{pngfiguresample}
+\end{figure}
+
 We then applied the wavelet filtering to real GPS data.
-Figure 2 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth.
+Figure 3 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth.
 In the data, we can see a sharp drop in displacement whenever there is a slow slip event.
@@
 We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal.
-We choose 16 points located on the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 3).
+We choose 16 points located on the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4).
 Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations.
@@
 Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform.
-Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5).
+Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5).
 Contrary to what happens for the GPS data, we see a sharp increase in the data whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
@@
 
-Explanation vespagram
+We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+The result is shown in the top panel of Figure 6, where each line represents one of the locations.
+To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold.
+To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6.
+We used the opposite of the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
 
-Correlation
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_8.pdf}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_8.pdf}
+\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
 

### 2021-08-24 src/missing_values.py

@@ -18,7 +18,6 @@ filename = '../data/PANGA/' + dataset + '/' + station + '.' + direction
 data = np.loadtxt(filename, skiprows=26)
 time = data[:, 0]
 disp = data[:, 1]
-error = data[:, 2]
 # Correct for the repeated value
 dt = np.diff(time)
 gap = np.where(dt < 1.0 / 365.0 - 0.001)[0]
@@ -33,15 +32,13 @@ ibegin = 2943
 iend = 4333
 time = time[ibegin + 1 : iend + 1]
 disp = disp[ibegin + 1 : iend + 1]
-error = error[ibegin + 1 : iend + 1]
-sigE2 = np.mean(np.square(error))
 N = np.shape(disp)[0]
 
 # Parameters
 name = 'LA8'
 g = MODWT.get_scaling(name)
 L = len(g)
-J = 6
+J = 8
 nmax = 42
 nstep = 3
 filling = 'interpolation'
@@ -51,23 +48,17 @@ np.random.seed(0)
 
 # Compute MODWT
 [W0, V0] = pyramid(disp, name, J)
-(nuH, nuG) = DWT.get_nu(name, J)
 (D0, S0) = get_DS(disp, W0, name, J)
-W0t = []
-for j in range(1, J + 1):
-    W0j = W0[j - 1]
-    deltaj = sqrt(2.0 * sigE2 * log(N) / (2.0 ** j))
-    W0jt = np.where(np.abs(W0j) >= deltaj, W0j, 0.0)
-    if (j == J):
-        V0t = np.where(np.abs(V0) >= deltaj, V0, 0.0)
-    W0t.append(W0jt)
-dispt = inv_pyramid(W0t, V0t, name, J)
 
 # Loop on missing values
 gaps = [605, 790]
-for gap in gaps:
-    # Loop on length of gap
-    for n in range(1, nmax, nstep):
+
+# Loop on length of gap
+for n in range(1, nmax, nstep):
+
+    plt.figure(1, figsize=(len(gaps) * 3, (J + 2) * 2))
+
+    for count, gap in enumerate(gaps):
         disp_interp = np.copy(disp)
         # Remove points, replace by interpolation or noise
         if (filling == 'interpolation'):
@@ -85,88 +76,30 @@ for gap in gaps:
         # Compute MODWT
         [W, V] = pyramid(disp_interp, name, J)
         (D, S) = get_DS(disp_interp, W, name, J)
-        Wt = []
[…]

### 2021-08-26 paper/agujournaltemplate.tex
Results part of paper

@@
 We used the opposite of the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
+Although the latitudinal extension of the events is not always the same for the GPS data and for the tremor data, we identify the same 10 events in both 8th wavelet decompositions for the 8th level: Summer 2010, Summer 2011, Summer 2012, Fall 2013, Summer-Fall 2014, Winter 2015-2016, Winter 2017, Spring 2018, Spring-Fall 2019, and Fall 2020 - Winter 2021.
+We can also see the end of an 11th event in Summer 2009.
+\\
 
@@
 Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
+
+Figures 7 and 8 show the same comparison between the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively.
+For the 7th level, we see the same events as for the 8th level, both for the GPS data and the tremor count data.
+The wavelet decomposition is more noisy for the GPS data in the earliest part of the time series, between 2010 and 2013, but it does not seem that there are more slow slip events visible in the 7th level.
+\\
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_7.pdf}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_7.pdf}
+\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
+
+For the 6th level detail, we see an additional event in Fall 2009 that is present both in the GPS and the tremor data.
+There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections.
+To summarize, all the 11 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 12 of the 15 events present on the 6th level detail of the wavelet decomposition are true detections.
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_6.pdf}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_6.pdf}
+\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
 \label{pngfiguresample}

### 2021-08-29 paper/agujournaltemplate.tex
Adde evnts from Micel et al to paper

@@
 To summarize, all the 11 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 12 of the 15 events present on the 6th level detail of the wavelet decomposition are true detections.
+\\
 
@@
 \section{Discussion}
+
+In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal.
+As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (levels 4 and 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog.
+Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events.
+Table 1 summarizes for each event its number as indicated in the catalog from ~\citet{MIC_2019}, the beginning and end times as indicated in the catalog from ~\citet{MIC_2019}, whether it is visible at the level 4 of the wavelet decomposition, whether it is visible at the level 5 of the wavelet decomposition, and whether it is part of a bigger magnitude 6 event.
+All 10 events that are sub-event of a bigger event are visible at both levels 4 and 5.
+However, this may be due because the bigger event is in at levels 6 to 8, and also at smaller time scales.
+For the 3 small events that are not part of a bigger event, only one is visible at both time scales, the other two are visible either for level either for level 5 of the wavelet decomposition.
+Therefore, it is difficult to conclude whether the method can indeed detect events of magnitude 5.
+
+ \begin{table}
+ \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
+ \centering
+ \begin{tabular}{c c c c c}
+ \hline
+ Event number & Time & Visible at level 4 & Visible at level 5 & Sub-event of bigger event \\
+ \hline
+ 25 & 2010.62-2010.67 & Yes & Yes & Yes \\
+ 29 & 2011.42-2011.45 & No & Yes & No \\
+ 31 & 2011.62-2011.68 & Yes & Yes & Yes \\
+ 32 & 2011.65-2011.68 & Yes & Yes & Yes \\
+ 35 & 2012.66-2012.72 & Yes & Yes & Yes \\
+ 42 & 2013.70-2013.78 & Yes & Yes & Yes \\
+ 44 & 2014.12-2014-20 & Yes & No & No \\
+ 45 & 2014.40-2014.48 & Yes & Yes & No \\
+ 49 & 2014.66-2014.71 & Yes & Yes & Yes \\
+ 52 & 2014.91-2014.95 & Yes & Yes & Yes \\
+ 57 & 2015.98-2016.08 & Yes & Yes & Yes \\
+ 60 & 2017.11-2017.15 & Yes & Yes & Yes \\
+ 61 & 2017.20-2017.24 & Yes & Yes & Yes \\
+ \hline
+ \end{tabular}
+ \end{table}
 

### 2021-08-29 src/vespagram.py

@@ -97,7 +97,7 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         (W, V) = pyramid(disp, wavelet, J)
         (D, S) = get_DS(disp, W, wavelet, J)
         # Save wavelets into file
-        filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
+        filename = 'MODWT_GPS_new/' + dataset + '_' + station + '_' + direction + '.pkl'
         pickle.dump([time, disp, W, V, D, S], open(filename, 'wb'))
 
         # Start figure
@@ -132,7 +132,7 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         
         # Save figure
         plt.tight_layout()
-        plt.savefig('MODWT_GPS/' + dataset + '_' + station + '_' + \
+        plt.savefig('MODWT_GPS_new/' + dataset + '_' + station + '_' + \
             direction + '.eps', format='eps')
         plt.close(1)
         
@@ -453,7 +453,7 @@ def vesp_map(station_file, tremor_file, tmin_tremor, tmax_tremor, lats, lons, \
 
         # Read output files from wavelet transform
         for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
-            filename = 'tmp/' + dataset + '_' + station + '_' + direction + '.pkl'
+            filename = 'MODWT_GPS_new/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
             if ((np.min(time) < tmax_GPS) and (np.max(time) > tmin_GPS)):
                 lon_stas.append(lon_sta)
@@ -539,7 +539,9 @@ def vesp_map(station_file, tremor_file, tmin_tremor, tmax_tremor, lats, lons, \
                 vespagram[:, (time_subset >= 2009.25) & (time_subset <= 2021.25)], cmap=plt.get_cmap('seismic'), \
                 norm=Normalize(vmin=-1.5, vmax=1.5))
 #                levels = np.linspace(-1.2, 1.2, 25))
-            plt.axvline(0.5 * (tmin_GPS + tmax_GPS), color='grey', linewidth=1)
[…]

### 2021-10-09 src/count_tremor.py
New tremor catalog

@@ -1,69 +1,0 @@
-import numpy as np
-import pandas as pd
-
-from datetime import datetime
-from math import cos, floor, pi, sin, sqrt
-
-lats = [47.20000, 47.30000, 47.40000, 47.50000, 47.60000, 47.70000, \
-        47.80000, 47.90000, 48.00000, 48.10000, 48.20000, 48.30000, 48.40000, \
-        48.50000, 48.60000, 48.70000]
-lons = [-122.74294, -122.73912, -122.75036, -122.77612, -122.81591, \
-        -122.86920, -122.93549, -123.01425, -123.10498, -123.20716, \
-        -123.32028, -123.44381, -123.57726, -123.72011, -123.87183, \
-        -124.03193]
-radius_tremor = 50
-    
-# Read tremor files (A. Wech)
-data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
-data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
-data_2010 = pd.read_csv('../data/tremor/tremor_events-2010-01-01T00 00 00-2010-12-31T23 59 59.csv')
-data_2010['time '] = pd.to_datetime(data_2010['time '], format='%Y-%m-%d %H:%M:%S')
-data_2011 = pd.read_csv('../data/tremor/tremor_events-2011-01-01T00 00 00-2011-12-31T23 59 59.csv')
-data_2011['time '] = pd.to_datetime(data_2011['time '], format='%Y-%m-%d %H:%M:%S')
-data_2012 = pd.read_csv('../data/tremor/tremor_events-2012-01-01T00 00 00-2012-12-31T23 59 59.csv')
-data_2012['time '] = pd.to_datetime(data_2012['time '], format='%Y-%m-%d %H:%M:%S')
-data_2013 = pd.read_csv('../data/tremor/tremor_events-2013-01-01T00 00 00-2013-12-31T23 59 59.csv')
-data_2013['time '] = pd.to_datetime(data_2013['time '], format='%Y-%m-%d %H:%M:%S')
-data_2014 = pd.read_csv('../data/tremor/tremor_events-2014-01-01T00 00 00-2014-12-31T23 59 59.csv')
-data_2014['time '] = pd.to_datetime(data_2014['time '], format='%Y-%m-%d %H:%M:%S')
[…]

### 2021-10-09 src/tremor_MODWT_longer.py

@@ -2,17 +2,22 @@
 Script to compute MODWT of cumulative tremor count
 """
 
+#import datetime
+#import matplotlib.cm as cm
 import matplotlib.pylab as pylab
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import pickle
 
+#from datetime import datetime
 from math import cos, floor, pi, sin, sqrt
+#from matplotlib.colors import Normalize
 from scipy.io import loadmat
 from scipy.signal import detrend
 
 import date
+#import DWT
 from MODWT import get_DS, pyramid
 
 def compute_wavelets_tremor(station_file, lats, lons, dataset, direction, \
@@ -23,36 +28,19 @@ def compute_wavelets_tremor(station_file, lats, lons, dataset, direction, \
     stations = pd.read_csv(station_file, sep=r'\s{1,}', header=None, engine='python')
     stations.columns = ['name', 'longitude', 'latitude']
 
-    # Read tremor files (A. Wech)
-    data_2009 = pd.read_csv('../data/tremor/tremor_events-2009-08-06T00 00 00-2009-12-31T23 59 59.csv')
-    data_2009['time '] = pd.to_datetime(data_2009['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2010 = pd.read_csv('../data/tremor/tremor_events-2010-01-01T00 00 00-2010-12-31T23 59 59.csv')
-    data_2010['time '] = pd.to_datetime(data_2010['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2011 = pd.read_csv('../data/tremor/tremor_events-2011-01-01T00 00 00-2011-12-31T23 59 59.csv')
-    data_2011['time '] = pd.to_datetime(data_2011['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2012 = pd.read_csv('../data/tremor/tremor_events-2012-01-01T00 00 00-2012-12-31T23 59 59.csv')
-    data_2012['time '] = pd.to_datetime(data_2012['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2013 = pd.read_csv('../data/tremor/tremor_events-2013-01-01T00 00 00-2013-12-31T23 59 59.csv')
-    data_2013['time '] = pd.to_datetime(data_2013['time '], format='%Y-%m-%d %H:%M:%S')
-    data_2014 = pd.read_csv('../data/tremor/tremor_events-2014-01-01T00 00 00-2014-12-31T23 59 59.csv')
[…]

### 2021-10-09 src/vespagram.py

@@ -4,7 +4,7 @@ Script to plot a vespagram-like figure of slow slip
 
 import cartopy.crs as ccrs
 import cartopy.io.shapereader as shapereader
-import datetime
+#import datetime
 import matplotlib.cm as cm
 import matplotlib.pylab as pylab
 import matplotlib.pyplot as plt
@@ -12,7 +12,7 @@ import numpy as np
 import pandas as pd
 import pickle
 
-from datetime import datetime
+#from datetime import datetime
 from math import cos, floor, pi, sin, sqrt
 from matplotlib.colors import Normalize
 from scipy.io import loadmat
@@ -97,7 +97,7 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         (W, V) = pyramid(disp, wavelet, J)
         (D, S) = get_DS(disp, W, wavelet, J)
         # Save wavelets into file
-        filename = 'MODWT_GPS_new/' + dataset + '_' + station + '_' + direction + '.pkl'
+        filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
         pickle.dump([time, disp, W, V, D, S], open(filename, 'wb'))
 
         # Start figure
@@ -113,26 +113,26 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         plt.subplot2grid((J + 2, 1), (J + 1, 0))
         plt.plot(time, disp, 'k', label='Data')
         plt.xlabel('Time (years)', fontsize=24)
-        plt.xlim([2009.25, 2021.25])
+        plt.xlim([2006.0, 2021.5])
         plt.ylim([np.min(disp), np.max(disp)])
         plt.legend(loc=3, fontsize=20)
         # Plot details
         for j in range(0, J):
             plt.subplot2grid((J + 2, 1), (J - j, 0))
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
-            plt.xlim([2009.25, 2021.25])
+            plt.xlim([2006.0, 2021.5])
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
         # Plot smooth
         plt.subplot2grid((J + 2, 1), (0, 0))
         plt.plot(time, S[J], 'k', label='S' + str(J))
-        plt.xlim([2009.25, 2021.25])
+        plt.xlim([2006.0, 2021.5])
[…]

### 2021-10-09 paper/agujournaltemplate.tex
Fix noise question

@@
 The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings.
-To deal with the gaps in the GNSS recordings, we simply replace the missing values by Gaussian noise with mean zero and standard deviation equal to the standard deviation of the whole time series.
+To deal with the gaps in the GNSS recordings, we simply replace the missing values by the sum of a straight line and a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series.
+The straight line starts at the mean of the five days before the gap and ends at the mean of the five days after the gap.
 We verify how the wavelet details may be affected by looking at a GPS time series without missing values and comparing the wavelet details with and without removing some data points.
@@
 We first computed the wavelet details without missing values.
-Then, we removed ten neighboring missing values, replaced them by Gaussian noise, and computed the wavelet details with the replaced values.
+Then, we removed ten neighboring missing values, replaced them by the sum of the straight line and the Gaussian noise component, and computed the wavelet details with the replaced values.
 Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values.
 We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition.
-However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels the details, and are barely visible for the levels we are mostly interested in (levels 6 to 8).
+However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels the details, and are barely visible for the levels we are mostly interested in (levels 6 and above).
 We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events.
@@
 \noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by Gaussian noise (red) for two locations of the missing values (left and right).
-Bottom to top: Corresponding eight details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by Gaussian noise (red).}
+\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
+Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).}
 \label{pngfiguresample}

### 2021-10-09 src/missing_values.py

@@ -1,5 +1,5 @@
 """ Script to look at MODWT of GPS data
-with defferent methods to fill in missing values """
+with different methods to fill in missing values """
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -38,10 +38,10 @@ N = np.shape(disp)[0]
 name = 'LA8'
 g = MODWT.get_scaling(name)
 L = len(g)
-J = 8
+J = 10
 nmax = 42
 nstep = 3
-filling = 'interpolation'
+filling = 'both'
 
 # Set random seed
 np.random.seed(0)
@@ -63,16 +63,19 @@ for n in range(1, nmax, nstep):
         # Remove points, replace by interpolation or noise
         if (filling == 'interpolation'):
             for i in range(0, n):
-                disp_interp[gap + i] = disp_interp[gap - 1] + (disp_interp[gap + n] \
-                    - disp_interp[gap - 1]) * (i + 1) / (n + 1)
+                before = np.mean(disp_interp[gap - 6 : gap - 1])
+                after = np.mean(disp_interp[gap + n : gap + n + 5])
+                disp_interp[gap + i] = before + (after - before) * (i + 1) / (n + 1)
         elif (filling == 'noise'):
             sigma = np.std(disp)
             disp_interp[gap : (gap + n)] = np.random.normal(0.0, sigma, n)
         else:
             sigma = np.std(disp)
             for i in range(0, n):
-                disp_interp[gap + i] = disp_interp[gap - 1] + (disp_interp[gap + n] \
-                    - disp_interp[gap - 1]) * (i + 1) / (n + 1) + np.random.normal(0.0, sigma, 1)
+                before = np.mean(disp_interp[gap - 6 : gap - 1])
+                after = np.mean(disp_interp[gap + n : gap + n + 5])
+                disp_interp[gap + i] = before + (after - before) * (i + 1) / (n + 1) + \
+                    np.random.normal(0.0, sigma, 1)
         # Compute MODWT
         [W, V] = pyramid(disp_interp, name, J)
         (D, S) = get_DS(disp_interp, W, name, J)

### 2021-10-09 src/synthetics.py
Remake MODWT

@@ -11,13 +11,13 @@ import MODWT
 # Time between events
 timesteps = [100, 200, 500, 1000]
 # Durations of slow slip events
-durations = [2, 5, 10, 20]
+durations = [5, 10, 20, 40]
 # MODWT wavelet filter
 name = 'LA8'
 # Duration of recording
 N = 2000
 # MODWT level
-J = 8
+J = 10
 
 # Create time vector
 time = np.arange(0, N)
@@ -47,7 +47,7 @@ for timestep in timesteps:
         maxD = max([np.max(Dj) for Dj in D])
         minD = min([np.min(Dj) for Dj in D])
         # Plot data
-        ax = plt.subplot2grid((J + 2, len(durations)), (J + 1, i))
+        ax = plt.subplot2grid((J + 2, len(durations)), (0, i))
         plt.plot(time, disp, 'k', label='Data')
         plt.xlabel('Time (days)', fontsize=20)
         plt.ylim(-2.0, 2.0)
@@ -56,7 +56,7 @@ for timestep in timesteps:
             ax.axes.yaxis.set_ticks([])
         # Plot details
         for j in range(0, J):
-            ax = plt.subplot2grid((J + 2, len(durations)), (J - j, i))
+            ax = plt.subplot2grid((J + 2, len(durations)), (j + 1, i))
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
@@ -64,7 +64,7 @@ for timestep in timesteps:
             if i != 0:
                 ax.axes.yaxis.set_ticks([])
         # Plot smooth
-        ax = plt.subplot2grid((J + 2, len(durations)), (0, i))
+        ax = plt.subplot2grid((J + 2, len(durations)), (J + 1, i))
         plt.plot(time, S[J], 'k', label='S' + str(J))
         plt.ylim(-2.0, 2.0)
         plt.legend(loc=3, fontsize=20)
@@ -75,8 +75,6 @@ for timestep in timesteps:
         plt.title(title, fontsize=20)
 
     # Save figure
-#    title = 'Time between events = ' + str(timestep)
-#    plt.suptitle(title, fontsize=30)
     plt.tight_layout()
     plt.savefig('synthetics/' + str(timestep) + '_DS.eps', format='eps')
     plt.close(1)

### 2021-10-09 src/vespagram.py

@@ -60,12 +60,14 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
     # Read GPS data and compute wavelet transform
     for station in subset['name']:
         filename = '../data/PANGA/' + dataset + '/' + station + '.' + direction
+
         # Load the data
         data = np.loadtxt(filename, skiprows=26)
         time = data[:, 0]
         disp = data[:, 1]
         error = data[:, 2]
         sigma = np.std(disp)
+
         # Correct for the repeated values
         dt = np.diff(time)
         gap = np.where(dt < 1.0 / 365.0 - 0.0001)[0]
@@ -81,23 +83,36 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
                     and (time[gap[i] + 3] - time[gap[i] + 2] < 2.0 / 365.0 + 0.0001)):
                         time[gap[i] + 1] = time[gap[i] + 2]
                         time[gap[i] + 2] = 0.5 * (time[gap[i] + 2] + time[gap[i] + 3])
+
         # Look for gaps greater than 1 day
         days = 2
         dt = np.diff(time)
         gap = np.where(dt > days / 365.0 - 0.0001)[0]
         duration = np.round((time[gap + 1] - time[gap]) * 365).astype(np.int)
+
         # Fill the gaps by interpolation
         for j in range(0, len(gap)):
             time = np.insert(time, gap[j] + 1, \
                 time[gap[j]] + np.arange(1, duration[j]) / 365.0)
-            disp = np.insert(disp, gap[j] + 1, \
-                np.random.normal(0.0, sigma, duration[j] - 1))
+            if gap[j] >= 4:
+                before = np.mean(disp[gap[j] - 4 : gap[j] + 1])
+            else:
+                before = np.mean(disp[0 : gap[j] + 1])
+            if len(disp) >= gap[j] + 6:
+                after = np.mean(disp[gap[j] + 1 : gap[j] + 6])
+            else:
+                after = np.mean(disp[gap[j] + 1 :])
+            disp_interp = before + (after - before) * (np.arange(0, duration[j] - 1) + 1) / duration[j] + \
+                np.random.normal(0.0, sigma, duration[j] - 1)
+            disp = np.insert(disp, gap[j] + 1, disp_interp)
             gap[j + 1 : ] = gap[j + 1 : ] + duration[j] - 1
+
         # MODWT
[…]

### 2021-10-09 src/synthetics.py
Remade synthetics

@@ -49,7 +49,6 @@ for timestep in timesteps:
         # Plot data
         ax = plt.subplot2grid((J + 2, len(durations)), (0, i))
         plt.plot(time, disp, 'k', label='Data')
-        plt.xlabel('Time (days)', fontsize=20)
         plt.ylim(-2.0, 2.0)
         plt.legend(loc=3, fontsize=20)
         if i != 0:
@@ -60,15 +59,14 @@ for timestep in timesteps:
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
-            ax.axes.xaxis.set_ticks([])
             if i != 0:
                 ax.axes.yaxis.set_ticks([])
         # Plot smooth
         ax = plt.subplot2grid((J + 2, len(durations)), (J + 1, i))
         plt.plot(time, S[J], 'k', label='S' + str(J))
         plt.ylim(-2.0, 2.0)
+        plt.xlabel('Time (days)', fontsize=20)
         plt.legend(loc=3, fontsize=20)
-        ax.axes.xaxis.set_ticks([])
         if i != 0:
             ax.axes.yaxis.set_ticks([])
         title = 'Duration of event = ' + str(duration) + ' days'

### 2021-10-10 paper/agujournaltemplate.tex
remake plots

@@
 To reproduce the ground displacement observed on the longitudinal component of GPS stations in Cascadia, we divide each period into two parts: In the first part of duration $T - N$, the displacement is linearly increasing and corresponds to the secular plate motion in the eastern direction; in the second part of duration $N$, the displacement is linearly decreasing and corresponds to a slow slip event on a reverse fault at depth triggering a ground displacement in the western direction.
-To see the effect of the magnitude of the slow slip event, we use different values for $N = 2, 5, 10, 20$ days.
-Figure 1 shows the synthetics, the details of the wavelet decomposition for levels 1 to 8, and the smooth for the four durations of a slow slip event.
+To see the effect of the magnitude of the slow slip event, we use different values for $N = 5, 10, 20, 40$ days.
+Figure 1 shows the synthetics, the details of the wavelet decomposition for levels 1 to 10, and the smooth for the four durations of a slow slip event.
 \\
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
+\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
 \caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
@@
 The ramp-like signal is transformed through the wavelet filtering into a waveform with first a positive peak and then a negative peak.
-The width of the waveform increases with the scale level.
+The shape of the waveform is the same for every level of the wavelet decomposition, but the width of the waveform increases with the scale level.
 For the 8th level of the wavelet decomposition, the width of the waveform is nearly as large as the time between two events.
-We do not show details at larger scales as the corresponding waveforms would start to merge two contiguous events together, and make the wavelet decomposition less interpretable.
-For an event of duration 2 days, the wavelet details at levels higher than 2 have a larger amplitude than the wavelet detail at level 1.
+At larger scales, the waveforms start to merge two contiguous events together, and make the wavelet decomposition less interpretable.
 For an event of duration 5 days, the wavelet details at levels higher than 3 have a larger amplitude than the wavelet details at lower scales.
-For an event of duration 10 days, the wavelet details at levels higher than 5 have a larger amplitude than the wavelet details at lower scales.
-For an event of duration 20 days, the wavelet details at levels higher than 6 have a larger amplitude than the wavelet details at lower scales.
+For an event of duration 10 days, the wavelet details at levels higher than 4 have a larger amplitude than the wavelet details at lower scales.
+For an event of duration 20 days, the wavelet details at levels higher than 5 have a larger amplitude than the wavelet details at lower scales.
+For an event of duration 40 days, the wavelet details at levels higher than 6 have a larger amplitude than the wavelet details at lower scales.
 Thus, the scale levels at which an event is being seen in the wavelet details give us an indication about the duration (and the magnitude) of the slow slip event.
-We expect the big slow slip events of magnitude 6-7 that lasts about 10 days to start being visible at the level 5 of the wavelet decomposition, but to not be noticeable at lower time scales.
+We expect the big slow slip events of magnitude 6-7 that last several weeks to start being visible at the level 5 of the wavelet decomposition, but to not be noticeable at lower time scales.
 \\

### 2021-10-10 src/compare_tremor_GPS.py

@@ -458,9 +458,9 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
 
         # Read output files from wavelet transform
         for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
-            filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
+            filename = 'MODWT_GPS_longer/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
-            if ((np.min(time) <= 2021.25) and (np.max(time) >= 2009.25)):
+            if ((np.min(time) <= 2021.5) and (np.max(time) >= 2006.0)):
                 times.append(time)
                 disps.append(disp)
                 Ws.append(W)
@@ -512,8 +512,8 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
         disps_stacked = np.concatenate(disps_stacked)
 
         # Filter displacement
-        disps_stacked = disps_stacked[(times_stacked  >= 2009.25) & (times_stacked <= 2021.25)]
-        times_stacked = times_stacked[(times_stacked  >= 2009.25) & (times_stacked  <= 2021.25)]
+        disps_stacked = disps_stacked[(times_stacked  >= 2006.0) & (times_stacked <= 2021.5)]
+        times_stacked = times_stacked[(times_stacked  >= 2006.0) & (times_stacked  <= 2021.5)]
 
         # Figure
         if len(times_stacked) > 0:
@@ -538,7 +538,7 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
 
             plt.plot(times_stacked, lat + 0.1 * disps_stacked, color='black')
 
-    plt.xlim([2009.25, 2021.25])
+    plt.xlim([2006.0, 2021.5])
     plt.xlabel('Time (years)', fontsize=24)
     plt.xticks(fontsize=24)
     plt.ylim([min(lats) - 0.15, max(lats) + 0.15])
[…]

### 2021-10-10 src/vespagram.py

@@ -112,7 +112,7 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         (D, S) = get_DS(disp, W, wavelet, J)
 
         # Save wavelets into file
-        filename = 'MODWT_GPS_longer/' + dataset + '_' + station + '_' + direction + '.pkl'
+        filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
         pickle.dump([time, disp, W, V, D, S], open(filename, 'wb'))
 
         # Start figure
@@ -128,26 +128,26 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         plt.subplot2grid((J + 2, 1), (J + 1, 0))
         plt.plot(time, disp, 'k', label='Data')
         plt.xlabel('Time (years)', fontsize=24)
-        plt.xlim([2006.0, 2021.5])
+        plt.xlim([2009.25, 2021.25])
         plt.ylim([np.min(disp), np.max(disp)])
         plt.legend(loc=3, fontsize=20)
         # Plot details
         for j in range(0, J):
             plt.subplot2grid((J + 2, 1), (J - j, 0))
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
-            plt.xlim([2006.0, 2021.5])
+            plt.xlim([2009.25, 2021.25])
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
         # Plot smooth
         plt.subplot2grid((J + 2, 1), (0, 0))
         plt.plot(time, S[J], 'k', label='S' + str(J))
-        plt.xlim([2006.0, 2021.5])
+        plt.xlim([2009.25, 2021.25])
         plt.ylim([np.min(disp), np.max(disp)])
         plt.legend(loc=3, fontsize=20)
         
         # Save figure
         plt.tight_layout()
-        plt.savefig('MODWT_GPS_longer/' + dataset + '_' + station + '_' + \
+        plt.savefig('MODWT_GPS/' + dataset + '_' + station + '_' + \
             direction + '.eps', format='eps')
         plt.close(1)
         

### 2021-10-10 src/compare_tremor_GPS.py
Remake figures

@@ -417,7 +417,7 @@ def plot_correlations():
         plt.savefig('events_' + str(level) + '.pdf', format='pdf')
         plt.close(1)        
 
-def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, threshold):
+def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, threshold, events):
     """
     """
     # Read station file
@@ -501,7 +501,10 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 if (len(indices) > 0):
                     nsta = nsta + 1
                     tj = time[indices]
-                    Dj = D[J][indices]
+#                    Dj = D[J][indices]
+                    Dj = np.zeros(len(indices))
+                    for j in J:
+                        Dj = Dj + D[j - 1][indices]
                     Dj_interp = np.interp(time_subset, tj, Dj)
                     Dj_stacked =  Dj_stacked + Dj_interp
             Dj_stacked = Dj_stacked / nsta
@@ -528,6 +531,9 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
             begin_times = times_slowslip[begin_jumps]
             end_times = times_slowslip[end_jumps]
 
+            for event in events:
+                plt.axvline(event, color='grey', linewidth=1)
+
             for i in range(0, len(jumps) + 1):
                 x0 = begin_times[i]
                 dx = end_times[i] - begin_times[i]
@@ -544,12 +550,15 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     plt.ylim([min(lats) - 0.15, max(lats) + 0.15])
     plt.ylabel('Latitude', fontsize=24)
     plt.yticks(fontsize=24)
-    plt.title('Detail at level {:d} of MODWT of GPS data'. \
-        format(J + 1), fontsize=24)
-    plt.savefig('GPS_longer_detail_' + str(J + 1) + '.pdf', format='pdf')
+#    plt.title('Detail at level {:d} of MODWT of GPS data'. \
+#        format(J + 1), fontsize=24)
+    plt.title('Detail at levels {} of MODWT of GPS data'. \
[…]

### 2021-10-10 paper/agujournaltemplate.tex
Send paper to advisors

@@
 
-\title{=enter title here=}
+\title{Detection of slow slip events using wavelet analysis of GNSS recordings}
 
@@
 
-\authors{=list all authors here=}
+\authors{A. Ducellier \affil{1}, K. C. Creager \affil{1}, D. A. Schmidt \affil{1}}
 
@@
 
-\affiliation{=number=}{=Affiliation Address=}
+\affiliation{1}{University of Washington}
 %(repeat as many times as is necessary)
@@
 
-\correspondingauthor{=name=}{=email address=}
+\correspondingauthor{Ariane Ducellier}{ducela@uw.edu}
 
@@
 \begin{abstract}
-enter abstract here
-
+Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks.
+Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles.
+In many places, tectonic tremor is also observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in GNSS noise.
+However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip.
+Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next.
+In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations.
+As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform.
+In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events.
+We assume that there is a slow slip event whenever there is a peak in the wavelet signal.
+We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
+The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
+~\citep{MIC_2019}).
 \end{abstract}
@@
 In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data.
-For instance, ~\citet{AGU_2009} computed the GPS-estimated moment release for 23 ETS events in Cascadia between 1997 and 2008.
-Simultaneously, they computed the cumulative number of hours of tectonic tremor recorded for each event.
+For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor.
+For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded.
 They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8.
-For all these events, at least 50 hours of tectonic tremor where observed simultaneously with the GPS deformation.
-However, many smaller bursts of tremor of duration 1 to 50 hours were also observed in between the big ETS events.
-Based on the relationship between slow slip moment and number of hours of tremor, ~\citet{AGU_2009} suggested that smaller slow slip events of magnitude 5-6 may occur simultaneously with the tremor bursts without being detectable in the GPS data.
+They also observed many smaller bursts of tremor of duration 1 to 50 hours in between the big ETS events, without any detectable signal in the GPS data.
+However, based on the relationship between slow slip moment and number of hours of tremor for bigger events, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with the tremor bursts.
+This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}).
 \\
 
-~\citet{FRA_2016} transformed the GPS time series into daily increments of surface motion by computing the first order differentiation of the time series.
-He then discarded the daily increments observed during known big slow slip events, and focused on the inter-events period.
-He divided the daily increments into two groups: the first group contains days when slow seismicity (tectonic tremor and LFEs) is detected, the second group contains days when the numbers of LFEs (for Guerrero) or tremor (for Cascadia) is lower than a threshold.
-He then stacked separately the two groups of daily increments and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
+~\citet{FRA_2016} divided GPS time series observations into two groups: the first group contains days when slow seismicity (tectonic tremor and LFEs) is detected, the second group contains days when the numbers of tremor or LFEs is lower than a threshold.
+He then stacked separately the two groups of observations and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
 He also observed a cumulative displacement in the southern direction (for Guerrero) and the western direction (for Cascadia) corresponding to the release period when tremor and LFEs are observed.
-This reverse displacement corresponds to smaller slow slip events not directly observable in the GPS data.
+He was thus able to observe a reverse displacement corresponding to smaller slow slip events not directly observable in the GPS data.
 \\
 
-However, in other subduction zones such as New Zealand, there is no clear relationship between tremor and slow slip occurrence and these methods cannot be applied.
+However, in other subduction zones such as New Zealand, there is no clear relationship between tremor and slow slip occurrence and these methods cannot be applied to detect smaller slow slip events that produce a GNSS signal with an amplitude too small compared to the noise.
 We thus need other methods to be able to better detect and quantify slow slip.
@@
 Wavelet methods have been widely used for geophysical applications ~\citep{KUM_1997}.
-However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events.
-\\
-
-~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011.
-The noise in the tide gauges data is associated with tides, and ocean and atmospheric noise on multiple timescales (a few days for storms to decades for oscillations between ocean basins), and is assumed to be coherent between each of the four tidal gauges studied.
-On the contrary, the uplift due to ETS events should be different at each tidal gauge.
-They first removed the tides using NOAA hourly harmonic tidal predictions.
-They then removed the residual noise using a method based on the DWT.
-More precisely, the authors applied a DWT to each of the four sites studied, and to the average of the four sites.
-Then, for each level of the DWT decomposition, they carried out a linear regression between the detail for one site and the detail for the average of the four sites.
-This process gives a coefficient for each level and for each site.
-They then constructed a noise signal for each site by multiplying the coefficient from the linear regression by the detail of the average over the four sites, and summing for all levels.
-The noise signal thus obtained was then removed from the time series.
-They then stacked multiple events to obtain an average event uplift rate, aligning the 12 ETS events using exact timing from GPS data.
-A difference in uplift between the two tidal gauges at Port Angeles and Port Townsend was then clearly seen in the stacked time series.
-Finally, the authors removed the long-term uplift rate and the long-term sea level rise to obtain an average inter-event uplift rate.
-They found that the inter-event deformation at a site is equal and opposite to the deformation during an ETS event, suggesting that ETS events are, on average, releasing the strain accumulated between ETS events.
+However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events ~\citep{SZE_2008,OHT_2010,WEI_2012,ALB_2019}.
 \\
@@
 ~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005.
-They stabilized the GPS time series using a reference set of stations from stable North America.
-They then modelled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, and Heaviside step functions corresponding to earthquakes and hardware upgrades.
-The linear system was then solved using a weighted QR decomposition.
-Finally, they applied a Gaussian wavelet transform to the residual time series to get the exact timing of the slow slip at each GPS station.
-The succeeding wavelet basis functions are increasingly sensitive to temporal localization of a given signal, and the onset of faulting appears on the wavelet spectrum as an amplitude spike present over all frequencies.
-The offset for each slow slip event was then used to invert for the slow slip at depth by assuming a thrust fault slip at each subfault of the plate boundary.
-An equivalent moment magnitude was thus obtained.
+They modeled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, Heaviside step functions corresponding to earthquakes and hardware upgrades, and a residual signal.
+They then applied a Gaussian wavelet transform to the residual time series to get the exact timing of the slow slip at each GPS station.
+The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains.
+A sharp change in the signal will be localized and seen at all levels of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform.
 \\
 
-Finally, instead of using wavelets in the time domain, ~\citet{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events.
+Instead of using wavelets in the time domain, ~\citet{OHT_2010} used 2D wavelet functions in the spatial domain to detect slow slip events.
 They designed the Network Stain Filter (NSF) to detect transient deformation signals from large-scale geodetic arrays.
-Contrary to their previous work on the Network Inversion Filter (NIF), there is no need to specify potential sources of deformation.
 They modeled the position of the GPS station by the sum of the secular velocity, a spatially coherent field, site-specific noise, reference frame errors, and observation errors.
-The spatial displacement field is modeled by the sum of basis wavelets (the Deslauriers-Dubuc wavelet of degree 3) with time-varying weights.
-The transient is considered to be nearly steady-sate, so that it has spatial weights for the displacement and the velocity, but the acceleration is modeled by a random walk with a time-varying variance.
-All the time varying coefficients are estimated using Kalman filtering, and the optimization problem is regularized with the spatial sum of the transient strain rate field.
+The spatial displacement field is modeled by the sum of basis wavelets with time-varying weights.
 Their method has been successfully used to detect a transient event in the Boso peninsula, Japan, and a slow slip event in the Alaska subduction zone ~\citep{WEI_2012}.
@@
 
-In this study, we use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia.
+Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011.
+Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events.
+The noise component is assumed to be coherent between all tidal gauges, while the uplift signal is different provided that the gauges are far enough from each other.
+By stacking the tidal records, the uplift signals cancel each other while the noise signal is amplified.
+By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal.
+The authors were then able to clearly see a difference in uplift between the two tidal gauges at Port Angeles and Port Townsend.
+\\
+
+In our study, we use a similar approach with a different reasoning.
+We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered.
+We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking.
+Finally, we suppose that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
+We use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia.
 Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
@@
 However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio.
-The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times eries with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}.
-For the seismic data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}.
+The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}.
+For the seismic data, we used the tremor catalog from \textcolor{red}{reference for the catalog.
+The following is to be modified.
+the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}.
 Tremor were detected and located using waveform envelope correlation and clustering and a centroid location is available for every given five-minute time window when tremor was detected.
-As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later.
-\\
+As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later.} \\
 
@@
 The details and smooths are associated with a zero phase filter, making it easy to line up meaningfully the features of the MRA with the original time series.
-The wavelet methods for time series analysis are explained in a more detailed way in ~\citet{PER_2000}).
+The wavelet methods for time series analysis are explained in a more detailed way in ~\citep{PER_2000}).
 \\
@@
 Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations.
-We thus have a stacked detail for each level 1 to 8 of the wavelet decomposition.
+We thus have a stacked detail for each level 1 to 10 of the wavelet decomposition.
 \\
@@
 We used the opposite of the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
-Although the latitudinal extension of the events is not always the same for the GPS data and for the tremor data, we identify the same 10 events in both 8th wavelet decompositions for the 8th level: Summer 2010, Summer 2011, Summer 2012, Fall 2013, Summer-Fall 2014, Winter 2015-2016, Winter 2017, Spring 2018, Spring-Fall 2019, and Fall 2020 - Winter 2021.
-We can also see the end of an 11th event in Summer 2009.
+In the tremor catalog from \textcolor{red}{reference?}, there are 17 tremor events with more than 150 hours of tremor recorded.
+The events are summarized in Table 1.
+The time of the event is the start date plus half the duration of the event.
+Although the latitudinal extension of the events is not always the same for the GPS data and for the tremor data, we identify the same 13 events in both 8th wavelet decompositions for the 8th level: January 2007, May 2008, May 2009, August 2010, August 2011, September 2012, September 2013, August-November 2014, January 2016, March 2017, June 2018, March-November 2019, and October 2020-January 2021.
+Although there are two events in the tremor catalog in August 2014 and November 2014, these two events are not distinguishable in the 8th level details and look more like a single event slowly propagating from South to North.
+The same phenomenon is observed in 2019 when two tremor events in March and November 2019 are merged into a single event propagating slowly from South to North.
+In 2020-2021, the wavelet decomposition of the tremor shows one event in the south in October-November 2020 and one event in the North in January 2021, but in the wavelet decomposition of the GPS data, these three events look like a single event propagating slowly from South to North.
 \\
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_8.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.pdf}
 
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_8.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.pdf}
 \caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
@@
 For the 7th level, we see the same events as for the 8th level, both for the GPS data and the tremor count data.
-The wavelet decomposition is more noisy for the GPS data in the earliest part of the time series, between 2010 and 2013, but it does not seem that there are more slow slip events visible in the 7th level.
+The wavelet decomposition is more noisy for the GPS data between 2010 and 2012, but it does not seem that there are more slow slip events visible in the 7th level.
 \\
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_7.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.pdf}
 
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_7.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.pdf}
 \caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
@@
 
-For the 6th level detail, we see an additional event in Fall 2009 that is present both in the GPS and the tremor data.
+For the 6th level detail, we see an additional event in the South in Fall 2009 that is present both in the GPS and the tremor data.
+It may correspond to the northern extent of a big ETS event occurring in Fall 2009 south of the study area (event 19 in the ~\citet{MIC_2019} catalog).
 There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections.
-To summarize, all the 11 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 12 of the 15 events present on the 6th level detail of the wavelet decomposition are true detections.
+To summarize, all the 13 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
 \\
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_detail_6.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.pdf}
 
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_detail_6.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.pdf}
 \caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
@@
 
+ \begin{table}
+     \caption{Big Episodic Tremor and Slip events.
+The duration and the number of tremor are from the tremor catalog of \textcolor{red}{reference?}.
+The event number and the magnitude are from the slow slip catalog of ~\citet{MIC_2019}.}
+ \centering
+ \begin{tabular}{c c c c c}
+ \hline
+ Time & Duration (days) & Number of tremor (hours) & Event number & Magnitude \\
+ \hline
+ 2007.06 & 28 & 398 & 3 & 6.68 \\
+ 2008.36 & 25 & 402 & 10 & 6.56 \\
+ 2009.35 & 24 & 248 & 16 & 6.49 \\
+ 2010.63 & 29 & 518 & 24 & 6.54 \\
+ 2011.60 & 37 & 479 & 30 & 6.47 \\
+ 2012.72 & 37 & 620 & 34 & 6.54 \\
+ 2013.71 & 27 & 423 & 41 & 6.58 \\
+ 2014.65 & 15 & 190 & 48 & 6.03 \\
+ 2014.89 & 38 & 385 & 51 & 6.40 \\
+ 2016.11 & 43 & 421 & 54 & 6.79 \\
+ 2017.23 & 19 & 279 & 59 & 6.61 \\
+ 2018.49 & 22 & 381 & & \\
+ 2019.23 & 34 & 195 & & \\
+ 2019.88 & 16 & 205 & & \\
+ 2020.79 & 26 & 193 & & \\
+ 2020.86 & 12 & 162 & & \\
+ 2021.09 & 14 & 230 & & \\
+ \hline
+ \end{tabular}
+ \end{table}
+
 \section{Discussion}
@@
 In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal.
-As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (levels 4 and 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog.
+As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (level 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog.
 Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events.
-Table 1 summarizes for each event its number as indicated in the catalog from ~\citet{MIC_2019}, the beginning and end times as indicated in the catalog from ~\citet{MIC_2019}, whether it is visible at the level 4 of the wavelet decomposition, whether it is visible at the level 5 of the wavelet decomposition, and whether it is part of a bigger magnitude 6 event.
-All 10 events that are sub-event of a bigger event are visible at both levels 4 and 5.
-However, this may be due because the bigger event is in at levels 6 to 8, and also at smaller time scales.
-For the 3 small events that are not part of a bigger event, only one is visible at both time scales, the other two are visible either for level either for level 5 of the wavelet decomposition.
-Therefore, it is difficult to conclude whether the method can indeed detect events of magnitude 5.
+Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event.
+\\
 
@@
  \centering
- \begin{tabular}{c c c c c}
+ \begin{tabular}{c c c c}
  \hline
- Event number & Time & Visible at level 4 & Visible at level 5 & Sub-event of bigger event \\
+ Time & Event number & Magnitude & Sub-event of bigger event \\
  \hline
- 25 & 2010.62-2010.67 & Yes & Yes & Yes \\
- 29 & 2011.42-2011.45 & No & Yes & No \\
- 31 & 2011.62-2011.68 & Yes & Yes & Yes \\
- 32 & 2011.65-2011.68 & Yes & Yes & Yes \\
- 35 & 2012.66-2012.72 & Yes & Yes & Yes \\
- 42 & 2013.70-2013.78 & Yes & Yes & Yes \\
- 44 & 2014.12-2014-20 & Yes & No & No \\
- 45 & 2014.40-2014.48 & Yes & Yes & No \\
- 49 & 2014.66-2014.71 & Yes & Yes & Yes \\
- 52 & 2014.91-2014.95 & Yes & Yes & Yes \\
- 57 & 2015.98-2016.08 & Yes & Yes & Yes \\
- 60 & 2017.11-2017.15 & Yes & Yes & Yes \\
- 61 & 2017.20-2017.24 & Yes & Yes & Yes \\
+ 2007.06 & 1 & 5.64 & Yes \\
+ 2007.08 & 2 & 5.91 & Yes \\
+ 2008.38 & 11 & 5.50 & Yes \\
+ 2009.16 & 14 & 5.50 & No \\
+ 2009.36 & 17 & 5.32 & Yes \\
+ 2010.63 & 25 & 5.76 & Yes \\
+ 2011.66 & 31 & 5.61 & Yes \\
+ 2011.66 & 32 & 5.32 & Yes \\
+ 2012.69 & 35 & 5.56 & Yes \\
+ 2013.74 & 42 & 5.71 & Yes \\
+ 2014.69 & 49 & 5.31 & Yes \\
+ 2014.93 & 52 & 5.39 & Yes \\
+ 2016.03 & 57 & 5.80 & Yes \\
+ 2017.13 & 60 & 5.43 & Yes \\
+ 2017.22 & 61 & 5.37 & Yes \\
  \hline
@@
 
+Figure 9 shows the 5th level detail wavelet decomposition of the GPS data.
+Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.pdf}
+\caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.}
+\label{pngfiguresample}
+\end{figure}
+
+All 14 events that are sub-event of a bigger event are visible at level 5.
+However, this may be due because the bigger event are clearly seen at levels 6 to 8, and also at smaller time scales.
+The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition.
+However, some other events that are not in ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
+Therefore, it is difficult to make the difference between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
+
 \section{Conclusion}
+
+In this paper, we have used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+We used detrended GNSS data, applied the MODWT transform and stack the wavelet details over several nearby GNSS stations.
+As an independent check on the timing of slow slip events, we also computed the cumulative number of tremor in the vicinity of the GNSS stations, detrended this signal, and applied the MODWT transform.
+In both time series, we could then see simultaneous waveforms whose timing corresponds to the timing of slow slip events.
+We assumed that there is a slow slip event whenever there is a peak in the wavelet signal.
+We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
+The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
+~\citep{MIC_2019}).
+We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to make the difference between true detections and false detections.
 
@@
 \acknowledgments
-Enter acknowledgments, including your data availability statement, here.
+This work was funded by the grant from the National Science Foundation XXX.
+A.D.
+would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington.
+The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA}.
+Figure 4 was done using GMT ~\citep{WES_1991}.
+The Python scripts used to analyze the data and make the  figures can be found on the  first author's Github account.
 

### 2021-10-10 src/compare_tremor_GPS.py

@@ -417,7 +417,7 @@ def plot_correlations():
         plt.savefig('events_' + str(level) + '.pdf', format='pdf')
         plt.close(1)        
 
-def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, threshold, events):
+def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, threshold, events, small_events):
     """
     """
     # Read station file
@@ -432,6 +432,11 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     params = {'xtick.labelsize':24,
               'ytick.labelsize':24}
     pylab.rcParams.update(params)
+
+    for event in events:
+        plt.axvline(event, color='red', linewidth=2)
+    for event in small_events:
+        plt.axvline(event, color='blue', linewidth=1)
 
     # Loop on latitude and longitude
     for index, (lat, lon) in enumerate(zip(lats, lons)):
@@ -501,10 +506,10 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 if (len(indices) > 0):
                     nsta = nsta + 1
                     tj = time[indices]
-#                    Dj = D[J][indices]
-                    Dj = np.zeros(len(indices))
-                    for j in J:
-                        Dj = Dj + D[j - 1][indices]
+                    Dj = D[J][indices]
+#                    Dj = np.zeros(len(indices))
+#                    for j in J:
+#                        Dj = Dj + D[j - 1][indices]
                     Dj_interp = np.interp(time_subset, tj, Dj)
                     Dj_stacked =  Dj_stacked + Dj_interp
             Dj_stacked = Dj_stacked / nsta
@@ -531,34 +536,31 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
             begin_times = times_slowslip[begin_jumps]
             end_times = times_slowslip[end_jumps]
 
-            for event in events:
-                plt.axvline(event, color='grey', linewidth=1)
-
-            for i in range(0, len(jumps) + 1):
-                x0 = begin_times[i]
[…]

### 2021-10-21 paper/paper_bssa.tex
Reformat paper for BSSA

--- 2021-10-10 paper/agujournaltemplate.tex
@@
-%% March 2018
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%
-% Step 1: Set the \documentclass
-%
-% There are two options for article format:
-%
-% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
-% The draft option produces double spaced output.
-%
-
-%% To submit your paper:
-\documentclass[draft]{agujournal2018}
+\documentclass{article}
 
@@
 \usepackage{amssymb}
+\usepackage{authblk}
+\usepackage{color}
+\usepackage{graphicx}
+\usepackage{lineno}
+\usepackage{natbib}
+\usepackage{url}
 
-\usepackage{apacite}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\usepackage{lineno}
 \linenumbers
-%%%%%%%
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
 
-\draftfalse
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Enter journal name here}
+\title{Detection of slow slip events using wavelet analysis of GNSS recordings}
+\author[1]{Ariane Ducellier}
+\author[1]{Kenneth C. Creager}
+\author[1]{David A. Schmidt}
+\affil[1]{University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
+\date{}
 
@@
 
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
+\maketitle
 
-% Example: \title{This is a test title}
+\section*{Key points}
 
-\title{Detection of slow slip events using wavelet analysis of GNSS recordings}
+\begin{itemize}
+\item We use a wavelet-based signal processing method to detect slow slip events in GNSS data.
+\item There is a good correlation between events detected with GNSS data and events detected with seismic data.
+\item The method could be applied in regions and no tremor are detected in conjonction with slow slip events.
+\end{itemize}
 
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
+\newpage
 
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
+\section*{Abstract}
 
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-\authors{A. Ducellier \affil{1}, K. C. Creager \affil{1}, D. A. Schmidt \affil{1}}
-
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-
-\affiliation{1}{University of Washington}
-%(repeat as many times as is necessary)
-
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
-\correspondingauthor{Ariane Ducellier}{ducela@uw.edu}
-
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 100 characters or less with no special characters or punctuation
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 100 characters or less with no special characters or punctuation
-% \end{keypoints}
-
-\begin{keypoints}
-\item enter point 1 here
-\item enter point 2 here
-\item enter point 3 here
-\end{keypoints}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT
-%
-% A good abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
-\begin{abstract}
-Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks.
-Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles.
+Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by GNSS networks.
+Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years).
 In many places, tectonic tremor is also observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in GNSS noise.
 However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip.
-Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next.
-In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+In this paper, we use the Maximal Overlap Discrete Wavelet Transform (MODWT) to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
 We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations.
@@
 The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
-~\citep{MIC_2019}).
-\end{abstract}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
+~\citet{MIC_2019}).
 
@@
 However, based on the relationship between slow slip moment and number of hours of tremor for bigger events, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with the tremor bursts.
-This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}).
+\\ %This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}).
 \\
@@
 
-\begin{figure}
-\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
-\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
-\label{pngfiguresample}
-\end{figure}
-
 The ramp-like signal is transformed through the wavelet filtering into a waveform with first a positive peak and then a negative peak.
@@
 
-\begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).}
-\label{pngfiguresample}
-\end{figure}
-
 We then applied the wavelet filtering to real GPS data.
@@
 \\
-
-\begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
-\caption{Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.}
-\label{pngfiguresample}
-\end{figure}
 
@@
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations.eps}
-\caption{GPS stations used in this study (black triangles).
-The black line represents the 40 km depth contour of the plate boundary model by ~\citet{PRE_2003}.
-The red triangles are the locations where we stack the GPS data.
-The small grey dots are all the tremor locations from the PNSN catalog.}
-\label{pngfiguresample}
-\end{figure}
-
 To compare slow slip events detected with GPS data and slow slip events detected with seismic data, we took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3.
@@
 Contrary to what happens for the GPS data, we see a sharp increase in the data whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
-
-\begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
-\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.}
-\label{pngfiguresample}
-\end{figure}
 
@@
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.pdf}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.pdf}
-\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
-\label{pngfiguresample}
-\end{figure}
-
 Figures 7 and 8 show the same comparison between the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively.
@@
 \\
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.pdf}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.pdf}
-\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
-\label{pngfiguresample}
-\end{figure}
 
@@
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.pdf}
+\section{Discussion}
 
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.pdf}
-\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
-\label{pngfiguresample}
-\end{figure}
+To better evaluate the number of true and false detections, we convert the wavelet details into binary time series.
+If the absolute value of the wavelet detail is higher that a threshold, we replace the value by 1 (for positive values) or -1 (for negative values), otherwise we replace the value by 0.
+We do this on both the wavelet details of the GPS data and of the tremor data.
+Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a true detection (true positive TP).
+If the GPS and the tremor time series have opposite signs, or if the absolute value of the GPS time series is 1 but the value of the tremor time series is 0, we have a false detection (false positive FP).
+If both time series take the value 0, we do not have detection (true negative TN).
+If the GPS time series take the value 0, but the absolute value of the tremor time series is 1, we miss a detection (false negative).
+We then define the sensitivity (true positive rate) and the specificity (equal to 1 minus the false positive rate) as:
 
- \begin{table}
+\begin{equation}
+\begin{aligned}
+\text{sensitivity} &= \frac{TP}{TP + FN} \\
+\text{specificity} &= \frac{TN} {TN + FP}
+\end{aligned}
+\end{equation}
+
+We can then evaluate the quality of the detections obtained with our method by plotting a receiver operating characteristic curve (ROC curve).
+The ROC curve is widely use for binary classification problems, and is plotted by varying the values of the threshold (here the thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
+If the classification was made randomly, all the points would fall on the first diagonal.
+If the classifier was perfect, the corresponding point would fall on the top left corner of the graph with true positive rate equal to 1 and false positive rate equal to 0.
+The bigger the area under the curve, the better the classifier is.
+\\
+
+As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a binary time series.
+We apply this transform to both the GPS and the tremor time series with varying thresholds.
+We then plot the ROC curve on Figure 8.
+The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10.
+We can see that there is a trade-off between sensitivity and specificity.
+If we decrease the false positive rate, we also decrease the number of true events detected.
+If we increase the number of true events detected, we also increase the false positive rate.
+\\
+
+In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal.
+As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (level 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog.
+Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events.
+Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event.
+\\
+
+Figure 11 shows the 5th level detail wavelet decomposition of the GPS data.
+Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
+
+All 14 events that are sub-event of a bigger event are visible at level 5.
+However, this may be due because the bigger event are clearly seen at levels 6 to 8, and also at smaller time scales.
+The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition.
+However, some other events that are not in ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
+Therefore, it is difficult to make the difference between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
+
+\section{Conclusion}
+
+In this paper, we have used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+We used detrended GNSS data, applied the MODWT transform and stack the wavelet details over several nearby GNSS stations.
+As an independent check on the timing of slow slip events, we also computed the cumulative number of tremor in the vicinity of the GNSS stations, detrended this signal, and applied the MODWT transform.
+In both time series, we could then see simultaneous waveforms whose timing corresponds to the timing of slow slip events.
+We assumed that there is a slow slip event whenever there is a peak in the wavelet signal.
+We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
+The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
+~\citep{MIC_2019}).
+We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to make the difference between true detections and false detections.
+
+\section*{Data and Resources}
+
+The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA} \url{http://www.panga.cwu.edu/}.
+\textcolor{red}{Reference for tremor catalog}.
+The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/slowslip}.
+Figure 4 was done using GMT ~\citep{WES_1991}.
+
+\section*{Acknowledgements}
+
+This work was funded by the grant from the National Science Foundation XXX.
+A.D.
+would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington.
+
+\section*{Declaration of Competing Interests}
+
+The authors declare no competing interests.
+
+\bibliographystyle{plainnat}
+\bibliography{bibliography}
+
+\newpage
+
+\section*{Addresses}
+
+Ariane Ducellier. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310. \\
+
+Kenneth C. Creager. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310. \\
+
+David A. Schmidt. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310.
+
+\newpage
+
+\section*{Tables}
+
+\begin{table}[hbt!]
 \caption{Big Episodic Tremor and Slip events.
@@
 
-\section{Discussion}
-
-In addition to the magnitude 6 events discussed above,  ~\citet{MIC_2019} have also identified several magnitude 5 events using a variational Bayesian Independent Component Analysis (vbICA) decomposition of the signal.
-As we expect smaller magnitude events to be more visible at smaller time scales of the wavelet decomposition (level 5), we verify for all these events whether a signal can be seen at the same time as the time given in their catalog.
-Most of these magnitude 5 events are also sub-events of bigger magnitude 6 events.
-Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event.
-\\
-
- \begin{table}
+ \begin{table}[hbt!]
  \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
@@
 
-Figure 9 shows the 5th level detail wavelet decomposition of the GPS data.
-Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
+\newpage
+
+\section*{Figure captions}
+
+\begin{itemize}
+
+\item Figure1.
+Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).
+
+\item Figure2.
+Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
+Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).
+
+\item Figure3. Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.
+
+\item Figure 4.
+GPS stations used in this study (black triangles).
+The black line represents the 40 km depth contour of the plate boundary model by ~\citet{PRE_2003}.
+The red triangles are the locations where we stack the GPS data.
+The small grey dots are all the tremor locations from the PNSN catalog.
+
+\item Figure 5. Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.
+
+\item Figure 6.
+Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+
+\item Figure 7.
+Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+
+\item Figure 8.
+Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+
+\item Figure 9.
+Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the sum of the 6th, 7th and 8th level details of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+
+\item Figure 10.
+ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
+
+\item Figure 11.
+Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+
+\end {itemize}
+
+\newpage
+
+\section*{Figures}
+
+\begin{figure}
+\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
+\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
+\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
+Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
+\caption{Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations.eps}
+\caption{GPS stations used in this study (black triangles).
+The black line represents the 40 km depth contour of the plate boundary model by ~\citet{PRE_2003}.
+The red triangles are the locations where we stack the GPS data.
+The small grey dots are all the tremor locations from the PNSN catalog.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
+\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.eps}
+\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
+\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.pdf}
+\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6-7-8.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
+\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Bottom: Opposite of the sum of the 6th, 7th and 8th level details of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/ROC_6-7-8.eps}
+\caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.}
+\label{pngfiguresample}
+\end{figure}
 
@@
 
-All 14 events that are sub-event of a bigger event are visible at level 5.
-However, this may be due because the bigger event are clearly seen at levels 6 to 8, and also at smaller time scales.
-The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition.
-However, some other events that are not in ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
-Therefore, it is difficult to make the difference between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
-
-\section{Conclusion}
-
-In this paper, we have used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
-We used detrended GNSS data, applied the MODWT transform and stack the wavelet details over several nearby GNSS stations.
-As an independent check on the timing of slow slip events, we also computed the cumulative number of tremor in the vicinity of the GNSS stations, detrended this signal, and applied the MODWT transform.
-In both time series, we could then see simultaneous waveforms whose timing corresponds to the timing of slow slip events.
-We assumed that there is a slow slip event whenever there is a peak in the wavelet signal.
-We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
-The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
-~\citep{MIC_2019}).
-We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to make the difference between true detections and false detections.
-
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURE
-%
-% \begin{figure}[h]
-% \centering
-% when using pdflatex, use pdf file:
-% \includegraphics[natwidth=800px,natheight=600px]{figsamp.pdf}
-%
-% when using dvips, use .eps file:
-% \includegraphics[natwidth=800px,natheight=600px]{figsamp.eps}
-%
-% \caption{Short caption}
-% \label{figone}
-%  \end{figure}
-%
-% We recommend that you provide the native width and height (natwidth, natheight) of your figures.
-% Specifying native dimensions ensures that your figures are properly scaled
-%
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-%
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$}
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% >>>>	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% 	All funding sources related to this work from all authors
-%
-% 	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% 	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-%
-% It is also the appropriate place to thank colleagues and other contributors.
-% AGU does not normally allow dedications.
-
-\acknowledgments
-This work was funded by the grant from the National Science Foundation XXX.
-A.D.
-would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington.
-The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA}.
-Figure 4 was done using GMT ~\citep{WES_1991}.
-The Python scripts used to analyze the data and make the  figures can be found on the  first author's Github account.
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% BibTeX is preferred:
-%
-\bibliography{bibliography.bib}
-%
-% don't specify bibliographystyle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Please use ONLY \citet and \citep for reference citations.
-% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
-%% Example \citet and \citep:
-%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
-%  \citet{Ghel00}, and \citet{Leit74}.
-
-%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
-%  \citep{Ghel00, Leit74}.
-
-%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.
-
 \end{document}
-
-More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2021-10-21 src/compare_tremor_GPS.py

@@ -434,9 +434,9 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     pylab.rcParams.update(params)
 
     for event in events:
-        plt.axvline(event, color='red', linewidth=2)
-    for event in small_events:
-        plt.axvline(event, color='blue', linewidth=1)
+        plt.axvline(event, color='black', linewidth=1)
+#    for event in small_events:
+#        plt.axvline(event, color='blue', linewidth=1)
 
     # Loop on latitude and longitude
     for index, (lat, lon) in enumerate(zip(lats, lons)):
@@ -485,7 +485,6 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
         # Initializations
         times_stacked = []
         disps_stacked = []
-        vesps = []
 
         # Loop on time blocks
         for t in range(0, len(tblocks) - 1):
@@ -506,10 +505,10 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 if (len(indices) > 0):
                     nsta = nsta + 1
                     tj = time[indices]
-                    Dj = D[J][indices]
-#                    Dj = np.zeros(len(indices))
-#                    for j in J:
-#                        Dj = Dj + D[j - 1][indices]
+#                    Dj = D[J][indices]
+                    Dj = np.zeros(len(indices))
+                    for j in J:
+                        Dj = Dj + D[j - 1][indices]
                     Dj_interp = np.interp(time_subset, tj, Dj)
                     Dj_stacked =  Dj_stacked + Dj_interp
             Dj_stacked = Dj_stacked / nsta
@@ -531,33 +530,35 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
             disps_slowslip = disps_stacked[slowslip]
             difference = np.diff(times_slowslip)
             jumps = np.where(difference > 1.5 / 365.25)[0]
-            begin_jumps = np.insert(jumps + 1, 0, 0)
-            end_jumps = np.append(jumps, len(times_slowslip) - 1)
-            begin_times = times_slowslip[begin_jumps]
-            end_times = times_slowslip[end_jumps]
+            
+            if len(jumps > 0):
[…]

### 2021-10-22 src/compare_tremor_GPS.py
Corrected bug

@@ -505,10 +505,10 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 if (len(indices) > 0):
                     nsta = nsta + 1
                     tj = time[indices]
-#                    Dj = D[J][indices]
-                    Dj = np.zeros(len(indices))
-                    for j in J:
-                        Dj = Dj + D[j - 1][indices]
+                    Dj = D[J][indices]
+#                    Dj = np.zeros(len(indices))
+#                    for j in J:
+#                        Dj = Dj + D[j - 1][indices]
                     Dj_interp = np.interp(time_subset, tj, Dj)
                     Dj_stacked =  Dj_stacked + Dj_interp
             Dj_stacked = Dj_stacked / nsta
@@ -553,12 +553,12 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     plt.ylim([min(lats) - 0.15, max(lats) + 0.15])
     plt.ylabel('Latitude', fontsize=24)
     plt.yticks(fontsize=24)
-#    plt.title('Detail at level {:d} of MODWT of GPS data'. \
-#        format(J + 1), fontsize=24)
-    plt.title('Details at levels {} of MODWT of GPS data'. \
-        format(J), fontsize=24)
-#    plt.savefig('GPS_longer_detail_' + str(J + 1) + '.eps', format='eps')
-    plt.savefig('GPS_longer_detail.eps', format='eps')
+    plt.title('Detail at level {:d} of MODWT of GPS data'. \
+        format(J + 1), fontsize=24)
+#    plt.title('Details at levels {} of MODWT of GPS data'. \
+#        format(J), fontsize=24)
+    plt.savefig('GPS_longer_detail_' + str(J + 1) + '.pdf', format='pdf')
+#    plt.savefig('GPS_longer_detail.eps', format='eps')
     plt.close(1)
 
 def plot_tremor(lats, J, threshold, events, small_events):
@@ -583,10 +583,10 @@ def plot_tremor(lats, J, threshold, events, small_events):
         filename = 'MODWT_tremor_longer/tremor_' + str(index) + '.pkl'
         MODWT_tremor = pickle.load(open(filename, 'rb'))
         times_stacked = MODWT_tremor[0]
-#        D_tremor = MODWT_tremor[4][J]
[…]

### 2021-10-24 paper/paper_bssa.tex
Remake figures

@@
 Figures 7 and 8 show the same comparison between the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively.
-For the 7th level, we see the same events as for the 8th level, both for the GPS data and the tremor count data.
+The events are harder to see in the 7th level than in the 8th level, both for the GPS data and the tremor count data.
 The wavelet decomposition is more noisy for the GPS data between 2010 and 2012, but it does not seem that there are more slow slip events visible in the 7th level.
@@
 There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections.
-To summarize, all the 13 events present on the 7th and 8th level details of the wavelet decomposition are true detections, 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
+To summarize, all the 13 events present on the 8th level detail of the wavelet decomposition are true detections, 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
 \\
@@
 We apply this transform to both the GPS and the tremor time series with varying thresholds.
-We then plot the ROC curve on Figure 8.
+We then plot the ROC curve on Figure 9.
 The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10.
@@
 If we increase the number of true events detected, we also increase the false positive rate.
+In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75, and the sensitivity is the highest possible.
 \\

### 2021-10-24 src/compare_tremor_GPS.py

@@ -505,10 +505,10 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 if (len(indices) > 0):
                     nsta = nsta + 1
                     tj = time[indices]
-                    Dj = D[J][indices]
-#                    Dj = np.zeros(len(indices))
-#                    for j in J:
-#                        Dj = Dj + D[j - 1][indices]
+#                    Dj = D[J][indices]
+                    Dj = np.zeros(len(indices))
+                    for j in J:
+                        Dj = Dj + D[j - 1][indices]
                     Dj_interp = np.interp(time_subset, tj, Dj)
                     Dj_stacked =  Dj_stacked + Dj_interp
             Dj_stacked = Dj_stacked / nsta
@@ -553,12 +553,12 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     plt.ylim([min(lats) - 0.15, max(lats) + 0.15])
     plt.ylabel('Latitude', fontsize=24)
     plt.yticks(fontsize=24)
-    plt.title('Detail at level {:d} of MODWT of GPS data'. \
-        format(J + 1), fontsize=24)
-#    plt.title('Details at levels {} of MODWT of GPS data'. \
-#        format(J), fontsize=24)
-    plt.savefig('GPS_longer_detail_' + str(J + 1) + '.pdf', format='pdf')
-#    plt.savefig('GPS_longer_detail.eps', format='eps')
+#    plt.title('Detail at level {:d} of MODWT of GPS data'. \
+#        format(J + 1), fontsize=24)
+    plt.title('Details at levels {} of MODWT of GPS data'. \
+        format(J), fontsize=24)
+#    plt.savefig('GPS_longer_detail_' + str(J + 1) + '.eps', format='eps')
+    plt.savefig('GPS_longer_detail.eps', format='eps')
     plt.close(1)
 
 def plot_tremor(lats, J, threshold, events, small_events):
@@ -583,10 +583,10 @@ def plot_tremor(lats, J, threshold, events, small_events):
         filename = 'MODWT_tremor_longer/tremor_' + str(index) + '.pkl'
         MODWT_tremor = pickle.load(open(filename, 'rb'))
         times_stacked = MODWT_tremor[0]
[…]

### 2021-10-24 paper/paper_bssa.tex
Start including comments from DS

@@
 \begin{itemize}
-\item We use a wavelet-based signal processing method to detect slow slip events in GNSS data.
-\item There is a good correlation between events detected with GNSS data and events detected with seismic data.
-\item The method could be applied in regions and no tremor are detected in conjonction with slow slip events.
+\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
+\item There is a good correlation between detections using GNSS data and independent detections using seismic data.
+\item The method could be applied in regions where no tremor are detected in conjunction with slow slip events.
 \end{itemize}
@@
 
-Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by GNSS networks.
-Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years).
-In many places, tectonic tremor is also observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in GNSS noise.
+In many places, tectonic tremor is observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in Global Navigation Satellite System (GNSS) noise.
 However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip.
-In this paper, we use the Maximal Overlap Discrete Wavelet Transform (MODWT) to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next.
+In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
 We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations.
@@
 In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events.
-We assume that there is a slow slip event whenever there is a peak in the wavelet signal.
-We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
-The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
-~\citet{MIC_2019}).
+We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal.
+We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data for northern Cascadia.
+The wavelet-based detection method detects well events of magnitude higher than 6 as determined by independent event catalogs (e.g.
+~\citep{MIC_2019}).
 
@@
 Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks.
-As with ordinary earthquakes, slow slip events are caused by slip on a fault, such as the plate boundary between a tectonic plate subducting under another tectonic plate.
-However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes, and they have a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}.
+As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate.
+However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes.
+They have a relatively short recurrence time (months to years) compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}.
 A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion.
-Slow slip events have been observed in many subduction zones, such as Cascadia, Nankai (southwest Japan), Alaska, Costa Rica, Mexico, and New Zealand ~\citep{BER_2011,AUD_2016}.
+Slow slip events have been observed in many places, such as Cascadia, Nankai (southwest Japan), Alaska, Costa Rica, Mexico, and New Zealand ~\citep{BER_2011,AUD_2016}.
 \\
 
-In many places, tectonic tremor are also observed in relation to slow slip.
+In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places.
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
-Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, that is small magnitude earthquakes (M $\sim$ 1), for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
-In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations~\citep{OBA_2002,ROG_2003}.
+Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
+In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\citep{OBA_2002,ROG_2003}.
 Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS).
 However, this is not always the case.
-For instance, in northern New Zealand, tremor are more challenging to detect, and seem to be located downdip of the slow slip on the plate boundary.
+For instance, in northern New Zealand, tremor are more challenging to detect, and seem to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}.
+In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal correlation between tremor and slow slip occurrence ~\citep{WEC_2016}.
 \\
 
-In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data.
+In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly observed in the GNSS data.
 For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor.
@@
 
-~\citet{FRA_2016} divided GPS time series observations into two groups: the first group contains days when slow seismicity (tectonic tremor and LFEs) is detected, the second group contains days when the numbers of tremor or LFEs is lower than a threshold.
-He then stacked separately the two groups of observations and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
+~\citet{FRA_2016} divided GPS time series observations from Guerrero and Cascadia into two groups: the first group contains days with abundant tremor and LFEs, the second group contains days when the number of tremor or LFEs is lower than a threshold.
+He then stacked separately the two groups of daily observations and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
 He also observed a cumulative displacement in the southern direction (for Guerrero) and the western direction (for Cascadia) corresponding to the release period when tremor and LFEs are observed.
-He was thus able to observe a reverse displacement corresponding to smaller slow slip events not directly observable in the GPS data.
+He was thus able to observe a reverse displacement corresponding to smaller slow slip events not directly observable in the GPS data for individual events.
 \\
@@
 
-Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next.
-Wavelet methods have been widely used for geophysical applications ~\citep{KUM_1997}.
+Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next.
+Wavelet methods have been widely used for geophysical applications (e.g.
+~\citep{KUM_1997}).
 However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events ~\citep{SZE_2008,OHT_2010,WEI_2012,ALB_2019}.
@@
 
-~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005.
+~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005 using wavelets.
 They modeled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, Heaviside step functions corresponding to earthquakes and hardware upgrades, and a residual signal.
-They then applied a Gaussian wavelet transform to the residual time series to get the exact timing of the slow slip at each GPS station.
+They then applied a Gaussian wavelet transform to the residual time series to get the exact timing of slow slip at each GPS station.
 The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains.

### 2021-10-26 paper/paper_bssa.tex
Included comments from DS

@@
 
-In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly observed in the GNSS data.
+In Cascadia, there are robust signals in both GNSS and tremor.
+This is also the case in Nankai, where tiltmeters are used instead of GNSS.
+It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data.
 For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor.
@@
 They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8.
-They also observed many smaller bursts of tremor of duration 1 to 50 hours in between the big ETS events, without any detectable signal in the GPS data.
-However, based on the relationship between slow slip moment and number of hours of tremor for bigger events, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with the tremor bursts.
-\\ %This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}).
+Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data.
 \\
 
-~\citet{FRA_2016} divided GPS time series observations from Guerrero and Cascadia into two groups: the first group contains days with abundant tremor and LFEs, the second group contains days when the number of tremor or LFEs is lower than a threshold.
-He then stacked separately the two groups of daily observations and observed a cumulative displacement in the northern direction (for Guerrero) and the eastern direction (for Cascadia) corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
-He also observed a cumulative displacement in the southern direction (for Guerrero) and the western direction (for Cascadia) corresponding to the release period when tremor and LFEs are observed.
+~\citet{FRA_2016} divided GPS time series observations from Cascadia and Guerrero, Mexico, into two groups: the first group contains days with abundant tremor and LFEs, the second group contains days when the number of tremor or LFEs is lower than a threshold.
+He then stacked separately the two groups of daily observations and observed a cumulative displacement in the direction corresponding to the loading period when few tremor or LFEs are observed and the surface deformation corresponds to the secular plate motion.
+He also observed a cumulative displacement in the opposite direction corresponding to the release period when tremor and LFEs are observed.
 He was thus able to observe a reverse displacement corresponding to smaller slow slip events not directly observable in the GPS data for individual events.
@@
 
-However, in other subduction zones such as New Zealand, there is no clear relationship between tremor and slow slip occurrence and these methods cannot be applied to detect smaller slow slip events that produce a GNSS signal with an amplitude too small compared to the noise.
+However, these methods cannot be applied to detect slow slip events in places where tremor and slow slip occurrence are not well spatially and temporary correlated, tremor is not abundant, or the seismic network is not robust enough.
 We thus need other methods to be able to better detect and quantify slow slip.
@@
 
-Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine vertical displacements, uplift rates between ETS events, and net uplift rates between 1996 and 2011.
+Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with ETS events between 1996 and 2011.
 Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events.
-The noise component is assumed to be coherent between all tidal gauges, while the uplift signal is different provided that the gauges are far enough from each other.
-By stacking the tidal records, the uplift signals cancel each other while the noise signal is amplified.
+The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other.
+By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified.
 By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal.
@@
 
-In our study, we use a similar approach with a different reasoning.
+In our study, we use a similar approach to previous studies with a different reasoning.
 We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered.
 We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking.
-Finally, we suppose that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
+Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
 We use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia.
 Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
-We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may be currently undetected with standard methods.
+We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods.
 \\
@@
 However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio.
-The wavelet method works best with data with zero mean, and no sharp discontinuities, so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}.
+The wavelet method works best with data with zero mean, and no sharp discontinuities; so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}.
 For the seismic data, we used the tremor catalog from \textcolor{red}{reference for the catalog.
@@
 The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$.
-If we denote $J$ the level of the wavelet decomposition, and we have $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
+If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
 , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$.
-Each wavelet vector $W_j$ is associated with changes on scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$.
-The scaling vector $V_J$ is associated with averages in scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$.
-We can also define for $j = 1 , \cdots , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to scale $\tau_j = dt 2^{j - 1}$.
+Each wavelet vector $W_j$ is associated with changes on time scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$.
+The scaling vector $V_J$ is associated with averages in time scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$.
+Wavelet vectors can be further decomposed into details and smooths, which are more easily interpretable.
+We define for $j = 1 , \cdots , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to time scale $\tau_j = dt 2^{j - 1}$.
 Similarly, we can define for $j = 1 , \cdots , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher.
+The basic idea is to reapply to $W_j$ the wavelet filter that was used to construct $W_j$ from the initial time series $X$.
 Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:
@@
 
-The DWT present several disadvantages.
+The DWT presents several disadvantages.
 First, the length of the time series must be a multiple of $2^J$ where $J$ is the level of the DWT decomposition.
@@
 Thus, the values of the wavelet coefficients, details and smooths are strongly dependent on the time when we start experimentally gathering the data.
-Finally, when we filter the time series to obtain the details and smooths, we introduce a phase shift, which makes difficult to line up meaningfully the features of the MRA with the original time series.
+Finally, when we filter the time series to obtain the details $D_j$ and smooths $S_j$, we introduce a phase shift, which makes it difficult to line up meaningfully the features of the MRA with the original time series.
 \\
 
-This is why we use instead the Maximal Overlap Discrete Wavelet Transform (MODWT).
+To overcome the disadvantages described above, we use instead the Maximal Overlap Discrete Wavelet Transform (MODWT).
 The MODWT transforms the time series $X_t \left( t = 0, ...
@@
 
-To illustrate the wavelet transform method, we first apply the MODWT to synthetics data.
+To illustrate the wavelet transform method, we first apply the MODWT to synthetic data.
 As slow slip events occur in Cascadia on a regular basis, every twelve to eighteen months, we create a synthetic signal of period $T = 500$ days.
-To reproduce the ground displacement observed on the longitudinal component of GPS stations in Cascadia, we divide each period into two parts: In the first part of duration $T - N$, the displacement is linearly increasing and corresponds to the secular plate motion in the eastern direction; in the second part of duration $N$, the displacement is linearly decreasing and corresponds to a slow slip event on a reverse fault at depth triggering a ground displacement in the western direction.
-To see the effect of the magnitude of the slow slip event, we use different values for $N = 5, 10, 20, 40$ days.
-Figure 1 shows the synthetics, the details of the wavelet decomposition for levels 1 to 10, and the smooth for the four durations of a slow slip event.
+To reproduce the ground displacement observed on the longitudinal component of GPS stations in Cascadia, we divide each period into two parts: In the first part of duration $T - N$, the displacement is linearly increasing and corresponds to the inter seismic plate motion in the eastern direction; in the second part of duration $N$, the displacement is linearly decreasing and corresponds to a slow slip event on a reverse fault at depth triggering a ground displacement in the western direction.
+To see the effect of the duration of the slow slip event, we use different values for $N = 5, 10, 20, 40$ days.
+The amplitude of the set is normalized to 1.
+Figure 1 shows the synthetics, the details $D_j$  of the wavelet decomposition for levels 1 to 10, and the smooth $S_{10}$ for the four durations of a slow slip event.
 \\
@@
 Thus, the scale levels at which an event is being seen in the wavelet details give us an indication about the duration (and the magnitude) of the slow slip event.
-We expect the big slow slip events of magnitude 6-7 that last several weeks to start being visible at the level 5 of the wavelet decomposition, but to not be noticeable at lower time scales.
+The big slow slip events of magnitude 6-7 typically trigger a signal that lasts about one week at an individual GPS station, and the whole event lasts several weeks.
+We expect them to start being visible at the level 5 of the wavelet decomposition, but to not be noticeable at lower time scales.
 \\
@@
 The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings.
-To deal with the gaps in the GNSS recordings, we simply replace the missing values by the sum of a straight line and a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series.
-The straight line starts at the mean of the five days before the gap and ends at the mean of the five days after the gap.
-We verify how the wavelet details may be affected by looking at a GPS time series without missing values and comparing the wavelet details with and without removing some data points.
-Station PGC5 has recorded during 1390 days between 2009 and 2013, without any missing values.
+To deal with the gaps in the GNSS recordings, we simply replace the missing values by interpolation.
+The value for the first day for which data are missing is equal to the mean of the five days before the gap.
+The value for the last day for which data are missing is equal to the mean of the five days after the gap.
+The remaining missing values are computed by doing a linear interpolation of the first and the last values and adding a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series.
+The straight line starts at and ends at .
+We verify how the wavelet details may be affected by looking at a GPS time series without missing values and compared the wavelet details with and without removing some data points.
+Station PGC5 recorded continuous 1390 days between 2009 and 2013 without any missing values.
 We first computed the wavelet details without missing values.
-Then, we removed ten neighboring missing values, replaced them by the sum of the straight line and the Gaussian noise component, and computed the wavelet details with the replaced values.
+Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values.
 Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values.
 We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition.
-However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels the details, and are barely visible for the levels we are mostly interested in (levels 6 and above).
+However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels of details, and are barely visible for the levels that are most relevant (levels 6 and above).
 We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events.
@@
 Figure 3 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth.
-In the data, we can see a sharp drop in displacement whenever there is a slow slip event.
-For levels 5 to 8, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data.
-We thus verify that the wavelet method can detect slow slip events.
+In the data, we can see a sharp drop in displacement whenever there is a documented slow slip event.
+For levels 5 to 8, which correspond to time scales 16, 32, 64 and 128 days, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data.
+We thus verify that the wavelet method can detect steps in the time series associated with slow slip events.
 \\
 
-To increase the signal-to-noise ratio and be able to better detect slow slip events, we stack the signal over several GPS stations.
+To increase the signal-to-noise ratio and better detect slow slip events, we stack the signal from several neighboring GPS stations.
 We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal.
-We choose 16 points located on the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4).
+We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4).
 Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations.
@@
 
-To compare slow slip events detected with GPS data and slow slip events detected with seismic data, we took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3.
+To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for ETS events.
+We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3.
 Then we computed the cumulative number of tremor within this circle.
@@
 Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5).
-Contrary to what happens for the GPS data, we see a sharp increase in the data whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
+Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
 
@@
 We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-The result is shown in the top panel of Figure 6, where each line represents one of the locations.
+The result is shown in the top panel of Figure 6, where each line represents one of the locations along strike.
 To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold.
 To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6.
-We used the opposite of the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
+We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
 In the tremor catalog from \textcolor{red}{reference?}, there are 17 tremor events with more than 150 hours of tremor recorded.
@@
 The time of the event is the start date plus half the duration of the event.
+\\
+
 Although the latitudinal extension of the events is not always the same for the GPS data and for the tremor data, we identify the same 13 events in both 8th wavelet decompositions for the 8th level: January 2007, May 2008, May 2009, August 2010, August 2011, September 2012, September 2013, August-November 2014, January 2016, March 2017, June 2018, March-November 2019, and October 2020-January 2021.
@@
 
-Figures 7 and 8 show the same comparison between the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively.
+A similar comparison is shown for the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively (Figures 7 and 8).
 The events are harder to see in the 7th level than in the 8th level, both for the GPS data and the tremor count data.
@@
 There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections.
-To summarize, all the 13 events present on the 8th level detail of the wavelet decomposition are true detections, 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
+To summarize, we assume that true detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series.
+Then, all the 13 events present on the 8th level detail of the wavelet decomposition are true detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
 \\
@@
 We do this on both the wavelet details of the GPS data and of the tremor data.
-Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a true detection (true positive TP).
-If the GPS and the tremor time series have opposite signs, or if the absolute value of the GPS time series is 1 but the value of the tremor time series is 0, we have a false detection (false positive FP).
-If both time series take the value 0, we do not have detection (true negative TN).
-If the GPS time series take the value 0, but the absolute value of the tremor time series is 1, we miss a detection (false negative).
+Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a true detection (true positive, TP).
+If the GPS and the tremor time series have opposite signs, or if the absolute value of the GPS time series is 1 but the value of the tremor time series is 0, we have a false detection (false positive, FP).
+If both time series take the value 0, we do not have detection (true negative, TN).
+If the GPS time series take the value 0, but the absolute value of the tremor time series is 1, we miss a detection (false negative, FN).
 We then define the sensitivity (true positive rate) and the specificity (equal to 1 minus the false positive rate) as:
@@
 We can then evaluate the quality of the detections obtained with our method by plotting a receiver operating characteristic curve (ROC curve).
-The ROC curve is widely use for binary classification problems, and is plotted by varying the values of the threshold (here the thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
+The ROC curve is widely use for binary classification problems in statistics and machine learning.
+We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
 If the classification was made randomly, all the points would fall on the first diagonal.
 If the classifier was perfect, the corresponding point would fall on the top left corner of the graph with true positive rate equal to 1 and false positive rate equal to 0.
-The bigger the area under the curve, the better the classifier is.
+The bigger the area under the curve, the better the classification method is.
 \\
@@
 We apply this transform to both the GPS and the tremor time series with varying thresholds.
-We then plot the ROC curve on Figure 9.
+We then plot the ROC curve on Figure 9, each dot representing a different threshold.
 The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10.
-We can see that there is a trade-off between sensitivity and specificity.
+We can see that there is a trade-off between sensitivity and specificity as we vary the threshold.
 If we decrease the false positive rate, we also decrease the number of true events detected.
 If we increase the number of true events detected, we also increase the false positive rate.
-In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75, and the sensitivity is the highest possible.
+In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
 \\
@@
 Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event.
+Figure 11 shows the 5th level detail wavelet decomposition of the GPS data.
+Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
 \\
 
-Figure 11 shows the 5th level detail wavelet decomposition of the GPS data.
-Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
-
-All 14 events that are sub-event of a bigger event are visible at level 5.
-However, this may be due because the bigger event are clearly seen at levels 6 to 8, and also at smaller time scales.
+All 14 events that are sub-events of a bigger event are visible at level 5.
+However, this may be because the bigger events are clearly seen at levels 6 to 8, and also at smaller time scales.
 The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition.
-However, some other events that are not in ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
-Therefore, it is difficult to make the difference between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
+However, some other events that are not in the catalog of ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
+Therefore, it is difficult to differentiate between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
 
@@
 
-In this paper, we have used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
-We used detrended GNSS data, applied the MODWT transform and stack the wavelet details over several nearby GNSS stations.
+In this paper, we develop and test a new approach for detecting transient events in GPS time series, such as slow slip events.
+We used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+We used detrended GNSS data, applied the MODWT transform, and stacked the wavelet details over several nearby GNSS stations.
 As an independent check on the timing of slow slip events, we also computed the cumulative number of tremor in the vicinity of the GNSS stations, detrended this signal, and applied the MODWT transform.
 In both time series, we could then see simultaneous waveforms whose timing corresponds to the timing of slow slip events.
-We assumed that there is a slow slip event whenever there is a peak in the wavelet signal.
+We assumed that there is a slow slip event whenever the wavelet signal gets above a threshold.
 We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
@@
 ~\citep{MIC_2019}).
-We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to make the difference between true detections and false detections.
+We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between true detections and false detections.
 
@@
 The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/slowslip}.
-Figure 4 was done using GMT ~\citep{WES_1991}.
+Figure 4 was created using GMT ~\citep{WES_1991}.
 
@@
 \begin{table}[hbt!]
-\caption{Big Episodic Tremor and Slip events.
+\caption{Episodic Tremor and Slip events with M > 6 identified by MODWT in both the GPS and the tremor data.
 The duration and the number of tremor are from the tremor catalog of \textcolor{red}{reference?}.
@@
 
-\item Figure1.
-Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).
+\item Figure 1.
+Demonstration of a wavelet decomposition for a synthetic dataset.
+A synthetic time series is created (top row) with steps of period 500 days, and transient durations of 2 days (left), 5 days, 10 days, and 20 days (right).
+The resulting details and smooths are shown in increasing level.
+The amplitude of the synthetic time series is normalized to 1, and the details and smooths show the relative amplitude.
 
-\item Figure2.
-Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).
+\item Figure 2.
+Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
+The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).
 
-\item Figure3. Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.
+\item Figure 3. Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.
 
@@
 
-\item Figure 5. Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.
+\item Figure 5. Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).
 
 \item Figure 6.
-Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 8th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 (for the GPS) or 0.003 (for the tremor).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 7.
-Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+\item Figure 7. Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).
 
-\item Figure 8.
-Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+\item Figure 8. Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).
 
-\item Figure 9.
-Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the sum of the 6th, 7th and 8th level details of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+\item Figure 9. Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
 
@@
 ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
 The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
@@
 \item Figure 11.
-Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
+Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+The red lines represent the timings of the ETS events from Table 1.
+The blue lines represent the timings of the magnitude 5 events from the catalog of ~\citet{MIC_2019}.
 
@@
 \begin{figure}
-\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
-\caption{Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).}
+\noindent\includegraphics[width=10cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
+\caption{Demonstration of a wavelet decomposition for a synthetic dataset.
+A synthetic time series is created (top row) with steps of period 500 days, and transient durations of 2 days (left), 5 days, 10 days, and 20 days (right).
+The resulting details and smooths are shown in increasing level.
+The amplitude of the synthetic time series is normalized to 1, and the details and smooths show the relative amplitude.}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).}
+\caption{Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
+The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
-\caption{Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.}
+\caption{Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
-\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.}
+\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.eps}
-\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 8th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 (for the GPS) or 0.003 (for the tremor).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
-\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\caption{Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.pdf}
-\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\caption{Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
-\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-Bottom: Opposite of the sum of the 6th, 7th and 8th level details of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
+\caption{Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).}
 \label{pngfiguresample}
@@
 \caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
 The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.pdf}
-\caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.}
+\caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+The red lines represent the timings of the ETS events from Table 1.
+The blue lines represent the timings of the magnitude 5 events from the catalog of ~\citet{MIC_2019}.}
 \label{pngfiguresample}

### 2021-10-26 src/missing_values.py

@@ -81,27 +81,25 @@ for n in range(1, nmax, nstep):
         (D, S) = get_DS(disp_interp, W, name, J)
 
         # Plot data
-        ax = plt.subplot2grid((J + 2, len(gaps)), (J + 1, count))
+        ax = plt.subplot2grid((J + 2, len(gaps)), (0, count))
         plt.plot(time, disp_interp, 'r')
         plt.plot(time, disp, 'k', label='Data')
         plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))
-        plt.legend(loc=1)
-        plt.xlabel('Time (years)', fontsize=14)
+        plt.legend(loc=1)        
         # Plot details at each level
         for j in range(0, J):
-            ax = plt.subplot2grid((J + 2, len(gaps)), (J - j, count))
+            ax = plt.subplot2grid((J + 2, len(gaps)), (j + 1, count))
             plt.plot(time, D[j], 'r')
             plt.plot(time, D0[j], 'k', label='D' + str(j + 1))
             plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))
             plt.legend(loc=1)
-            ax.axes.xaxis.set_ticks([])
         # Plot scaling coefficients for the last level
-        ax = plt.subplot2grid((J + 2, len(gaps)), (0, count))
+        ax = plt.subplot2grid((J + 2, len(gaps)), (J + 1, count))
         plt.plot(time, S[J], 'r')
         plt.plot(time, S0[J], 'k', label='S' + str(J))
         plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))
         plt.legend(loc=1)
-        ax.axes.xaxis.set_ticks([])
+        plt.xlabel('Time (years)', fontsize=14)
 
     plt.tight_layout()
     plt.savefig('missing_values/' + filling + '/DS_' + str(n) + '.eps', format='eps')

### 2021-10-26 src/vespagram.py

@@ -112,7 +112,7 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         (D, S) = get_DS(disp, W, wavelet, J)
 
         # Save wavelets into file
-        filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
+        filename = 'MODWT_GPS_longer/' + dataset + '_' + station + '_' + direction + '.pkl'
         pickle.dump([time, disp, W, V, D, S], open(filename, 'wb'))
 
         # Start figure
@@ -125,29 +125,29 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
         minD = min([np.min(Dj) for Dj in D])
 
         # Plot data
-        plt.subplot2grid((J + 2, 1), (J + 1, 0))
+        plt.subplot2grid((J + 2, 1), (0, 0))
         plt.plot(time, disp, 'k', label='Data')
-        plt.xlabel('Time (years)', fontsize=24)
-        plt.xlim([2009.25, 2021.25])
+        plt.xlim([2006.0, 2021.5])
         plt.ylim([np.min(disp), np.max(disp)])
         plt.legend(loc=3, fontsize=20)
         # Plot details
         for j in range(0, J):
-            plt.subplot2grid((J + 2, 1), (J - j, 0))
+            plt.subplot2grid((J + 2, 1), (j + 1, 0))
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
-            plt.xlim([2009.25, 2021.25])
+            plt.xlim([2006.0, 2021.5])
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
         # Plot smooth
-        plt.subplot2grid((J + 2, 1), (0, 0))
+        plt.subplot2grid((J + 2, 1), (J + 1, 0))
         plt.plot(time, S[J], 'k', label='S' + str(J))
-        plt.xlim([2009.25, 2021.25])
+        plt.xlim([2006.0, 2021.5])
         plt.ylim([np.min(disp), np.max(disp)])
         plt.legend(loc=3, fontsize=20)
+        plt.xlabel('Time (years)', fontsize=24)
         
         # Save figure
         plt.tight_layout()
-        plt.savefig('MODWT_GPS/' + dataset + '_' + station + '_' + \
+        plt.savefig('MODWT_GPS_longer/' + dataset + '_' + station + '_' + \
             direction + '.eps', format='eps')
         plt.close(1)
         

### 2021-10-30 paper/paper_bssa.tex
last paragraph

@@
 \item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
-\item There is a good correlation between detections using GNSS data and independent detections using seismic data.
+\item There is a good correlation between detections of slow slip using GNSS data and using tremor data.
 \item The method could be applied in regions where no tremor are detected in conjunction with slow slip events.
@@
 We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal.
-We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data for northern Cascadia.
+We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia.
 The wavelet-based detection method detects well events of magnitude higher than 6 as determined by independent event catalogs (e.g.
@@
 Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
-We use wavelet methods to analyze GPS and seismic recordings of slow slip events in Cascadia.
-Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
+We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia.
+Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
 We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods.
@@
 The wavelet method works best with data with zero mean, and no sharp discontinuities; so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}.
-For the seismic data, we used the tremor catalog from \textcolor{red}{reference for the catalog.
-The following is to be modified.
-the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}.
-Tremor were detected and located using waveform envelope correlation and clustering and a centroid location is available for every given five-minute time window when tremor was detected.
-As the catalog starts in August 2009, we only looked at GPS data recorded in 2009 or later.} \\
+For the tremor data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}.
+\\
 
@@
 We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
-In the tremor catalog from \textcolor{red}{reference?}, there are 17 tremor events with more than 150 hours of tremor recorded.
+In the tremor catalog from the PNSN, there are 17 tremor events with more than 150 hours of tremor recorded.
 The events are summarized in Table 1.
@@
 It may correspond to the northern extent of a big ETS event occurring in Fall 2009 south of the study area (event 19 in the ~\citet{MIC_2019} catalog).
-There are three small signals in the GPS data in Spring 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and are probably false detections.
+There are three small signals in the GPS data in Winter 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and may be false detections.
 To summarize, we assume that true detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series.
@@
 Therefore, it is difficult to differentiate between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
+\\
+
+In Figure 9, we see four smaller events that are not in the catalog of ~\citet{MIC_2019}: at about 2007.5, there is a negative peak followed by a positive peak (that is an event in the opposite direction of what would be expected from slow slip), at about 2010.2, 2012.2 and 2020.2, there are positive peaks followed by negative peaks for all the sixteen locations studied in this paper.
+Looking back at the original GPS data, there is a small increase in the displacement in the eastern direction that lasts about one or two months at about 2007.5.
+However, the direction of the displacement does not correspond to a slow slip event, and another cause should be found to explain this signal.
+There is a decrease in displacement that lasts several months at about 2010.2.
+This transient may correspond to a long duration slow slip event.
+There is a small decrease in displacement at about 2012.2.
+Its amplitude is small but the duration and direction correspond to a slow slip event, so this transient could be a very small slow slip event.
+Finally, there is also a small decrease in displacement at about 2020.2 that is difficult to interpret.
 
@@
 In this paper, we develop and test a new approach for detecting transient events in GPS time series, such as slow slip events.
-We used wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia.
+We used wavelet methods to analyze GNSS time series and tremor recordings of slow slip events in Cascadia.
 We used detrended GNSS data, applied the MODWT transform, and stacked the wavelet details over several nearby GNSS stations.
@@
 We assumed that there is a slow slip event whenever the wavelet signal gets above a threshold.
-We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data.
+We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
 The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
@@
 The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA} \url{http://www.panga.cwu.edu/}.
-\textcolor{red}{Reference for tremor catalog}.
 The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/slowslip}.
@@
 
-This work was funded by the grant from the National Science Foundation XXX.
+This work was funded by the grant from the National Science Foundation EAR-1358512.
 A.D.
-would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington.
+would like to thank Professor Donald Percival for introducing her to wavelet methods during his excellent class on Wavelets: Data Analysis, Algorithms and Theory taught at University of Washington.
 
@@
 \caption{Episodic Tremor and Slip events with M > 6 identified by MODWT in both the GPS and the tremor data.
-The duration and the number of tremor are from the tremor catalog of \textcolor{red}{reference?}.
+The duration and the number of tremor are from the tremor catalog of the PNSN.
 The event number and the magnitude are from the slow slip catalog of ~\citet{MIC_2019}.}

### 2021-12-01 paper/paper_bssa.tex
Make figures for movie

@@
 \usepackage{natbib}
+\usepackage{setspace}
 \usepackage{url}
@@
 \author[1]{Ariane Ducellier}
-\author[1]{Kenneth C. Creager}
-\author[1]{David A. Schmidt}
-\affil[1]{University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
+\author[2]{Kenneth C. Creager}
+\author[2]{David A. Schmidt}
+\affil[1]{Corresponding author. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
+\affil[2]{University of Washington, Department of Earth and Space Sciences}
 \date{}
@@
 \newpage
+
+\doublespacing
 

### 2021-12-01 src/compare_tremor_GPS.py

@@ -417,6 +417,8 @@ def plot_correlations():
         plt.savefig('events_' + str(level) + '.pdf', format='pdf')
         plt.close(1)        
 
+
+
 def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, threshold, events, small_events):
     """
     """
@@ -912,6 +914,10 @@ if __name__ == '__main__':
     dataset = 'cleaned'
     direction = 'lon'
     radius_GPS = 50
+    lonmin = -125.4
+    lonmax = -121.4
+    latmin = 46.3
+    latmax = 49.6
     events = [2007.0630, 2008.3566, 2009.3452, 2010.6342, 2011.6041, 2012.7172, \
         2013.7137, 2014.6479, 2014.8904, 2016.1082, 2017.2342, 2018.4932, \
         2019.2329, 2019.8767, 2020.7923, 2020.8552, 2021.0904]
@@ -942,15 +948,16 @@ if __name__ == '__main__':
 
     thresh_GPS = np.arange(0.1, 1.6, 0.1)
     thresh_tremor = np.arange(0.001, 0.011, 0.001)
-    chosen_GPS = 0.5
+    chosen_GPS = 0.8
     chosen_tremor = 0.01
-    J = [6, 7]
+    J = [6, 7, 8]
 
-    plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, chosen_GPS, events, small_events)
-    plot_tremor(lats, J, chosen_tremor, events, small_events)
+#    plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, chosen_GPS, events, small_events)
     
-    (TP, TN, FP, FN) = find_false_detections(station_file, lats, lons, dataset, direction, \
-        radius_GPS, J, chosen_GPS, chosen_tremor, events, True)
+#    plot_tremor(lats, J, chosen_tremor, events, small_events)
     
-    (sensitivity, specificity) = plot_ROC_curve(station_file, lats, lons, dataset, direction, \
-        radius_GPS, J, thresh_GPS, thresh_tremor, events, chosen_GPS, chosen_tremor)
+#    (TP, TN, FP, FN) = find_false_detections(station_file, lats, lons, dataset, direction, \
+#        radius_GPS, J, chosen_GPS, chosen_tremor, events, True)
+    
+#    (sensitivity, specificity) = plot_ROC_curve(station_file, lats, lons, dataset, direction, \
[…]

### 2021-12-03 src/make_moviefigure.py
Start movie

@@ -283,7 +283,7 @@ if __name__ == '__main__':
     latmin = 46.3
     latmax = 49.6
 
-    for image in range(0, 20):
+    for image in range(17, 21):
 
         tmin_GPS = 2010.25 + image * 0.02 - 1.0
         tmax_GPS = 2010.25 + image * 0.02 + 1.0

### 2021-12-03 src/synthetics.py

@@ -53,6 +53,8 @@ for timestep in timesteps:
         plt.legend(loc=3, fontsize=20)
         if i != 0:
             ax.axes.yaxis.set_ticks([])
+        title = 'Duration of event = ' + str(duration) + ' days'
+        plt.title(title, fontsize=20)
         # Plot details
         for j in range(0, J):
             ax = plt.subplot2grid((J + 2, len(durations)), (j + 1, i))
@@ -69,8 +71,6 @@ for timestep in timesteps:
         plt.legend(loc=3, fontsize=20)
         if i != 0:
             ax.axes.yaxis.set_ticks([])
-        title = 'Duration of event = ' + str(duration) + ' days'
-        plt.title(title, fontsize=20)
 
     # Save figure
     plt.tight_layout()

### 2021-12-03 src/synthetics_poster.py

@@ -1,0 +1,64 @@
+"""
+Script to make synthetic slow slip events and run MODWT on them
+"""
+
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
+import numpy as np
+
+import MODWT
+
+# Time between events
+timestep = 500
+# Durations of slow slip events
+durations = [2, 5, 10, 20]
+# MODWT wavelet filter
+name = 'LA8'
+# Duration of recording
+N = 2000
+# MODWT level
+J = 10
+
+# Create time vector
+time = np.arange(0, N)
+
+# Set random seed
+np.random.seed(0)
+
+# Start figure
+params = {'xtick.labelsize':20,
+          'ytick.labelsize':20}
+pylab.rcParams.update(params)   
+fig = plt.figure(1, figsize=(5 * len(durations), 3 * (2)))
+
+# Loop on duration
+for i, duration in enumerate(durations):
+    # Create displacements vectors
+    part1 = np.arange(0, timestep - duration) / (timestep - duration)
+    part2 = (timestep - np.arange(timestep - duration, timestep)) / duration
+    signal = np.concatenate([part1, part2])
+    disp = np.tile(signal, int(N / timestep))
+    # Compute MODWT
+    (W, V) = MODWT.pyramid(disp, name, J)
+    (D, S) = MODWT.get_DS(disp, W, name, J)
+    # Plot data
+    ax = plt.subplot2grid((2, len(durations)), (0, i))
+    plt.plot(time, disp, 'k', label='Data')
+    plt.ylim(-2.0, 2.0)
+    plt.legend(loc=3, fontsize=20, frameon=False)
+    if i != 0:
+        ax.axes.yaxis.set_ticks([])
+    title = 'Duration of event = ' + str(duration) + ' days'
+    plt.title(title, fontsize=20)
+    # Plot sum of details 6, 7 and 8
+    ax = plt.subplot2grid((2, len(durations)), (1, i))
+    plt.plot(time, D[5] + D[6] + D[7], 'k', label='D6 + D7 + D8')
+    plt.ylim(-2.0, 2.0)
+    plt.legend(loc=3, fontsize=20, frameon=False)
+    if i != 0:
+        ax.axes.yaxis.set_ticks([])
+    plt.xlabel('Time (days)', fontsize=20)
+
+plt.tight_layout()
+plt.savefig('synthetics/D678.eps', format='eps')
+plt.close(1)

### 2022-03-10 paper/paper_bssa.tex
Add citations to paper

@@
 \item There is a good correlation between detections of slow slip using GNSS data and using tremor data.
-\item The method could be applied in regions where no tremor are detected in conjunction with slow slip events.
+\item The method could be applied in regions where no tremor is detected in conjunction with slow slip events.
 \end{itemize}
@@
 
-Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks.
+Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}.
 As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate.
@@
 A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion.
-Slow slip events have been observed in many places, such as Cascadia, Nankai (southwest Japan), Alaska, Costa Rica, Mexico, and New Zealand ~\citep{BER_2011,AUD_2016}.
+Slow slip events have been observed in many places ~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}.
 \\
 
-In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places.
+In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places ~\citep{HAL_2018}.
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
@@
 However, this is not always the case.
-For instance, in northern New Zealand, tremor are more challenging to detect, and seem to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}.
+For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}.
 In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal correlation between tremor and slow slip occurrence ~\citep{WEC_2016}.
@@
 
-In Cascadia, there are robust signals in both GNSS and tremor.
-This is also the case in Nankai, where tiltmeters are used instead of GNSS.
+In Cascadia, there are robust signals in both GNSS and tremor ~\citep{HAW_2013}.
+This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS.
 It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data.
@@
 
-However, these methods cannot be applied to detect slow slip events in places where tremor and slow slip occurrence are not well spatially and temporary correlated, tremor is not abundant, or the seismic network is not robust enough.
+However, these methods cannot be applied to detect slow slip events in places where tremor and slow slip occurrence are not well spatially and temporally correlated, tremor is not abundant, or the seismic network is not robust enough.
 We thus need other methods to be able to better detect and quantify slow slip.
@@
 
-Wavelets methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next.
+Wavelet methods such as the Discrete Wavelet Transform (DWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next.
 Wavelet methods have been widely used for geophysical applications (e.g.
-~\citep{KUM_1997}).
+~\citet{KUM_1997}).
 However, few studies have used wavelet methods to analyze recordings of slow slip, and their scope was limited to the detection of the bigger (magnitude 6-7) short-term (a few weeks) events ~\citep{SZE_2008,OHT_2010,WEI_2012,ALB_2019}.
@@
 
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.eps}
 \caption{Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).}
@@
 \begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.pdf}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.eps}
 \caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.

### 2022-04-09 paper/answer_reviewers.tex
Work on revisions

\documentclass[letterpaper, 12pt]{article}

\usepackage{caption}
\usepackage{float}
\usepackage[T1]{fontenc}
\usepackage[lmargin=1 in, rmargin=1 in, tmargin=1 in, bmargin=1 in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{times}
\usepackage{xcolor}

\begin{document}

\textbf{Reviewer 1}

\bigskip

\textit{In "Detection of slow slip events using wavelet analysis of GNSS recordings", the authors propose a technique to detect slow slip events using the Maximal Overlap Discrete Wavelet Transform (MODWT). While previous authors have used or advocated the use of wavelets for the detection of slow slip in geodetic time series, this manuscript provides a rubric for detection and provides a connection to slow slip duration. Applying this rubric to slow slip events in northern Cascadia, the authors provide a method for using the MODWT on both geodetic and tremor duration time series to demonstrate the general applicability of the technique. The authors advocate using this technique in situations where the connection between slow slip and tectonic tremor is less obvious.} \\

\textit{This manuscript provide a useful and much needed exploration and explanation of the relationship between wavelet levels and offset duration. In particular, advocating for the use of the MODWT over the standard discrete wavelet transform provides a nice way to avoid boundary effects at the edge of each wavelet octave and provides better temporal location for identification of wavelet peaks. The applicability to the case of Cascadia presented in the manuscript is clear and concise as well. My one big wish for this manuscript is a demonstration of its power in the situations described in the Introduction (lines 69-74), in particular, for those subduction zones where the connection between tremor and slow slip is not straightforward. To that end, I think the biggest test missing from this manuscript is a test in one of the problematic subduction zones, such as New Zealand or Alaska. Much of the demonstration of the power of this technique lies in selecting the threshold. For instance, how applicable are the thresholds determined from northern Cascadia data to other regions (say NZ or AK)? If those subduction zones were examined, how would one go about selecting a threshold? In short, is it possible to select a threshold in the absence of a catalog or does threshold selection depend on slow slip events already being know from other methods?}

\bigskip

We added an analysis of New Zealand GPS data.

\bigskip

\textit{While Figure 10 does show that nearly all the thresholds examined demonstrate a higher true positive rate than false positive, not all threshold value have equal identification power. Additionally, the actual value of the threshold is missing from the diagram (it would be a third dimension in the diagram). It is therefore difficult to ascertain whether the threshold levels explored provide complete coverage of reasonable values.}

\bigskip

We modified Figure 10 to better show the results corresponding to the different values of the threshold, by using a different color for each point depending on the values taken by the two thresholds.

\bigskip

\textit{Thus, while I feel that this manuscript highlights a potentially useful mathematical technique for identifying slow slip in geodetic time series, its broader applicability is not demonstrated. This manuscript would be improved through a demonstration of the proposed method, along with a comparison to an existing slow slip catalog, in a separate slow slip area.}

\bigskip

We added a comparison of slow slip events detected with this method and slow slip events detected by Todd and Schwartz (2016) in New Zealand.

\bigskip

\textit{In addition, I note the following items:} \\

\textit{(Line 172): I'm pretty sure the reference frame for the PANGA GPS time series is ITRF14. Orbits from JPL in the IGb08 reference frame end in early-mid 2018 and the time series for PGC5 shown in Figure 3 extend at least half-way through 2020. Therefore, the orbits used to process PGC5 can't be in IGb08.}

\bigskip

Thank you for pointing out this mistake. We corrected the sentence.

\bigskip

\textit{(Line 391) If the time series is categorized by the values -1, 0, and 1, wouldn't that be a trinary time series, not a binary time series?}

\bigskip

We replaced binary by trinary.

\bigskip

\textit{(Lines 458-460) These lines discuss a small displacement in the time series at 2007.5, but I don't see it. Should I be looking at Figure 9? If so, which latitude band should I look at and which time series are they composed of? That would be the place to look for causal offsets.}

\bigskip

\textit{(Figure 2) The caption text is not clear, in particular the phrase "for two locations of the missing values" is vague. The text (lines 290-295) does provide a better explanation of this figure however. Rewording the caption to provide a more self-contained explanation would improve the caption. Also, the values on the time axis for the right hand column are difficult to read as formatted.}

\bigskip

We modified the caption and the values on the time axis.

\bigskip

\textit{(Figure 5) The ordering of the data and the detail levels in this figure are opposite that of the Figures 1-3. For the sake of comparison, these should all be in the same order.}

\bigskip

We reordered the data and the detail levels.

\bigskip

\textit{(Figure 6) When producing the stacked time series, is there overlap/redundancy? Looking at the map in Figure 4 suggests that most triangles will be made with an overlapping number of stations. How will this affect the mapping of strange station behavior? In particular, could the displacement discussed on Lines 458-460 all be caused by just one misbehaving station that shows up in multiple bins?}

(Figure 7) Why do the obvious red-to-blue transitions in the tremor data seem to disappear from the detail level 7 figure, but are present in the detail level 8 (Figure 6) and detail level 6 (Figure 8) data? Can anything be interpreted from that pattern?

(Figure 9) When stacking levels 6, 7 and 8, won't the lower numbered levels dominate because of their (generally) larger amplitudes (as seen in Figure 3)? Is it valuable to normalize the levels before summing, or will this cause other issues?

(Figure 10) Is it possible to select a threshold in the absence of a catalog or does threshold selection depend on slow slip events already being know from other methods?

(Figure 11) What are the dates selected by this algorithm? How do they compare with the dates given in Michel et al., 2019? Most look spot on, but others, in particular the 2009 event look like they are quite distant in time.

There are some awkward sentences as well:

 (Line 190-191): "where n is some integer greater than or equal to" sounds better than "higher or equal".

(Line 199): The terminology "details" and "smooths" sounds awkward, is that the standard terminology in the wavelet literature?

(Line 287-288) The sentence "The straight line starts at and ends at." is incomplete.

\textbf{Reviewer 2}

The authors present a wavelet-based method to identify slow slip events within the GNSS position time series. Rather than looking at the geodetic record of slow slip as a sum of periodic basis functions, the wavelet method allows for the isolation of slow slip signals across different time scales at different times during the period of study. The approach is of interest to the community and I think that this manuscript will eventually be publishable.

That said, my major concern is that the method as presented will be very difficult to apply to another region of study, especially one without a systematically generated catalog of tectonic tremor or low-frequency earthquakes. Here the authors validate the detections (i.e. establish a detection threshold) within the geodetic data with detections made in the seismic data. Looking at figures 6-8, there are many wiggles within the time series, and it's not obvious to me how to interpret these signals without the well-established catalogs of A. Wech and the PNSN (tremor) and Michel et al. (2019) (slow slip); such detailed and long-duration catalogs are not available outside of Japan to validate the results of the presented method. Aggravating my concern is the premise that the authors themselves set up in the introduction of the manuscript, that a method is needed to detect slow slip when tremor is not available as an independent control; the presented method does not reach that
goal. I would also highlight that slow slip in northern Cascadia is some of the most well behaved slow slip across all subduction zones, with short-ish recurrence intervals and consistently clear geodetic signals captured by a relatively dense GNSS network, that will likely show up more reliably than elsewhere with this sort of approach.

The figures for the most part are clear, but there are some improvements to be made (see comments below). I also find that the introduction is lacking details and there are way too few articles cited for background reading for interested readers (cf. first paragraph with only two unique citations that discusses observations made across many subduction zones).

I think a revised version of this manuscript that has reframed the purpose of the approach will be an interesting contribution that is appropriate to be published in BSSA. I list other minor comments below:

general - There is no discussion of common mode signals that will surely stack constructively across GNSS stations. The removal of sinusoidal seasonal signals will remove some common mode signal, but there can still exist some non-tectonic signals that will stack constructively within the bandwidth of interest here. There is a reasonable chance that this will have little impact on the results, but it should at least be verified/discussed given the proposed processing.

general - I would avoid using the term ETS as it only adds confusion, as lines 69-74 highlight. I would suggest using "slow slip" to describe geodetic observations of aseismic slip and "tectonic tremor" for the accompanying seismic signals. The following paragraph further underlines this potential confusion: discussing the moment release of an "ETS event" is ambiguous. As described, it is not obvious to those unfamiliar with slow slip and tremor that the aseismic moment of slow slip dominates the seismic moment of tremor, because ETS conflates the geodetic observations with the seismic ones. I would simply use the term slow slip when referring to the dominant aseismic transient.

general - I'm not sure that the "true detection" terminology used here is quite appropriate. This implies that we have an absolute knowledge of what's happening on the plate interface, and that there can't be new detections made here without already be corroborated by other catalogs. Perhaps "robust detection" might be more appropriate? In any case, this underlines my above major comment that this method is difficult to interpret without other catalogs.

line 42 - "feature" is a strange word choice and does not tell the reader much at all

line 60 - "it is more abundant in some places" doesn't tell us much

line 64 - "for which" -> whose

line 67 - Obara (2002) did not discuss slow slip, only tectonic tremor. I would suggest something like Obara et al., GRL, 2004

line 126 - "levels" does not mean very much without context

lines 148 - "details" same comment as above; this general term doesn't mean anything without context

line 151-152 - why is there supposed to be a difference between the two tidal gauges and what does that mean? Not clear without the reader being forced to go and read the cited article.

line 156 - I assume longitudinal displacements is supposed to mean east-west motion?

line 287-288 - missing text here

figure 2 - I would say figure 2 could be moved to the appendix or supplementary material. It's not key to the message of the manuscript

figures 6-8 - I think this figure would be more easily interpretable as well as much effective if only one panel was used, with alternating geodetic and seismic data for a given location (red triangle from Figure 4)in two different colors (black and something else). The red/blue rectangles that indicate detections could be four similar colors for geodetic/seismic detections (blue/purple for negative peaks and red/orange for positive peaks for example). I'd also suggest that figures 7 and 8 could be made into supplementary figures as they show similar information to figure 6

figure 7 - I'm not sure I understand why the threshold selected here is less effective (there are less events identified) than for levels levels 6 and 8. What could explain this?

figure 10 - I think this figure would be a bit clearer if the 6/7/8 detail levels were in different colors

\end{document}

### 2022-04-09 paper/paper_bssa_annotated.tex
Work on revisions

--- 2022-03-10 paper/paper_bssa.tex
@@
 
-Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}.
+Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks \textbf{~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}}.
 As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate.
@@
 A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion.
-Slow slip events have been observed in many places ~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}.
+Slow slip events have been observed in many places \textbf{~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}}.
 \\
 
-In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places ~\citep{HAL_2018}.
+In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places \textbf{~\citep{HAL_2018}}.
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
@@
 
-In Cascadia, there are robust signals in both GNSS and tremor ~\citep{HAW_2013}.
-This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS.
+In Cascadia, there are robust signals in both GNSS and tremor \textbf{~\citep{HAW_2013}}.
+This is also the case in Nankai \textbf{~\citep{HIR_2008}}, where tiltmeters are used instead of GNSS.
 It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data.
@@
 For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University.
-These are network solutions in ITRF2008 with phase ambiguities resolved.
-Solutions are computed with JPL/NASA orbits and satellite clocks.
+These are network solutions \textbf{in ITRF2014 with phase ambiguities resolved with wide-lane phase-biases and orbits and satellite clocks provided by the Jet Propulsion Laboratory/NASA}.
 North, East, and Vertical directions are available.
@@
 The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$.
-If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
+If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ \textbf{is some integer greater than or equal to 1}, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
 , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$.
@@
 
-To better evaluate the number of true and false detections, we convert the wavelet details into binary time series.
+To better evaluate the number of true and false detections, we convert the wavelet details into \textbf{trinary} time series.
 If the absolute value of the wavelet detail is higher that a threshold, we replace the value by 1 (for positive values) or -1 (for negative values), otherwise we replace the value by 0.
@@
 The ROC curve is widely use for binary classification problems in statistics and machine learning.
-We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
+We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into \textbf{trinary} time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
 If the classification was made randomly, all the points would fall on the first diagonal.
@@
 
-As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a binary time series.
+As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a \textbf{trinary} time series.
 We apply this transform to both the GPS and the tremor time series with varying thresholds.
@@
 If we increase the number of true events detected, we also increase the false positive rate.
-In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
+\textbf{If we increase the threshold for the tremor, the curve goes farther away from the first diagonal, that is we get better classification results.
+If we increase the threshold for the GPS, the false positive rate and the the number of events detected decrease.} In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75 \textbf{(that is the false positive rate is lower than 0.25)}, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
 \\
@@
 \item Figure 2.
-Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).
+Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}.
+The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).
 
@@
 
-\item Figure 9. Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
+\item Figure 9.
+ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
+The \textbf{black} cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure \textbf{10}.
+\textbf{The values of the threshold are color-coded.
+Reds correspond to the lowest value of the threshold for the tremor (0.001), while oranges, greens, blues, purples correspond to increasing values of the threshold for the tremor (up to 0.01).
+The brightest colors correspond to the highest values of the threshold for the GPS (1.5), while the darker and darker colors correspond to decreasing values of the threshold for the GPS (0.1).}
 
-\item Figure 10.
-ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
-Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
+\item Figure 10. Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
 
@@
 \begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).}
+\noindent\includegraphics[width=4.5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
+\caption{Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}.
+The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).}
 \label{pngfiguresample}
@@
 \begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/ROC_6-7-8.eps}
+\caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
+The \textbf{black} cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure \textbf{10}.
+\textbf{The values of the threshold are color-coded.
+Reds correspond to the lowest value of the threshold for the tremor (0.001), while oranges, greens, blues, purples correspond to increasing values of the threshold for the tremor (up to 0.01).
+The brightest colors correspond to the highest values of the threshold for the GPS (1.5), while the darker and darker colors correspond to decreasing values of the threshold for the GPS (0.1).}}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6-7-8.eps}
@@
 \caption{Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/ROC_6-7-8.eps}
-\caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
-Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.}
 \label{pngfiguresample}

### 2022-04-09 src/compare_tremor_GPS.py

@@ -17,6 +17,7 @@ from matplotlib.colors import Normalize
 
 import date
 import DWT
+from matplotlib_cmap_generator import red, green, blue, generateCmap
 from MODWT import get_DS, get_scaling, pyramid
 
 def correlate(X, Y, ncor):
@@ -468,6 +469,7 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
             filename = 'MODWT_GPS_longer/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
             if ((np.min(time) <= 2021.5) and (np.max(time) >= 2006.0)):
+                print(index, station)
                 times.append(time)
                 disps.append(disp)
                 Ws.append(W)
@@ -507,10 +509,10 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 if (len(indices) > 0):
                     nsta = nsta + 1
                     tj = time[indices]
-                    Dj = D[J][indices]
-#                    Dj = np.zeros(len(indices))
-#                    for j in J:
-#                        Dj = Dj + D[j - 1][indices]
+#                    Dj = D[J][indices]
+                    Dj = np.zeros(len(indices))
+                    for j in J:
+                        Dj = Dj + D[j - 1][indices]
                     Dj_interp = np.interp(time_subset, tj, Dj)
                     Dj_stacked =  Dj_stacked + Dj_interp
             Dj_stacked = Dj_stacked / nsta
@@ -542,26 +544,26 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
                 for i in range(0, len(jumps) + 1):
                     x0 = begin_times[i]
                     dx = end_times[i] - begin_times[i]
-#                    if disps_slowslip[begin_jumps[i]] > 0:
-#                        ax.add_patch(Rectangle((x0, lat + 0.01), dx, 0.03, facecolor='red'))
-#                    else:
-#                        ax.add_patch(Rectangle((x0, lat - 0.04), dx, 0.03, facecolor='blue'))
[…]

### 2022-04-09 src/matplotlib_cmap_generator.py

@@ -1,0 +1,142 @@
+# -*- coding: utf-8 -*-
+"""
+@author: Yichen Wang
+@date: 06/30/2020
+"""
+import matplotlib.pyplot as plt
+from matplotlib import cm, colors
+import numpy as np
+
+def red(x, xmin, xmax):
+    """
+    Amount of red in color for value x between xmin and xmax
+    """
+    if x < xmin + 0.2 * (xmax - xmin):
+        level = 1.0
+    elif x < xmin + 0.4 * (xmax - xmin):
+        level = 1.0 - (x - xmin - 0.2 * (xmax - xmin)) / (0.2 * (xmax - xmin))
+    elif x < xmin + 0.8 * (xmax - xmin):
+        level = 0.0
+    else:
+        level = (x - xmin - 0.8 * (xmax - xmin)) / (0.2 * (xmax - xmin))
+    return level
+
+def green(x, xmin, xmax):
+    """
+    Amount of green in color for value x between xmin and xmax
+    """
+    if x < xmin + 0.2 * (xmax - xmin):
+        level = (x - xmin) / (0.2 * (xmax - xmin))
+    elif x < xmin + 0.6 * (xmax - xmin):
+        level = 1.0
+    elif x < xmin + 0.8 * (xmax - xmin):
+        level = 1.0 - (x - xmin - 0.6 * (xmax - xmin)) / (0.2 * (xmax - xmin))
+    else:
+        level = 0
+    return level
+
+def blue(x, xmin, xmax):
+    """
+    Amount of blue in color for value x between xmin and xmax
+    """
+    if x < xmin + 0.4 * (xmax - xmin):
+        level = 0.0
+    elif x < xmin + 0.6 * (xmax - xmin):
+        level = (x - xmin - 0.4 * (xmax - xmin)) / (0.2 * (xmax - xmin))
+    else:
+        level = 1.0
+    return level
+
+def generateCmap(breakPointColors, breakPoints=None, sensitivity=256, 
+               name="userCmap"):
+    '''
+    Aimed at generate user customized ListedColormap instance, to be used when 
+    plotting with matplotlib.
+    Arguments:
+    ----------
+    breakPointColors: 
+        An array of color difinitions. Can be "rgba" vector or a string (like 
+        "red") that `matplotlib.colors.to_rgba()` accepts.
+    breakPoints: 
+        Optional. An array of numbers that defines the border of color 
+        gradients. The values does not necessarily match the border of the 
+        data to plot, but will be reflected as ratio on the plot. Must be of 
[…]

### 2022-04-09 src/missing_values.py

@@ -5,6 +5,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 from math import log, sqrt
+from matplotlib import ticker
 
 import DWT, MODWT
 from MODWT import get_DS, get_scaling, inv_pyramid, pyramid
@@ -82,6 +83,7 @@ for n in range(1, nmax, nstep):
 
         # Plot data
         ax = plt.subplot2grid((J + 2, len(gaps)), (0, count))
+        ax.xaxis.set_major_formatter(ticker.FormatStrFormatter('%.1f'))
         plt.plot(time, disp_interp, 'r')
         plt.plot(time, disp, 'k', label='Data')
         plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))
@@ -89,12 +91,14 @@ for n in range(1, nmax, nstep):
         # Plot details at each level
         for j in range(0, J):
             ax = plt.subplot2grid((J + 2, len(gaps)), (j + 1, count))
+            ax.xaxis.set_major_formatter(ticker.FormatStrFormatter('%.1f'))
             plt.plot(time, D[j], 'r')
             plt.plot(time, D0[j], 'k', label='D' + str(j + 1))
             plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))
             plt.legend(loc=1)
         # Plot scaling coefficients for the last level
         ax = plt.subplot2grid((J + 2, len(gaps)), (J + 1, count))
+        ax.xaxis.set_major_formatter(ticker.FormatStrFormatter('%.1f'))
         plt.plot(time, S[J], 'r')
         plt.plot(time, S0[J], 'k', label='S' + str(J))
         plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))

### 2022-04-09 src/tremor_MODWT_longer.py

@@ -140,7 +140,7 @@ def compute_wavelets_tremor(station_file, lats, lons, dataset, direction, \
         minD = min([np.min(Dj) for Dj in D])
 
         # Plot data
-        plt.subplot2grid((J + 2, 1), (J + 1, 0))
+        plt.subplot2grid((J + 2, 1), (0, 0))
         plt.plot(times_stacked, tremor_detrend, 'k', label='Data')
         plt.xlabel('Time (years)', fontsize=24)
         plt.xlim([2006.0, 2021.5])
@@ -148,13 +148,13 @@ def compute_wavelets_tremor(station_file, lats, lons, dataset, direction, \
         plt.legend(loc=3, fontsize=20)
         # Plot details
         for j in range(0, J):
-            plt.subplot2grid((J + 2, 1), (J - j, 0))
+            plt.subplot2grid((J + 2, 1), (j + 1, 0))
             plt.plot(times_stacked, D[j], 'k', label='D' + str(j + 1))
             plt.xlim([2006.0, 2021.5])
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
         # Plot smooth
-        plt.subplot2grid((J + 2, 1), (0, 0))
+        plt.subplot2grid((J + 2, 1), (J + 1, 0))
         plt.plot(times_stacked, S[J], 'k', label='S' + str(J))
         plt.xlim([2006.0, 2021.5])
         plt.ylim([np.min(tremor_detrend), np.max(tremor_detrend)])

### 2022-04-10 src/DWT.py
Finish NZ analysys plus start convert KV script

@@ -21,55 +21,55 @@ def get_scaling(name):
         g = Vector of coefficients of the scaling filter
     """
     if (name == 'Haar'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/Haar.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/Haar.dat')
     elif (name == 'D4'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D4.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D4.dat')
     elif (name == 'D6'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D6.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D6.dat')
     elif (name == 'D8'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D8.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D8.dat')
     elif (name == 'D10'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D10.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D10.dat')
     elif (name == 'D12'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D12.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D12.dat')
     elif (name == 'D14'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D14.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D14.dat')
     elif (name == 'D16'):
-        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D16.dat')
+        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D16.dat')
[…]

### 2022-04-10 src/compare_tremor_GPS_NZ.py

@@ -20,7 +20,7 @@ import DWT
 from MODWT import get_DS, get_scaling, pyramid
 
 
-def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold):
+def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, events, possible_events):
     """
     """
     # Read station file
@@ -28,14 +28,16 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold):
     stations.columns = ['name', 'longitude', 'latitude']
 
     # Time limits
-    tmin = 2000 + 909 / 365.25
-    tmax = 2000 + 8127 / 365.25
+#    tmin = 2000 + 909 / 365.25
+#    tmax = 2000 + 8127 / 365.25
+    tmin = 2010.0
+    tmax = 2016.0
 
     a = 6378.136
     e = 0.006694470
 
     # Start figure
-    fig, ax = plt.subplots(figsize=(16, 8))
+    fig, ax = plt.subplots(figsize=(16, 9))
     params = {'xtick.labelsize':24,
               'ytick.labelsize':24}
     pylab.rcParams.update(params)
@@ -149,6 +151,15 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold):
 
             plt.plot(times_stacked, lat + 0.1 * disps_stacked, color='black')
 
+    # Add slow slip events from Todd and Schwartz (2016)
+#    for event in events:
+#        plt.plot([event['time'], event['time']], \
+#            [event['latmin'], event['latmax']], color='red', linewidth=4)
+#    for event in possible_events:
+#        plt.plot([event['time'], event['time']], \
+#            [event['latmin'], event['latmax']], color='red', linewidth=4, \
+#            linestyle='dotted')
+
     plt.xlim([tmin, tmax])
     plt.xlabel('Time (years)', fontsize=24)
     plt.xticks(fontsize=24)
@@ -166,21 +177,52 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold):
 if __name__ == '__main__':
 
     station_file = '../data/GeoNet/stations.txt'
[…]

### 2022-04-10 src/find_ETS.py

@@ -1,0 +1,100 @@
+"""
+Script to find automatically the ETS events
+"""
+
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import pickle
+
+#  analyze the sum of wavelets for details 6-8
+details = '6-8'
+
+# start and end times of wavelets and tremor catalogs
+startTime = 2006.8
+endTime = 2021.3
+
+# Latitudes and Longitudes for 16 reference points along strike
+lats = [47.20000, 47.30000, 47.40000, 47.50000, 47.60000, 47.70000, 47.80000, 47.90000, 48.00000, \
+    48.10000, 48.20000, 48.30000, 48.40000, 48.50000, 48.60000, 48.70000]
+lons = [-122.74294, -122.73912, -122.75036, -122.77612, -122.81591, -122.86920, -122.93549, \
+    -123.01425, -123.10498, -123.20716, -123.32028, -123.44381, -123.57726, -123.72011, -123.87183, -124.03193]
+Nlat = len(lats)
+
+# Read in all the GPS wavelets
+# Redefine time vectors with errors up to +/- 4 days
+# time for 1-4 are the same but have 10 extra samples so remove 10 evenly spaced time samples
+# time for 16 has one extra sample so remove the middle one
+# time for 5-15 are the same
+# time for each latitude point is replaced by time vector for latitude 12
+# start and end times are the same for 1-16
+
+Dat_T = np.zeros((4, 16, 5578))
+Dat_Y = np.zeros((4, 16, 5578))
+
+for detail in [5, 6, 7, 8]:
+    for lat_id in np.arange(0, 16):
+        fileName = 'files/GPS_detail_' + str(detail) + '_loc_' + str(lat_id) + '.txt'
+        tmp = np.loadtxt(fileName)
+        klat = lat_id + 1
+        tmp1 = tmp[:, 0]
+        tmp2 = tmp[:, 1]
+        if klat <= 4:
+            krm = np.int_(np.floor(np.arange(0, 10) * len(tmp) / 11))
+            tmp1 = np.delete(tmp1, krm)
+            tmp2 = np.delete(tmp2, krm)
+        if klat == 16:
+            krm = np.int_(np.floor(0.5 * len(tmp)))
[…]

### 2022-04-10 src/vespagram_NZ.py

@@ -49,6 +49,7 @@ def compute_wavelets(station_file, direction, wavelet, J):
         time = np.array(data['date-time'].apply( \
             lambda x: datetime.date.toordinal(x)))
         cname = ' ' + direction + ' (mm)'
+
         time = 2000 + (time - 730120) / 365.25
         disp = np.array(data[cname])
         sigma = np.std(disp)
@@ -126,7 +127,7 @@ def compute_wavelets(station_file, direction, wavelet, J):
         plt.tight_layout()
         plt.savefig('MODWT_GPS_NZ/' + station + '_' + direction + '.pdf', format='pdf')
         plt.close(1)
-        
+
 if __name__ == '__main__':
 
     station_file = '../data/GeoNet/stations.txt'

### 2022-04-14 paper/answer_reviewers.tex
First version of revisions

@@
 
-We added an analysis of New Zealand GPS data.
+We added a section about the analysis of GPS data from New Zealand to show how the method could still be applied if tremor data are not available.
 
@@
 
+\textcolor{red}{I am not sure what to answer to this one.
+Should we try to make a zoom of Figure 9 over a shorter time window around the time of this possible slow slip event and also show the corresponding raw GPS data?}
+
+\bigskip
+
 \textit{(Figure 2) The caption text is not clear, in particular the phrase "for two locations of the missing values" is vague.
@@
 
-(Figure 7) Why do the obvious red-to-blue transitions in the tremor data seem to disappear from the detail level 7 figure, but are present in the detail level 8 (Figure 6) and detail level 6 (Figure 8) data?
-Can anything be interpreted from that pattern?
+\bigskip
 
-(Figure 9) When stacking levels 6, 7 and 8, won't the lower numbered levels dominate because of their (generally) larger amplitudes (as seen in Figure 3)?
-Is it valuable to normalize the levels before summing, or will this cause other issues?
+We added a table summarizing the number of stations and the number of overlapping stations for each location.
+There is no evidence for a single station misbehaving at time 2007.5 and several GPS stations show a small increase in the displacement in the eastern direction at that time.
+However, there are many missing values around time 2007.5, so it is difficult to conclude.
 
-(Figure 10) Is it possible to select a threshold in the absence of a catalog or does threshold selection depend on slow slip events already being know from other methods?
+\bigskip
 
-(Figure 11) What are the dates selected by this algorithm?
+\textit{(Figure 7) Why do the obvious red-to-blue transitions in the tremor data seem to disappear from the detail level 7 figure, but are present in the detail level 8 (Figure 6) and detail level 6 (Figure 8) data?
+Can anything be interpreted from that pattern?}
+
+\bigskip
+
+The number of red-to-blue transitions depends on the threshold chosen.
+For detail level 7, the false positive rate is smaller than for details levels 6 and 8, so all the transitions may not be seen clearly.
+
+\bigskip
+
+\textit{(Figure 9) When stacking levels 6, 7 and 8, won't the lower numbered levels dominate because of their (generally) larger amplitudes (as seen in Figure 3)?
+Is it valuable to normalize the levels before summing, or will this cause other issues?}
+
+\bigskip
+
+The amplitude of level 6 is not much higher than the amplitude of level 8, so we do not think that it would dominate the stacking.
+A more complex way of stacking the coefficients would be to apply a wavelet-based denoising method, where each wavelet vector $\widetilde{W}_j$ is modified before reconstructing the initial time series by applying an inverse wavelet transform to the wavelet vectors.
+We could then stacked the denoised GPS time series instead of stacking the wavelet details.
+
+\bigskip
+
+\textit{(Figure 10) Is it possible to select a threshold in the absence of a catalog or does threshold selection depend on slow slip events already being know from other methods?}
+
+\bigskip
+
+It is possible to choose a threshold without a tremor catalog.
+However, without the comparison with the tremor, we cannot decide which slow slip events are real and which slow slip events are false detections and we cannot produce the ROC curve.
+
+\bigskip
+
+\textit{(Figure 11) What are the dates selected by this algorithm?
 How do they compare with the dates given in Michel et al., 2019?
-Most look spot on, but others, in particular the 2009 event look like they are quite distant in time.
+Most look spot on, but others, in particular the 2009 event look like they are quite distant in time.}
 
-There are some awkward sentences as well:
+\bigskip
 
- (Line 190-191): "where n is some integer greater than or equal to" sounds better than "higher or equal".
+The timing of the events are taken from the plot of moment rate as a function of time from Figure 8 of Michel et al.
+(2019).
+For each event, we choose the middle of the time window as the time of the slow slip event.
 
-(Line 199): The terminology "details" and "smooths" sounds awkward, is that the standard terminology in the wavelet literature?
+\bigskip
 
-(Line 287-288) The sentence "The straight line starts at and ends at." is incomplete.
+\textit{There are some awkward sentences as well:} \\
+
+\textit{ (Line 190-191): "where n is some integer greater than or equal to" sounds better than "higher or equal".}
+
+\bigskip
+
+We modified this sentence.
+
+\bigskip
+
+\textit{(Line 199): The terminology "details" and "smooths" sounds awkward, is that the standard terminology in the wavelet literature?}
+
+\bigskip
+
+Yes, this is the standard terminology in the wavelet literature.
+
+\bigskip
+
+\textit{(Line 287-288) The sentence "The straight line starts at and ends at." is incomplete.}
+
+\bigskip
+
+We removed the incomplete sentence.
+
+\bigskip
 
@@
 
-The authors present a wavelet-based method to identify slow slip events within the GNSS position time series.
+\bigskip
+
+\textit{The authors present a wavelet-based method to identify slow slip events within the GNSS position time series.
 Rather than looking at the geodetic record of slow slip as a sum of periodic basis functions, the wavelet method allows for the isolation of slow slip signals across different time scales at different times during the period of study.
-The approach is of interest to the community and I think that this manuscript will eventually be publishable.
+The approach is of interest to the community and I think that this manuscript will eventually be publishable.} \\
 
-That said, my major concern is that the method as presented will be very difficult to apply to another region of study, especially one without a systematically generated catalog of tectonic tremor or low-frequency earthquakes.
+\textit{That said, my major concern is that the method as presented will be very difficult to apply to another region of study, especially one without a systematically generated catalog of tectonic tremor or low-frequency earthquakes.
 Here the authors validate the detections (i.e.
@@
 (2019) (slow slip); such detailed and long-duration catalogs are not available outside of Japan to validate the results of the presented method.
-Aggravating my concern is the premise that the authors themselves set up in the introduction of the manuscript, that a method is needed to detect slow slip when tremor is not available as an independent control; the presented method does not reach that
-goal.
-I would also highlight that slow slip in northern Cascadia is some of the most well behaved slow slip across all subduction zones, with short-ish recurrence intervals and consistently clear geodetic signals captured by a relatively dense GNSS network, that will likely show up more reliably than elsewhere with this sort of approach.
+Aggravating my concern is the premise that the authors themselves set up in the introduction of the manuscript, that a method is needed to detect slow slip when tremor is not available as an independent control; the presented method does not reach that goal.
+I would also highlight that slow slip in northern Cascadia is some of the most well behaved slow slip across all subduction zones, with short-ish recurrence intervals and consistently clear geodetic signals captured by a relatively dense GNSS network, that will likely show up more reliably than elsewhere with this sort of approach.}
 
-The figures for the most part are clear, but there are some improvements to be made (see comments below).
+\bigskip
+
+We added an application to GPS data in New Zealand, where tremor is not well spatially and temporally correlated, to show the broader applicability of the method.
+
+\bigskip
+
+\textit{The figures for the most part are clear, but there are some improvements to be made (see comments below).
 I also find that the introduction is lacking details and there are way too few articles cited for background reading for interested readers (cf.
-first paragraph with only two unique citations that discusses observations made across many subduction zones).
+first paragraph with only two unique citations that discusses observations made across many subduction zones).}
 
-I think a revised version of this manuscript that has reframed the purpose of the approach will be an interesting contribution that is appropriate to be published in BSSA.
-I list other minor comments below:
+\bigskip
 
-general - There is no discussion of common mode signals that will surely stack constructively across GNSS stations.
+We added several references in the introduction.
+
+\bigskip
+
+\textit{I think a revised version of this manuscript that has reframed the purpose of the approach will be an interesting contribution that is appropriate to be published in BSSA.
+I list other minor comments below:} \\
+
+\textit{general - There is no discussion of common mode signals that will surely stack constructively across GNSS stations.
 The removal of sinusoidal seasonal signals will remove some common mode signal, but there can still exist some non-tectonic signals that will stack constructively within the bandwidth of interest here.
-There is a reasonable chance that this will have little impact on the results, but it should at least be verified/discussed given the proposed processing.
+There is a reasonable chance that this will have little impact on the results, but it should at least be verified/discussed given the proposed processing.}
 
-general - I would avoid using the term ETS as it only adds confusion, as lines 69-74 highlight.
+\bigskip
+
+We added a short discussion on common modes.
+
+\bigskip
+
+\textit{general - I would avoid using the term ETS as it only adds confusion, as lines 69-74 highlight.
 I would suggest using "slow slip" to describe geodetic observations of aseismic slip and "tectonic tremor" for the accompanying seismic signals.
@@
 As described, it is not obvious to those unfamiliar with slow slip and tremor that the aseismic moment of slow slip dominates the seismic moment of tremor, because ETS conflates the geodetic observations with the seismic ones.
-I would simply use the term slow slip when referring to the dominant aseismic transient.
+I would simply use the term slow slip when referring to the dominant aseismic transient.}
 
-general - I'm not sure that the "true detection" terminology used here is quite appropriate.
+\bigskip
+
+We replaced "ETS" by "slow slip" wherever it seemed more appropriate.
+
+\bigskip
+
+\textit{general - I'm not sure that the "true detection" terminology used here is quite appropriate.
 This implies that we have an absolute knowledge of what's happening on the plate interface, and that there can't be new detections made here without already be corroborated by other catalogs.
 Perhaps "robust detection" might be more appropriate?
-In any case, this underlines my above major comment that this method is difficult to interpret without other catalogs.
+In any case, this underlines my above major comment that this method is difficult to interpret without other catalogs.}
 
-line 42 - "feature" is a strange word choice and does not tell the reader much at all
+\bigskip
 
-line 60 - "it is more abundant in some places" doesn't tell us much
+"true detection" is the standard term usually used in the literature on machine learning methods for classification.
+We agree that it would be valid only if there was perfect correlation between slow slip and tremor.
+As it may not be the case, we replace the term by "robust detection".
 
-line 64 - "for which" -> whose
+\bigskip
 
-line 67 - Obara (2002) did not discuss slow slip, only tectonic tremor. I would suggest something like Obara et al., GRL, 2004
+\textit{line 42 - "feature" is a strange word choice and does not tell the reader much at all}
 
-line 126 - "levels" does not mean very much without context
+\bigskip
 
-lines 148 - "details" same comment as above; this general term doesn't mean anything without context
+We replaced "feature" by "phenomenon".
 
-line 151-152 - why is there supposed to be a difference between the two tidal gauges and what does that mean? Not clear without the reader being forced to go and read the cited article.
+\bigskip
 
-line 156 - I assume longitudinal displacements is supposed to mean east-west motion?
+\textit{line 60 - "it is more abundant in some places" doesn't tell us much}
 
-line 287-288 - missing text here
+\bigskip
 
-figure 2 - I would say figure 2 could be moved to the appendix or supplementary material. It's not key to the message of the manuscript
+We modified this sentence.
 
-figures 6-8 - I think this figure would be more easily interpretable as well as much effective if only one panel was used, with alternating geodetic and seismic data for a given location (red triangle from Figure 4)in two different colors (black and something else).
+\bigskip
+
+\textit{line 64 - "for which" -> whose}
+
+\bigskip
+
+We modified this sentence.
+
+\bigskip
+
+\textit{line 67 - Obara (2002) did not discuss slow slip, only tectonic tremor. I would suggest something like Obara et al., GRL, 2004}
+
+\bigskip
+
+We replaced the reference by Obara et al., 2004.
+
+\bigskip
+
+\textit{line 126 - "levels" does not mean very much without context}
+
+\bigskip
+
+We replaced "levels" by "time scales".
+
+\bigskip
+
+\textit{lines 148 - "details" same comment as above; this general term doesn't mean anything without context}
+
+\bigskip
+
+We replaced "details" by "components at different time scales".
+
+\bigskip
+
+\textit{line 151-152 - why is there supposed to be a difference between the two tidal gauges and what does that mean? Not clear without the reader being forced to go and read the cited article.}
+
+\bigskip
+
+We clarified this sentence.
+
+\bigskip
+
+\textit{line 156 - I assume longitudinal displacements is supposed to mean east-west motion?}
+
+\bigskip
+
+Yes, we modified this sentence.
+
+\bigskip
+
+\textit{line 287-288 - missing text here}
+
+\bigskip
+
+We removed this sentence.
+
+\bigskip
+
+\textit{figure 2 - I would say figure 2 could be moved to the appendix or supplementary material. It's not key to the message of the manuscript}
+
+\bigskip
+
+We moved Figure 2 to the Supplementary Material.
+
+\bigskip
+
+\textit{figures 6-8 - I think this figure would be more easily interpretable as well as much effective if only one panel was used, with alternating geodetic and seismic data for a given location (red triangle from Figure 4)in two different colors (black and something else).
 The red/blue rectangles that indicate detections could be four similar colors for geodetic/seismic detections (blue/purple for negative peaks and red/orange for positive peaks for example).
-I'd also suggest that figures 7 and 8 could be made into supplementary figures as they show similar information to figure 6
+I'd also suggest that figures 7 and 8 could be made into supplementary figures as they show similar information to figure 6}
 
-figure 7 - I'm not sure I understand why the threshold selected here is less effective (there are less events identified) than for levels levels 6 and 8. What could explain this?
+\bigskip
 
-figure 10 - I think this figure would be a bit clearer if the 6/7/8 detail levels were in different colors
+\textcolor{red}{I am not sure about this one. Wouldn't that make the figure even more difficult to understand?}
+
+\bigskip
+
+\textit{figure 7 - I'm not sure I understand why the threshold selected here is less effective (there are less events identified) than for levels levels 6 and 8. What could explain this?}
+
+\bigskip
+
+This is probably due to the value we have chosen for the threshold.
+In all cases, we tried to maximize the true positive rate and minimize the false positive rate.
+The true positive rate is higher for level 7 than for level 8 and the false positive rate is lower for level than for level 6 and 8.
+The downsize is that there are more false negative (missed detections) for the level 7.
+
+\bigskip
+
+\textit{figure 10 - I think this figure would be a bit clearer if the 6/7/8 detail levels were in different colors}
+
+\bigskip
+
+In Figure 10, we sum details 6, 7 and 8, we did not plot them separately.
 

### 2022-04-14 paper/paper_bssa_annotated.tex
First version of revisions

@@
 ~\citep{MIC_2019}).
+\textbf{We then apply the wavelet method to data from New Zealand and detect slow slip events spatially and temporally close to those detected previously by ~\citet{TOD_2016}.}
 
@@
 
-Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks \textbf{~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}}.
+Slow slip events are new \textbf{phenomena} discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks \textbf{~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}}.
 As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate.
@@
 
-In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places \textbf{~\citep{HAL_2018}}.
+In many places, tectonic tremor is also observed in relation to slow slip, but \textbf{the spatial correlation between tremor and slow slip may vary along the strike of the plate boundary ~\citep{HAL_2018}}.
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
-Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
-In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\citep{OBA_2002,ROG_2003}.
+Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) \textbf{whose} frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
+In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations \textbf{~\citep{ROG_2003,OBA_2004}}.
 Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS).
@@
 For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded.
-They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8.
+They observed a linear relationship between moment release and number of hours of tremor for \textbf{slow slip} events of moment magnitude 6.3 to 6.8.
 Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data.
@@
 The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains.
-A sharp change in the signal will be localized and seen at all levels of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform.
+A sharp change in the signal will be localized and seen at all \textbf{time scales} of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform.
 \\
@@
 
-Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with ETS events between 1996 and 2011.
-Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events.
+Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with \textbf{slow slip} events between 1996 and 2011.
+Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the \textbf{slow slip} events.
 The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other.
 By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified.
-By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal.
-The authors were then able to clearly see a difference in uplift between the two tidal gauges at Port Angeles and Port Townsend.
-\\
+By stacking the \textbf{components at different time scales} of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal.
+\textbf{Due to the relative location of the tidal gauges at Port Angeles and Port Townsend compared to the slo slip region on the plate boundary, a slow slip event should result in uplift in Port Angeles (western part) and in subsidence in Port Townsend (eastern part).
+Indeed, the authors were able to clearly see a difference in the sign of the uplift at these two tidal gauges.} \\
 
 In our study, we use a similar approach to previous studies with a different reasoning.
-We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered.
+We only stack signals at nearby GPS stations, assuming that the \textbf{east-west} displacement due to the \textbf{slow slip} events will then be the same at each of the GPS stations considered.
 We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking.
-Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
+Finally, we assume that the noise and the longitudinal displacement due to the \textbf{slow slip} events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the \textbf{slow slip} events.
 We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia.
@@
 We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods.
-\\
+\textbf{Finally, we apply the method to GPS data in New Zealand and successfully detect several slow slip events without needing to rely on the tremor data.} \\
 
@@
 
-We focused our study on northwest Washington State.
+We \textbf{first} focused our study on northwest Washington State.
 For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University.
@@
 \\
+
+\textbf{For the application to slow slip events in New Zealand, we used the GPS time series provided by the Geological hazard information for New Zealand (GeoNet).
+The coordinates have been extracted by GeoNet during the GLOBK run from the combined daily solution files, and converted to (east, north, up) displacement in millimeters with respect to an a priori position and epoch in the ITRF2008 realization.
+The time series provided by GeoNet have no adjustments made to them, so they may, for example, contain offsets due to earthquakes, offsets due to equipment changes at individual sites, and seasonal (annual and semi-annual) signals due to various causes.
+Here again, the direction of the secular plate motion is close to the West direction, so we only used the East direction of the GPS time series for the data analysis.
+We detrended the data before applying the wavelet transform by carrying a linear regression of the whole time series and removing the straight line obtained from the regression.}
 
@@
 The remaining missing values are computed by doing a linear interpolation of the first and the last values and adding a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series.
-The straight line starts at and ends at .
 We verify how the wavelet details may be affected by looking at a GPS time series without missing values and compared the wavelet details with and without removing some data points.
@@
 Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values.
-Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values.
+\textbf{Figure S1 in the Supplementary Material} shows a comparison of the two wavelet details for two different locations of the missing values.
 We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition.
@@
 We then applied the wavelet filtering to real GPS data.
-Figure 3 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth.
+Figure 2 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth.
 In the data, we can see a sharp drop in displacement whenever there is a documented slow slip event.
@@
 We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal.
-We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4).
+We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 3).
 Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations.
@@
 
-To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for ETS events.
+To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for \textbf{slow slip} events.
 We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3.
@@
 Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform.
-Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5).
+Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5).
 Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
 
-\section{Results}
+\section{Application to data from Cascadia}
 
 We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-The result is shown in the top panel of Figure 6, where each line represents one of the locations along strike.
+The result is shown in the top panel of Figure 5, where each line represents one of the locations along strike.
 To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold.
-To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6.
+To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 5.
 We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
@@
 
-A similar comparison is shown for the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively (Figures 7 and 8).
+A similar comparison is shown for the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively (Figures 6 and 7).
 The events are harder to see in the 7th level than in the 8th level, both for the GPS data and the tremor count data.
@@
 There are three small signals in the GPS data in Winter 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and may be false detections.
-To summarize, we assume that true detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series.
-Then, all the 13 events present on the 8th level detail of the wavelet decomposition are true detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
+To summarize, we assume that \textbf{robust} detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series.
+Then, all the 13 events present on the 8th level detail of the wavelet decomposition are \textbf{robust} detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are \textbf{robust} detections.
 \\
 
-\section{Discussion}
-
-To better evaluate the number of true and false detections, we convert the wavelet details into \textbf{trinary} time series.
+To better evaluate the number of \textbf{robust} and false detections, we convert the wavelet details into \textbf{trinary} time series.
 If the absolute value of the wavelet detail is higher that a threshold, we replace the value by 1 (for positive values) or -1 (for negative values), otherwise we replace the value by 0.
 We do this on both the wavelet details of the GPS data and of the tremor data.
-Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a true detection (true positive, TP).
+Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a \textbf{robust} detection (true positive, TP).
 If the GPS and the tremor time series have opposite signs, or if the absolute value of the GPS time series is 1 but the value of the tremor time series is 0, we have a false detection (false positive, FP).
@@
 We apply this transform to both the GPS and the tremor time series with varying thresholds.
-We then plot the ROC curve on Figure 9, each dot representing a different threshold.
-The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10.
+We then plot the ROC curve on Figure 8, each dot representing a different threshold.
+The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 9.
 We can see that there is a trade-off between sensitivity and specificity as we vary the threshold.
@@
 \textbf{If we increase the threshold for the tremor, the curve goes farther away from the first diagonal, that is we get better classification results.
-If we increase the threshold for the GPS, the false positive rate and the the number of events detected decrease.} In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75 \textbf{(that is the false positive rate is lower than 0.25)}, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
+If we increase the threshold for the GPS, the false positive rate and the the number of events detected decrease.} In Figure 9, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75 \textbf{(that is the false positive rate is lower than 0.25)}, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
 \\
@@
 Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event.
-Figure 11 shows the 5th level detail wavelet decomposition of the GPS data.
-Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
+Figure 10 shows the 5th level detail wavelet decomposition of the GPS data.
+Red lines show the timing of the big \textbf{slow slip} events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
 \\
@@
 However, some other events that are not in the catalog of ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
-Therefore, it is difficult to differentiate between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
+Therefore, it is difficult to differentiate between a \textbf{robust} detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
 \\
@@
 Finally, there is also a small decrease in displacement at about 2020.2 that is difficult to interpret.
+\\
+
+\textbf{Due to the short distances between the GPS stations and the locations of the red triangles on the map from Figure 3, the same station could be used multiple times for the stacking at different locations.
+When considering two different locations, the stacking is thus made over an overlapping number of stations.
+Table 3 summarizes the number of stations and the number of overlapping stations for each location on Figure 3.
+We hypothesize that the small displacement in the eastern direction seen at about 2007.5 could be due to a misbehaving station common to several locations.
+However, several GPS stations indeed show an increase in the displacement in the eastern direction at about 2007.5.
+There are many missing data around that time, so it is difficult to conclude.} \\
+
+\textbf{Another possibility is that common mode signals could stack constructively across GNSS stations and produce peaks in the wavelet details that are actually due to non-tectonic signals.
+We computed common mode signals for different latitude bins (each bin has width equal to half-a-degree of latitude) following the same method as \citet{NUY_2021}.
+We first stacked all the time series for the stations in each latitude bin that are located more than 100 km east of the 40 km depth contour of the plate boundary.
+We assume that these stations are not sensitive to the deformation on the plate interface.
+We then apply a yearly moving average to each common mode signal in order to remove any leftover noise.
+The common mode signal was then removed from the GNSS time series depending on each site’s latitude.
+Figure S2 in the Supplementary Information shows the corresponding sum of the stacks of the 6th, 7th and 8th wavelet details obtained from the resulting time series.
+The common modes seem to have little impact on the results and do not explain the additional four small events that we noted in Figure 9.}
+
+\section{Application to data from New Zealand}
+
+\textbf{We now apply our wavelet-based method to detect slow slip events in New Zealand.
+The tectonics of the North Island of New Zealand are dominated by the westward subduction of the Pacific Plate under the Australian Plate at the Hikurangi Trench.
+Two types of slow slip events have been observed at the Hikurangi margin.
+Shallow (10-15 km depth), shorter (1-3 weeks), and usually smaller (Mw 6.3-6.8) slow slip events have been observed every 18-24 months in the northern part of the margin.
+Deeper (35-60 km depth), longer (12-18 months), and larger (Mw 7.0) slow slip events have been observed every 5 years in the southern part of the
+margin ~\citep{WAL_2010,TOD_2016}.
+It used to be thought that there were no tremor associated with slow slip events in northern Hikurangi.
+~\citet{DEL_2009} observed an increase in the rate of microseismicity downdip
+of the 2004 Gisborne slow slip event.
+More recently, however, ~\citep{KIM_2011} detected a low level of tremor activity that increased during the 2010 Gisborne slow slip event.
+As was the case for the microearthquakes, the source of the tremor was located downdip of the slow slip patch determined from GNSS data.
+~\citep{IDE_2012} detected tremor downdip of the location of two deep slow slip events observed by ~\citet{WAL_2013} in 2006 and 2008.
+However, contrary to ETS events in Cascadia and Nankai, the tremor activity did not seem to increase during the slow slip events.
+~\citet{TOD_2016}  detected tremor associated with most of the shallow slow slip events between 2010 and 2015, and located downdip of the geodetically inferred slip area.
+They also detected deeper tremor between 20 and 50 km depth with unclear origin.
+They hypothesized that these tremor may be related to currently undetected deep long-term slow slip events.} \\
+
+\textbf{As we have done for Cascadia, we take all GPS stations located in a 50 km radius of a given point, for the 18 locations indicated in Figure 11, and we stack the 6th level details, the 7th level details or the 8th level details over all the GPS stations.
+We then sum together the 6th, 7th and 8th levels stacked wavelet details (Figure 12, top panel).
+We highlight positive and negative peaks with red and blue colors as was done in Figure 9.
+As we cannot use the tremor data to decide what is the appropriate threshold above which we consider that there is a slow slip event, we  assume that slow slip events in New Zealand will result in a surface displacement of the same amplitude as the displacement observed in Cascadia.
+Therefore, the amplitudes of the peaks in the wavelet details should also be the same and we take identical thresholds for both regions.
+As a slow slip event in northern New Zealand results in a displacement in the east direction at the Earth’s surface, the slow slip events are indicated by a negative peak followed by a positive peak in the stacked wavelet details.
+We compare the results to the timings and locations of the slow slip events detected by ~\citet{TOD_2016}.
+As they only used data from five GPS stations (PUKE?
+ANAU, GISB, MAHI and CKID), we indicate by a vertical grey bar on the bottom panel of Figure 12 each time a slow slip event was detected for this stations.
+The grey bars are centered on the latitudes of the GPS stations.
+If a slow slip event was detected by more than one station, all the corresponding grey bars are linked together to show the spatial extent of the slow slip.
+~\citet{TOD_2016} indicated by a question mark (on their Figure 2 and their Table 1) additional possible events, and those are indicated by a dotted grey bar on Figure 12.} \\
+
+\textbf{We observed that there is a good correlation between the events detected with the wavelet method and the events previously detected by ~\citet{TOD_2016}.
+It is unclear if the event near station ANAU in early 2010 (event 1 from ~\citet{TOD_2016}) is visible in the wavelet details, but we clearly see an event propagating from south to north in January-February and March-April 2010, (events 2 and 3 from ~\citet{TOD_2016}).
+The June-July 2010 event (event 4 in ~\citet{TOD_2016}) and the August 2010 event (event 5 in ~\citet{TOD_2016}) are not clearly visible in the wavelet details.
+There is an event in April-May 2011 in the northern part of the region studied (events 6 and 7 in ~\citet{TOD_2016}).
+We then see an event propagating south to north in August-September and September-October 2011 (events 8 and 9 in ~\citet{TOD_2016}), followed by another event in December 2011 (event 10 in ~\citet{TOD_2016}).
+Although ~\citep{TOD_2016} only detected this event for GPS station GISB, it seems that this event may have also extended farther to the north and the south.
+We do not detect event 11 from ~\citet{TOD_2016} in March 2012 with our wavelet method.
+We then see a slow slip event in the northern part of the region studied in August 2012 (event 12 in ~\citet{TOD_2016}), followed by another event in December 2012-January 2013 (event 13 in ~\citet{TOD_2016}).
+There is a slow slip event in the southern part of the region studied in February-March 2013 (event 14 in ~\citet{TOD_2016}.
+We can an event propagating from south to north in June-July and July-August 2013 (events 15 and 16 in ~\citet{TOD_2016}).
+The events in September-October 2013, December 2013 and May-June 2014 (events 17, 18 and 19 in ~\citet{TOD_2016}) are not clearly seen in the wavelet details, but there could be a small negative peak followed by a small positive peak at these times.
+We can then see a big slow slip event in September 2014 (events 20 and 21 in ~\citet{TOD_2016}) and a slow slip event in the southern part of the region studied in December 2014-January 2015 (events 22 and 23 in ~\citet{TOD_2016}).
+The events in January-February and February 2015 (events 24 and 25 in ~\citet{TOD_2016}) are not clearly visible in the wavelet details, but there could be a small negative peak followed by a small positive peak at these times.
+Finally, we can see an event in June-July 2015 in the northern part of the region studied (event 26 in ~\citet{TOD_2016}).
+Additionally, there could be two other events that are not in \citep{TOD_2016} in Fall 2010 (southern part of the region studied) and in Fall 2015.} \\
+
+\textbf{Our wavelet-based method thus works well to detect transients in GPS data that could be slow slip events, even in the absence of tremor data.
+The choice of the appropriate threshold to decide that there is a transient and the levels of the wavelet details that we look at for the detection may still not be easily made.
+In particular, as there is more time between two slow slip events in New Zealand than in Cascadia, the biggest slow slip events (early 2010, late 2011, 2013 and late 2014) can also be ween on the 9th level detail for New Zealand, whereas they could not be seen for Cascadia.} \\
 
@@
 In this paper, we develop and test a new approach for detecting transient events in GPS time series, such as slow slip events.
-We used wavelet methods to analyze GNSS time series and tremor recordings of slow slip events in Cascadia.
+We used wavelet methods to analyze GNSS time series and tremor recordings of slow slip events in Cascadia\textbf{, and GNSS time series in New Zealand}.
 We used detrended GNSS data, applied the MODWT transform, and stacked the wavelet details over several nearby GNSS stations.
@@
 ~\citep{MIC_2019}).
-We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between true detections and false detections.
+We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between \textbf{robust} detections and false detections.
+\textbf{We then applied the method to GNSS data in New Zealand and detected slow slip events consistent with the events previously detected by ~\citet{TOD_2016}.}
 
@@
 The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/slowslip}.
-Figure 4 was created using GMT ~\citep{WES_1991}.
+Figures 3 and 11 were created using GMT ~\citep{WES_1991}.
 
@@
 
+The authors would like to anonymous reviewers and the Associate Editor Jeanne Hardback, whose comments greatly helped improve the manuscript.
 This work was funded by the grant from the National Science Foundation EAR-1358512.
@@
 
+ \begin{table}[hbt!]
+ \caption{Number of GPS stations used for the stacking for each location on Figure 3 and number of common stations with the location immediately to the north and the location immediately to the south.}
+ \centering
+ \begin{tabular}{c c c c c}
+ \hline
+ Index & Latitude & Number of stations & Common stations (north) & Common stations (south) \\
+ \hline
+0 & 47.2 & 15 & 14 & \\
+1 & 47.3 & 18 & 17 & 14 \\
+2 & 47.4 & 24 & 20 & 17 \\
+3 & 47.5 & 21 & 20 & 20 \\
+4 & 47.6 & 22 & 14 & 20 \\
+5 & 47.7 & 17 & 12 & 14 \\
+6 & 47.8 & 13 & 8 & 12 \\
+7 & 47.9 & 10 & 9 & 8 \\
+8 & 48.0 & 10 & 7 & 9 \\
+9 & 48.1 & 8 & 7 & 7 \\
+10 & 48.2 & 10 & 8 & 7 \\
+11 & 48.3 & 9 & 9 & 8 \\
+12 & 48.4 & 9 & 5 & 9 \\
+13 & 48.5 & 7 & 5 & 5 \\
+14 & 48.6 & 6 & 5 & 5 \\
+15 & 48.7 & 5 & & 5 \\
+ \hline
+ \end{tabular}
+ \end{table}
+
 \newpage
@@
 
-\item Figure 2.
-Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}.
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).
+\item Figure 2. Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.
 
-\item Figure 3. Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.
-
-\item Figure 4.
+\item Figure 3.
 GPS stations used in this study (black triangles).
@@
 
-\item Figure 5. Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).
+\item Figure 4. Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).
 
-\item Figure 6.
+\item Figure 5.
 Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
@@
 
-\item Figure 7. Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).
+\item Figure 6. Same as Figure 5 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).
 
-\item Figure 8. Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).
+\item Figure 7. Same as Figure 5 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).
 
-\item Figure 9.
+\item Figure 8.
 ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
 Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The \textbf{black} cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure \textbf{10}.
+The \textbf{black} cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
 \textbf{The values of the threshold are color-coded.
@@
 
-\item Figure 10. Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
+\item Figure 9. Same as Figure 5 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
 
-\item Figure 11.
+\item Figure 10.
 Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
@@
 The blue lines represent the timings of the magnitude 5 events from the catalog of ~\citet{MIC_2019}.
+
+\item \textbf{Figure 11.
+GPS stations used for the slow slip detection in New Zealand (black triangles).
+The red triangles are the locations where we stack the GPS data.
+They are located close to the 20 km depth contour of the plate boundary from ~\citet{WIL_2013}.}
+
+\item \textbf{Figure 12.
+Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 11.
+We mark by a red rectangle every time where the amplitude is higher than a threshold equal to 0.4.
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
+Bottom: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition.
+We mark with a grey bar the slow slip events detected by ~\citet{TOD_2016}.}
 
@@
 The amplitude of the synthetic time series is normalized to 1, and the details and smooths show the relative amplitude.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=4.5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}.
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
-\caption{Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).}
+\caption{Same as Figure 5 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.eps}
-\caption{Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).}
+\caption{Same as Figure 5 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).}
 \label{pngfiguresample}
@@
 Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The \textbf{black} cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure \textbf{10}.
+The \textbf{black} cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
 \textbf{The values of the threshold are color-coded.
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
-\caption{Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).}
+\caption{Same as Figure 5 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).}
 \label{pngfiguresample}
@@
 
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations_NZ.eps}
+\caption{GPS stations used for the slow slip detection in New Zealand (black triangles).
+The red triangles are the locations where we stack the GPS data.
+They are located close to the 20 km depth contour of the plate boundary from ~\citet{WIL_2013}.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_without_events.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_only_events.eps}
+\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 11.
+We mark by a red rectangle every time where the amplitude is higher than a threshold equal to 0.4.
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
+Bottom: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition.
+We mark with a grey bar the slow slip events detected by ~\citet{TOD_2016}.}
+\label{pngfiguresample}
+\end{figure}
+
 \end{document}

### 2022-04-14 paper/supplement.tex
First version of revisions

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{authblk}
\usepackage{color}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{url}

\linenumbers

\title{Supplement to Detection of slow slip events using wavelet analysis of GNSS recordings}
\author[1]{Ariane Ducellier}
\author[2]{Kenneth C. Creager}
\author[2]{David A. Schmidt}
\affil[1]{Corresponding author. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
\affil[2]{University of Washington, Department of Earth and Space Sciences}
\date{}

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\begin{document}

\maketitle

\begin{figure}
\noindent\includegraphics[width=4.5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
\caption{Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}. The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_mode_details_6-7-8.eps}
\caption{Same as top panel of Figure 9 from the main text, but using the GNSS time series where common modes have been removed before applying the wavelet transform.}
\label{pngfiguresample}
\end{figure}

\end{document}

### 2022-04-14 src/DWT.py

@@ -21,55 +21,55 @@ def get_scaling(name):
         g = Vector of coefficients of the scaling filter
     """
     if (name == 'Haar'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/Haar.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/Haar.dat')
     elif (name == 'D4'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D4.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D4.dat')
     elif (name == 'D6'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D6.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D6.dat')
     elif (name == 'D8'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D8.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D8.dat')
     elif (name == 'D10'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D10.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D10.dat')
     elif (name == 'D12'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D12.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D12.dat')
     elif (name == 'D14'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D14.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D14.dat')
     elif (name == 'D16'):
-        g = np.loadtxt('/Users/ariane/Documents/ResearchProject/wmtsa/scalingcoeff/D16.dat')
+        g = np.loadtxt('/Users/ariane/Documents/Travail/wmtsa/scalingcoeff/D16.dat')
[…]

### 2022-04-14 src/common_modes.py

@@ -2,9 +2,16 @@
 Script to compute common modes for GPS stations
 """
 
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import pickle
+
+from math import cos, floor, pi, sin, sqrt
+
+import DWT
+from MODWT import get_DS, get_scaling, pyramid
 
 def detrend(time, disp):
     """
@@ -17,7 +24,7 @@ def detrend(time, disp):
         disp_detrend[i] = np.mean(disp[indices])
     return disp_detrend
 
-def stack(station_file, dataset, direction, lat_min, lat_max):
+def stack(station_file, dataset, direction, lat_min, lat_max, lon_min, lon_max):
     """
     Stack all GPS stations within a latitude bin
     """
@@ -26,7 +33,8 @@ def stack(station_file, dataset, direction, lat_min, lat_max):
         engine='python')
     stations.columns = ['name', 'longitude', 'latitude']
     subset = stations.loc[(stations['latitude'] >= lat_min) & \
-                          (stations['latitude'] <= lat_max)]
+                          (stations['latitude'] <= lat_max) & \
+                          (stations['longitude'] >= 0.5 * (lon_min + lon_max))]
 
     # Stack the detrended signal over GPS stations
     times = []
@@ -36,7 +44,6 @@ def stack(station_file, dataset, direction, lat_min, lat_max):
         data = np.loadtxt(filename, skiprows=26)
         time = data[:, 0]
         disp = data[:, 1]
-        disp = detrend(time, disp)
         times.append(time)
         disps.append(disp)
 
@@ -72,18 +79,20 @@ def stack(station_file, dataset, direction, lat_min, lat_max):
                 nsta = nsta + 1
                 disp_interp = np.interp(time_subset, time[indices], disp[indices])
                 disp_subset = disp_subset + disp_interp
-            disp_subset = disp_subset / nsta
+        disp_subset = disp_subset / nsta
         time_stack.append(time_subset)
         disp_stack.append(disp_subset)
     time_subset = np.concatenate(time_stack)
     disp_subset = np.concatenate(disp_stack)
     return(time_subset, disp_subset)
 
-def compute_common_modes(station_file, dataset, direction, latitudes):
[…]

### 2022-04-14 src/compare_tremor_GPS.py

@@ -437,9 +437,9 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     pylab.rcParams.update(params)
 
     for event in events:
-        plt.axvline(event, color='red', linewidth=2)
-    for event in small_events:
-        plt.axvline(event, color='blue', linewidth=1)
+        plt.axvline(event, color='black', linewidth=2)
+#    for event in small_events:
+#        plt.axvline(event, color='blue', linewidth=1)
 
     # Loop on latitude and longitude
     for index, (lat, lon) in enumerate(zip(lats, lons)):
@@ -466,10 +466,9 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
 
         # Read output files from wavelet transform
         for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
-            filename = 'MODWT_GPS_longer/' + dataset + '_' + station + '_' + direction + '.pkl'
+            filename = 'MODWT_GPS_mode/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
             if ((np.min(time) <= 2021.5) and (np.max(time) >= 2006.0)):
-                print(index, station)
                 times.append(time)
                 disps.append(disp)
                 Ws.append(W)
@@ -563,7 +562,7 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
     plt.title('Details at levels {} of MODWT of GPS data'. \
         format(J), fontsize=24)
 #    plt.savefig('GPS_longer_detail_' + str(J + 1) + '.eps', format='eps')
-    plt.savefig('GPS_longer_detail.eps', format='eps')
+    plt.savefig('GPS_mode_detail.eps', format='eps')
     plt.close(1)
 
 def plot_tremor(lats, J, threshold, events, small_events):
@@ -957,7 +956,7 @@ if __name__ == '__main__':
 
     thresh_GPS = np.arange(0.1, 1.6, 0.1)
     thresh_tremor = np.arange(0.001, 0.011, 0.001)
-    chosen_GPS = 0.7
+    chosen_GPS = 0.8
     chosen_tremor = 0.01
     J = [6, 7, 8]
 

### 2022-04-14 src/compare_tremor_GPS_NZ.py

@@ -144,21 +144,21 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, even
                 for i in range(0, len(jumps) + 1):
                     x0 = begin_times[i]
                     dx = end_times[i] - begin_times[i]
-                    if disps_slowslip[begin_jumps[i]] > 0:
-                        ax.add_patch(Rectangle((x0, lat + 0.01), dx, 0.03, facecolor='red'))
-                    else:
-                        ax.add_patch(Rectangle((x0, lat - 0.04), dx, 0.03, facecolor='blue'))
+#                    if disps_slowslip[begin_jumps[i]] > 0:
+#                        ax.add_patch(Rectangle((x0, lat + 0.01), dx, 0.03, facecolor='red'))
+#                    else:
+#                        ax.add_patch(Rectangle((x0, lat - 0.04), dx, 0.03, facecolor='blue'))
 
             plt.plot(times_stacked, lat + 0.1 * disps_stacked, color='black')
 
     # Add slow slip events from Todd and Schwartz (2016)
-#    for event in events:
-#        plt.plot([event['time'], event['time']], \
-#            [event['latmin'], event['latmax']], color='red', linewidth=4)
-#    for event in possible_events:
-#        plt.plot([event['time'], event['time']], \
-#            [event['latmin'], event['latmax']], color='red', linewidth=4, \
-#            linestyle='dotted')
+    for event in events:
+        plt.plot([event['time'], event['time']], \
+            [event['latmin'], event['latmax']], color='grey', linewidth=4)
+    for event in possible_events:
+        plt.plot([event['time'], event['time']], \
+            [event['latmin'], event['latmax']], color='grey', linewidth=4, \
+            linestyle='dotted')
 
     plt.xlim([tmin, tmax])
     plt.xlabel('Time (years)', fontsize=24)
@@ -170,8 +170,8 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, even
 #        format(J + 1), fontsize=24)
     plt.title('Details at levels {} of MODWT of GPS data'. \
         format(J), fontsize=24)
-#    plt.savefig('GPS_NZ_detail_' + str(J + 1) + '.pdf', format='pdf')
-    plt.savefig('GPS_NZ_detail.pdf', format='pdf')
+#    plt.savefig('GPS_NZ_detail_' + str(J + 1) + '.eps', format='eps')
[…]

### 2022-04-14 src/vespagram_NZ.py

@@ -125,7 +125,7 @@ def compute_wavelets(station_file, direction, wavelet, J):
         
         # Save figure
         plt.tight_layout()
-        plt.savefig('MODWT_GPS_NZ/' + station + '_' + direction + '.pdf', format='pdf')
+        plt.savefig('MODWT_GPS_NZ/' + station + '_' + direction + '.eps', format='eps')
         plt.close(1)
 
 if __name__ == '__main__':

### 2022-06-29 paper/answer_reviewers.tex
Ready to submit

@@
 \begin{document}
+
+\textbf{Associate Editor}
+
+\bigskip
+
+\textit{Both reviewers expressed that the proposed MODWT method for slow-slip detection shows potential, but were concerned about whether the method could be applied successfully to regions without existing high-quality slow-slip or tremor catalogs.
+An existing catalog appears to be necessary to select the detection threshold from the ROC diagram, and to identify false detections.
+One way to address this concern would be, as Reviewer \#1 suggests, by applying the method to another subduction zone.
+Alternatively, as Reviewer \#2 suggests, the manuscript could be revised to reframe the proposed method.
+Successfully applying the method to another subduction zone would make the strongest case for the MODWT method, although it would be more work.
+Either of these options for revision are acceptable to me, so I leave the choice to the authors.}
+
+\bigskip
+
+We added a section about the analysis of GPS data from New Zealand to show how the method could still be applied if tremor data are not available.
+
+\bigskip
 
@@
 
-We added a section about the analysis of GPS data from New Zealand to show how the method could still be applied if tremor data are not available.
+We added a section about the analysis of GPS data from New Zealand to show how the method could still be applied if tremor data are not available. \\
+
+To select the threshold in the absence of a tremor catalog, a first possibility is to fit the value of the threshold using data from a subduction zone where tremor data are available.
+That is what we did in the manuscript: we use for New Zealand the same threshold as we use for northern Cascadia.
+Another possibility is to try different thresholds, and keep the one that give the most plausible results, with some slow slip events detected but without too much transients that look like noise.
 
@@
 
-\textcolor{red}{I am not sure what to answer to this one.
-Should we try to make a zoom of Figure 9 over a shorter time window around the time of this possible slow slip event and also show the corresponding raw GPS data?}
+We have added a Figure in the Supplement where we highlight the four possible small displacement that we see on the wavelet details.
 
@@
 
-\textcolor{red}{I am not sure about this one. Wouldn't that make the figure even more difficult to understand?}
+To better compare between events detected with GPS data and events detected with tremor data, we added figures where both set events are added on top on the wavelet details.
+We believe the figures are thus more easily readable and interpretable.
 

### 2022-06-29 paper/answer_reviewers_2.tex
Ready to submit

\documentclass[letterpaper, 12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{float}
\usepackage[T1]{fontenc}
\usepackage[lmargin=1 in, rmargin=1 in, tmargin=1 in, bmargin=1 in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{times}
\usepackage{xcolor}

\begin{document}

\textbf{Reviewer 1}

\bigskip

\textit{In "Detection of slow slip events using wavelet analysis of GNSS recordings", the authors propose a methodology for detecting slow slip events using a wavelet decomposition of GNSS time series. In Cascadia, the results of this method are corroborated by examining a wavelet decomposition of tremor time series over the same region and time span. Comparison of results from the application of the methodology using both GNSS and tremor time series shows that slow slip events are detectable at approximately the same time in both data sets for slow slip events down to around Mw 6.0.  As an example of the utility of this approach it is then applied to GNSS time series in New Zealand.} \\

\textit{It is clear from the author's responses to the initial review that this manuscript has been greatly improved. Some questions still remain for me. In particular, how would one choose the wavelet threshold parameters for subduction zones without tremor and where slow slip event amplitudes in GNSS time series are much larger or much smaller than those in Cascadia and New Zealand. It seems that the success seen in utilizing the wavelet parameters derived for Cascadia with the New Zealand time series mostly stems from the similarity in GNSS signal amplitudes (as noted by the authors on lines 604-607). Even under this condition, the exact wavelet detail levels to stack are, in fact, different for Cascadia and New Zealand - a result stemming primarily from the differences in slow slip recurrence intervals between the two subduction zones. How would a researcher determine which wavelet levels to stack unless they already knew when the slow slip events were?}

\bigskip

We think the best way to choose which wavelet details to stack is to look individually at each level detail. The researcher should then look for spatially coherent patterns and a high signal-to-noise ratio. The researcher should think about the expected direction of the slow slip and look for a clear sequence of alternating positive and negative peaks in the direction of the plate motion. The datasets we used have a sampling rate of one observation per day. For a smaller sampling interval, the researcher should look at higher level details. The time between big slow slip events is about 12 months for northern Cascadia and 18 months for New Zealand. For shorter time between events, the researcher should look at lower level wavelet details (5, 6, 7). For longer time between events, the researcher should look at higher level wavelet details (up to 9 or 10). We added a few sentences in the text.

\bigskip

\textit{While this approach is interesting, it does not seem to provide any new insight into the slow slip process nor does it seem to be able to uncover slow slip offsets not otherwise obvious to the naked eye (e.g. buried in the noise). To that end, it provides results no different than previous approaches to recognizing slow slip events. While this may sound like a harsh criticism, in my mind, the main utility of this wavelet approach is to provide a more objective way of identifying and demarcating slow slip events in GNSS time series. For example, it could be used to provide a starting point for identifying slow slip events in a large dataset, thereby saving future investigators time. In the interest of informing the research community of the results of this approach, I would recommend publication of this manuscript.} \\

\textit{Additional questions. Line numbers refer to the manuscript with the bolded changes:} \\

\textit{line 363-365: Wouldn't there be a unit for the wavelet decomposition? If it is the wavelet transform of cumulative tremor count shouldn't the unit still be cumulative tremor count even if a linear trend was removed? In a GNSS time series, the units are mm no matter whether a linear trend has been removed or not.}

\bigskip

The unit would be the fraction of tremor in a day divided by the total number of days. The average value is 1 divided by the total number of days. The threshold we used is about 10 times the average rate. We modified the text.

\bigskip

\textit{line 646-648: slow slip events at the beginning and end of a time series are not readily detectable due to edge effects in the wavelet transform. For each detail level, how much of the time series is eliminated? In other words, assuming one data point per day, at detail level 8, how long is the window of valid data? How about for level 6?}

\bigskip

If we denote $L$ the length of the base wavelet filter used for the wavelet decomposition (in our study, we used a Least Asymmetric wavelet filter of length $L = 8$, see Percival and Walden, 2000, section 4.8, page 107) and $j$ the level of the wavelet detail, the length of the wavelet filter at level $j$ used to compute the wavelet detail $D_j$ is:

\begin{equation*}
L_j = \left( 2^j - 1 \right) \left( L - 1 \right) + 1
\end{equation*}

If we have a time series of length $X_t \left( t = 0, \cdots , N - 1 \right)$, the wavelet coefficients of the detail at level $j$ affected by the boundary conditions at the edges would be the coefficients with indices $t = 0 , \cdots , L_j - 2$ or $t = N - L_j + 1 , \cdots, N - 1$ (see Percival and Walden, 2000, section 5.11, page 199). We get $L_j = 442$ for $j = 6$, $L_j = 890$ for $j = 7$ and $L_j = 1786$ for $j = 8$. In practice, the part of the wavelet details affected by the boundary conditions is much shorter than that. We added a figure in the supplement comparing the wavelet details computed when using only the data between 2008 and 2012 and the wavelet details computed when using the entire time series from 2000 to 2021. Even at level 8 only about 6 months of data on each side are effected by the boundary conditions. We added a paragraph in the text.

\bigskip

\textbf{Reviewer 2}

\bigskip

\textit{I thank the authors for the careful revisions and the brand new application to New Zealand. All of my comments have been adequately addressed, and I think this manuscript should be accepted for publication.} \\

\textit{I have just a few very minor comments, noted below, that I jotted down while rereading the manuscript:
\begin{itemize}
\item L547 backwards -> towards
\item L564-566 "tremor is not as robust" is not very clear; it would be simpler to state that no long-term or high-resolution catalog exists
\item L595 point -> location
\item L602-604 best to cite actual numbers here rather than just "bigger"
\item L667 ween -> seen
\item Figure 2 longitudinal -> east-west
\item Figure 5 units for the tremor threshold
\end{itemize}
}

\bigskip

We modified these sentences in the text.

\bigskip

\textbf{Editor-in-Chief}

\bigskip

\textit{I am a bit concerned about your wording "good correlation", which you use in KP 2, in the abstract, and throughout the text. I understand where it comes from. However, this is a vague statement. What is a "good" correlation? Can you quantify this correlation by a correlation coefficient? Also, from my understanding of your study, you really examine 'temporal correlation", where correlation is perhaps better stated as "agreement". I kindly ask you to review your paper in this context, and consider rephrasing "correlation", or include some precise wording/definition of how want use this term. But recall: a "correlation" should be statistically quantifiable.}

\bigskip

We replaced "correlation" by "agreement" in the text.

\bigskip

\textit{Are these given in full sentences? Do they accurately describe the problem you are addressing, state the main findings and conclusions of your paper? Are they enticing to motivate potential readers to read your paper in detail?}

\bigskip

We modified Key Points 2 and 3.

\bigskip

\textit{Many figures can be simplified and thereby improved visually by removing redundant information/labels/axis; this may also help to reduce their dimensions to make the more suited for printing/publishing. For example:
\begin{itemize}
\item Fig 1: remove redundant x-axis labels in panels 'data' to 'D10'.
\item Fig 2: remove redundant x-axis labels in panels 'data' to 'D10'. Perhaps try to resize panel widths for an optimized single-column printing
\item Fig 3: add latitude and longitude labels
\item Fig 4: remove redundant labels; move x-axis unit 'Time (years)' from top to bottom panel; see also Fig 2
\end{itemize}
}

\bigskip

We modified the figures.

\end{document}

### 2022-06-29 paper/paper_bssa.tex
Ready to submit

@@
 \item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
-\item There is a good correlation between detections of slow slip using GNSS data and using tremor data.
-\item The method could be applied in regions where no tremor is detected in conjunction with slow slip events.
+\item In Northern Cascadia, there is good agreement between detections of slow slip using GNSS data and using tremor data.
+\item The method can be applied as a starting point to identify slow slip without prior information. New Zealand is shown as on example.
 \end{itemize}
@@
 We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal.
-We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia.
-The wavelet-based detection method detects well events of magnitude higher than 6 as determined by independent event catalogs (e.g.
+We verify that there is a good agreement between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia.
+The wavelet-based detection method effectively detects events of magnitude higher than 6 as determined by independent event catalogs (e.g.
 ~\citep{MIC_2019}).
+As a demonstration of using the wavelet analysis in a region without significant tremor, we also analyze GNSS data from New Zealand and detect slow slip events that are spatially and temporally close to those detected previously by other studies.
 
@@
 
-Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}.
+Slow slip events are new phenomena discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}.
 As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate.
@@
 
-In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places ~\citep{HAL_2018}.
+In many places, tectonic tremor is also observed in relation to slow slip, but the spatial agreement between tremor and slow slip may vary along the strike of the plate boundary ~\citep{HAL_2018}.
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
-Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
-In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\citep{OBA_2002,ROG_2003}.
-Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS).
+Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) whose frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
+In subduction zones such as Nankai and Cascadia, tectonic tremor observations agree spatially and temporally with slow slip observations ~\citep{ROG_2003,OBA_2004}.
+Due to this agreement, these paired phenomena have been called Episodic Tremor and Slip (ETS).
 However, this is not always the case.
 For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}.
-In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal correlation between tremor and slow slip occurrence ~\citep{WEC_2016}.
+In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal agreement between tremor and slow slip occurrence ~\citep{WEC_2016}.
 \\
 
-In Cascadia, there are robust signals in both GNSS and tremor ~\citep{HAW_2013}.
+In Cascadia, there are robust signals in both slow slip and tremor ~\citep{HAW_2013}.
 This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS.
@@
 For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded.
-They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8.
+They observed a linear relationship between moment release and number of hours of tremor for slow slip events of moment magnitude 6.3 to 6.8.
 Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data.
@@
 The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains.
-A sharp change in the signal will be localized and seen at all levels of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform.
+A sharp change in the signal will be localized and seen at all time scales of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform.
 \\
@@
 
-Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with ETS events between 1996 and 2011.
-Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events.
+Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with slow slip events between 1996 and 2011.
+Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the slow slip events.
 The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other.
 By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified.
-By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal.
-The authors were then able to clearly see a difference in uplift between the two tidal gauges at Port Angeles and Port Townsend.
+By stacking the components at different time scales of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal.
+Due to the relative location of the tidal gauges at Port Angeles and Port Townsend compared to the slow slip region on the plate boundary, a slow slip event should result in uplift in Port Angeles (western part) and in subsidence in Port Townsend (eastern part).
+Indeed, the authors were able to clearly see a difference in the sign of the uplift at these two tidal gauges.
 \\
@@
 In our study, we use a similar approach to previous studies with a different reasoning.
-We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered.
+We only stack signals at nearby GPS stations, assuming that the east-west displacement due to the slow slip events will then be the same at each of the GPS stations considered.
 We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking.
-Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
+Finally, we assume that the noise and the longitudinal displacement due to the slow slip events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the slow slip events.
 We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia.
-Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
+Our objective is to verify that there is a good agreement between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
 We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods.
+Finally, we apply the method to GNSS data in New Zealand and successfully detect several slow slip events without needing to rely on the tremor data.
 \\
@@
 
-We focused our study on northwest Washington State.
+We first focused our study on northwest Washington State.
 For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University.
-These are network solutions in ITRF2008 with phase ambiguities resolved.
-Solutions are computed with JPL/NASA orbits and satellite clocks.
+These are network solutions in ITRF2014 with phase ambiguities resolved with wide-lane phase-biases.
+Orbits and satellite clocks provided by the Jet Propulsion Laboratory/NASA.
 North, East, and Vertical directions are available.
@@
 
+For the application to slow slip events in New Zealand, we used the GPS time series provided by the Geological hazard information for New Zealand (GeoNet).
+The coordinates have been extracted by GeoNet during the GLOBK run from the combined daily solution files, and converted to (east, north, up) displacement in millimeters with respect to an a priori position and epoch in the ITRF2008 realization.
+The time series provided by GeoNet have no adjustments made to them, so they may, for example, contain offsets due to earthquakes, offsets due to equipment changes at individual sites, and seasonal (annual and semi-annual) signals due to various causes.
+Here again, the direction of the secular interseismic plate motion is close to the West direction, so we only used the East-West component of the GPS time series for the data analysis.
+We detrended the data before applying the wavelet transform by carrying a linear regression of the whole time series and removing the straight line obtained from the regression.
+
 \section{Method}
@@
 The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$.
-If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
+If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ is some integer greater than or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ...
 , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$.
@@
 
+The boundary conditions at the two edges of the time series will affect the wavelet coefficients.
+For the MODWT, if we denote $L$ the length of the base wavelet filter used for the wavelet decomposition (in our study, we used a Least Asymmetric wavelet filter of length $L = 8$, see ~\citep{PER_2000}, section 4.8, page 107), the length of the wavelet filter at level $j$ used to compute the wavelet detail $D_j$ is:
+
+\begin{equation*}
+L_j = \left( 2^j - 1 \right) \left( L - 1 \right) + 1
+\end{equation*}
+
+The wavelet coefficients of the detail al level $j$ affected by the boundary conditions at the edges would then be the coefficients with indices $t = 0 , \cdots , L_j - 2$ or $t = N - L_j + 1 , \cdots, N - 1$ (see ~\citep{PER_2000}, section 5.11, page 199).
+We get $L_j = 442$ for $j = 6$, $L_j = 890$ for $j = 7$ and $L_j = 1786$ for $j = 8$.
+In practice, the part of the wavelet details affected by the boundary conditions is much shorter than that.
+We compared the wavelet details computed when using only the data between 2008 and 2012 and the wavelet details computed when using the entire time series from 2000 to 2021 (Figure S1 in the Supplementary Material).
+Even at level 8 only about 6 months of data on each side are effected by the boundary conditions.
+
 \subsection{Application to synthetic data}
@@
 Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values.
-Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values.
+Figure S2 in the Supplementary Material shows a comparison of the two wavelet details for two different locations of the missing values.
 We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition.
@@
 
-To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for ETS events.
+To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for slow slip events.
 We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3.
@@
 Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform.
-Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5).
+Because of the preprocessing applied to the tremor data before that wavelet transform, the measurement unit associated with the corresponding wavelet details is the fraction of tremor in a day divided by the total number of days.
+The average value is 1 divided by the total number of days.
+Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5).
 Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
 
-\section{Results}
+\section{Application to data from Cascadia}
 
 We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-The result is shown in the top panel of Figure 6, where each line represents one of the locations along strike.
+The result is shown in the top panel of Figure 5, where each line represents one of the locations along strike.
 To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold.
-To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6.
+To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 5.
 We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
@@
 There are three small signals in the GPS data in Winter 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and may be false detections.
-To summarize, we assume that true detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series.
-Then, all the 13 events present on the 8th level detail of the wavelet decomposition are true detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections.
+To summarize, we assume that robust detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series.
+Then, all the 13 events present on the 8th level detail of the wavelet decomposition are robust detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are robust detections.
 \\
 
-\section{Discussion}
-
-To better evaluate the number of true and false detections, we convert the wavelet details into binary time series.
+To better evaluate the number of robust and false detections, we convert the wavelet details into trinary time series.
 If the absolute value of the wavelet detail is higher that a threshold, we replace the value by 1 (for positive values) or -1 (for negative values), otherwise we replace the value by 0.
 We do this on both the wavelet details of the GPS data and of the tremor data.
-Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a true detection (true positive, TP).
+Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a robust detection (true positive, TP).
 If the GPS and the tremor time series have opposite signs, or if the absolute value of the GPS time series is 1 but the value of the tremor time series is 0, we have a false detection (false positive, FP).
@@
 The ROC curve is widely use for binary classification problems in statistics and machine learning.
-We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
+We calculate an ROC value by varying the values of the threshold (here the two thresholds used to convert the GPS and the tremor time series into trinary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate.
 If the classification was made randomly, all the points would fall on the first diagonal.
@@
 
-As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a binary time series.
+As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a trinary time series.
 We apply this transform to both the GPS and the tremor time series with varying thresholds.
-We then plot the ROC curve on Figure 9, each dot representing a different threshold.
-The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10.
+We then plot the ROC curve on Figure 8, each dot representing a different threshold.
+The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 9.
 We can see that there is a trade-off between sensitivity and specificity as we vary the threshold.
@@
 If we increase the number of true events detected, we also increase the false positive rate.
-In Figure 10, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75, and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
+If we increase the threshold for the tremor, the curve goes farther away from the first diagonal, that is we get better classification results.
+If we increase the threshold for the GPS, the false positive rate and the the number of events detected decrease.
+In Figure 9, we have chosen thresholds for the GPS time series and the tremor time series such that the specificity is higher than 0.75 (that is the false positive rate is lower than 0.25), and the sensitivity is the highest possible, that is we have chosen the thresholds corresponding to the dot that is farthest from the diagonal, which is random.
 \\
@@
 Table 2 summarizes for each event its timing, its number and its magnitude as indicated in the catalog from ~\citet{MIC_2019}, and whether it is part of a bigger magnitude 6 event.
-Figure 11 shows the 5th level detail wavelet decomposition of the GPS data.
-Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
+Figure 10 shows the 5th level detail wavelet decomposition of the GPS data.
+Red lines show the timing of the big slow slip events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.
 \\
@@
 The one small event that is not part of a bigger event (Winter 2009) is visible at level 5 of the wavelet decomposition.
-However, some other events that are not in the catalog of ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and late 2016.
-Therefore, it is difficult to differentiate between a true detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
+However, some other events that are not in the catalog of ~\citet{MIC_2019}'s catalog are also visible in late 2007, early 2010, early 2012, and early 2020.
+Therefore, it is difficult to differentiate between a robust detection and a false detection, and to conclude whether the method can indeed detect events of magnitude 5.
 \\
@@
 In Figure 9, we see four smaller events that are not in the catalog of ~\citet{MIC_2019}: at about 2007.5, there is a negative peak followed by a positive peak (that is an event in the opposite direction of what would be expected from slow slip), at about 2010.2, 2012.2 and 2020.2, there are positive peaks followed by negative peaks for all the sixteen locations studied in this paper.
+These events are highlighted in Figure S4 in the Supplementary Information.
 Looking back at the original GPS data, there is a small increase in the displacement in the eastern direction that lasts about one or two months at about 2007.5.
@@
 Finally, there is also a small decrease in displacement at about 2020.2 that is difficult to interpret.
+\\
+
+Due to the short distances between the GPS stations and the locations of the red triangles on the map from Figure 3, the same station could be used multiple times for the stacking at different locations.
+When considering two different locations, the stacking is thus made over an overlapping number of stations.
+Table 3 summarizes the number of stations and the number of overlapping stations for each location on Figure 3.
+We hypothesize that the small displacement in the eastern direction seen at about 2007.5 could be due to a misbehaving station common to several locations.
+However, several GPS stations indeed show an increase in the displacement in the eastern direction at about 2007.5.
+There are many missing data around that time, so it is difficult to conclude.
+\\
+
+Another possibility is that common mode signals could stack constructively across GNSS stations and produce peaks in the wavelet details that are actually due to non-tectonic signals.
+We computed common mode signals for different latitude bins (each bin has width equal to half-a-degree of latitude) following the same method as \citet{NUY_2021}.
+We first stacked all the time series for the stations in each latitude bin that are located more than 100 km east of the 40 km depth contour of the plate boundary.
+We assume that these stations are not sensitive to the deformation on the plate interface.
+We then apply a yearly moving average to each common mode signal in order to remove any leftover noise.
+The common mode signal was then removed from the GNSS time series depending on each site’s latitude.
+Figure S3 in the Supplementary Information shows the corresponding sum of the stacks of the 6th, 7th and 8th wavelet details obtained from the resulting time series.
+The common modes seem to have little impact on the results and do not explain the additional four small events that we noted in Figure 9.
+\\
+
+In order to convert our filtered eastward displacement time series into a slow slip event catalog we note that red bars represent displacements exceeding a threshold of 0.8 mm (east), and blue marks displacements less than minus -0.8 mm (west).
+During times with no slow slip GPS stations on the overriding plate are pushed slowly eastward by the locked subducting plate.
+Slow slip events represent GPS motion towards the west.
+Thus, we infer that slow slip events happen when red bars are immediately followed by blue bars in the wavelet details.
+We have identified everywhere that this has happened and mark it with a green line in Figure 11 and as a row in Table 4.
+We find 17 possible SSEs by this method using filtered GPS data only.
+For each of these 17 events we determine the time difference between the mid time of the GPS catalog and the nearest time from the tremor catalog (Table 1).
+These time differences are in column 6 (Table 4).
+Every event in the GPS catalog has a match in the tremor catalog except for the tremor event at 2010.15.
+There is also only one event in the tremor catalog that is not in the GPS catalog.
+It occurs at 2014.65 with a duration of 15 days and 190 hours of tremor.
+It occurs 0.25 years after the nearest GPS event.
+There are also two marginal events in the tremor catalog with time differences of 0.13 and 0.10 years, but those are also among the smaller events with 162 and 193 hours of tremor.
+
+\section{Application to data from New Zealand}
+
+We now apply our wavelet-based method to detect slow slip events in New Zealand, a location where the spatial and temporal agreement between tremor and slow slip is not as good as in other subduction zones.
+The tectonics of the North Island of New Zealand are dominated by the westward subduction of the Pacific Plate under the Australian Plate at the Hikurangi Trench.
+Two types of slow slip events have been observed at the Hikurangi margin.
+Shallow (10-15 km depth), shorter (1-3 weeks), and usually smaller (Mw 6.3-6.8) slow slip events have been observed every 18-24 months in the northern part of the margin.
+Deeper (35-60 km depth), longer (12-18 months), and larger (Mw 7.0) slow slip events have been observed every 5 years in the southern part of the margin ~\citep{WAL_2010,TOD_2016}.
+The detection of tremor has been elusive in northern Hikurangi.
+~\citet{DEL_2009} observed an increase in the rate of microseismicity downdip of the 2004 Gisborne slow slip event.
+More recently, however, ~\citep{KIM_2011} detected a low level of tremor activity that increased during the 2010 Gisborne slow slip event.
+As was the case for the microearthquakes, the source of the tremor was located downdip of the slow slip patch determined from GNSS data.
+~\citep{IDE_2012} detected tremor downdip of the location of two deep slow slip events observed by ~\citet{WAL_2013} in 2006 and 2008.
+However, contrary to ETS events in Cascadia and Nankai, the tremor activity did not seem to increase during the slow slip events.
+~\citet{TOD_2016}  detected tremor associated with most of the shallow slow slip events between 2010 and 2015, and located downdip of the geodetically inferred slip area.
+They also detected deeper tremor between 20 and 50 km depth with unclear origin.
+They hypothesized that these tremor may be related to undetected deep long-term slow slip events.
+\\
+
+To evaluate whether the wavelet analysis is effective in a region without robust tremor, we take all the New Zealand GPS stations located in a 50 km radius of a given location, for the 18 locations indicated in Figure 12, and we stack the 6th level details, the 7th level details or the 8th level details over all the GPS stations.
+We then sum together the 6th, 7th and 8th levels stacked wavelet details (Figure 13, top panel).
+We highlight positive and negative peaks with red and blue colors as was done in Figure 9.
+We cannot use the tremor data to decide what is the appropriate threshold above which we consider that there is a slow slip event.
+Slow slip events in New Zealand result in surface displacements that are similar in amplitude to twice as large as those observed in Cascadia.
+Therefore, the amplitudes of the peaks in the wavelet details should be similar in New Zealand and in Cascadia and we choose identical thresholds for both regions.
+As a slow slip event in northern New Zealand results in a displacement in the east direction at the earth’s surface, the slow slip events are indicated by a negative peak followed by a positive peak in the stacked wavelet details.
+We compare the results of the timings and locations of the slow slip events to those events detected by ~\citet{TOD_2016}.
+As they only used data from five GPS stations (PUKE, ANAU, GISB, MAHI and CKID), we indicate by a vertical orange bar on the bottom panel of Figure 13 each time a slow slip event was detected for these stations.
+The orange bars are centered on the latitudes of the GPS stations.
+If a slow slip event was detected by more than one station, all the corresponding orange bars are linked together to show the spatial extent of the slow slip.
+~\citet{TOD_2016} indicated by a question mark (on their Figure 2 and their Table 1) additional possible events, and those are indicated by a dotted orange bar on Figure 13.
+To compare with the slow slip events detected with the wavelet method, we also mark by a green bar every time a negative peak lower than the threshold is followed by a positive peak higher than the threshold.
+Table 5 summarizes the slow slip events detected with the wavelet method for 2010-2016.
+\\
+
+We observe that there is a good agreement between the events detected with the wavelet method and the events previously detected by ~\citet{TOD_2016}.
+We clearly see an event propagating from south to north in January-February (event 2 from ~\citet{TOD_2016}), an event in March-April 2010 (event 3), an event in April-May 2011 in the northern part of the region studied (events 6 and 7), an event propagating south-to-north in August-September and September-October 2011 (events 8 and 9), and an event in December 2011 (event 10).
+Although ~\citet{TOD_2016} only detected this last event for GPS station GISB, it seems that this event may have also extended farther to the north and the south.
+We then clearly see an event in the northern part of the region studied in August 2012 (event 12), an event in December 2012-January 2013 (event 13), an event in the southern part of the region studied in February-March 2013 (event 14), an event propagating from south to north in June-July and July-August 2013 (events 15 and 16), an event in September 2014 (events 20 and 21), an event in the southern part of the region studied in December 2014-January 2015 (events 22 and 23), and an event in June-July 2015 in the northern part of the region studied (event 26).
+It is unclear if the event near station ANAU in early 2010 (event 1) is visible in the wavelet details as it is too close to the beginning of the time series.
+The June-July 2010 event (event 4), the August 2010 event (event 5), and the March 2012 event (event 11), are not clearly visible in the wavelet details.
+The events in September-October 2013 (event 17), December 2013 (event 18), May-June 2014 (event 19), January-February (event 24) and February 2015 (events 25) are not clearly seen in the wavelet details, but there could be a small negative peak followed by a small positive peak at these times.
+Additionally, there could be two other events that are not in \citep{TOD_2016} in Fall 2010 (southern part of the region studied) and in Fall 2015.
+\\
+
+Our wavelet-based method thus works well to detect transients in GPS data that could be slow slip events, even in the absence of tremor data.
+The choice of the appropriate threshold to decide that there is a transient and the levels of the wavelet details that we look at for the detection may still not be easily made.
+There is a difference between Cascadia and New Zealand in terms of which wavelet details to stack.
+In particular, as there is more time between two slow slip events in New Zealand than in Cascadia, the biggest slow slip events (early 2010, late 2011, 2013 and late 2014) can also be seen on the 9th level detail for New Zealand, whereas they could not be seen for Cascadia.
+We then use the method to detect slow slip events during the period 2016-2022, which was not covered by ~\citet{TOD_2016} (Figure 14).
+We note four large transients that could be slow slip events in late 2016, late 2017, early 2019 and mid-2021.
+There are also possible smaller events in the northern part of the area in mid-2018 and in most of the area studied in early 2020.
+Table 6 summarizes the slow slip events detected with the wavelet method for 2016-2022.
+\\
+
+The method is thus applicable in regions where tremor data are not usable.
+To determine which wavelet levels to stack, we recommend analyzing each level detail.
+Look for spatially coherent patterns, wavelet details with energy at similar times and high signal-to-noise ratios.
+Look for alternating positive and negative peaks that are consistent with the expected direction of slow slip.
+Consider wavelet details with time scales ranging from the expected duration of slow slip events to the expected recurrence times between slow slip events.
+For Cascadia and New Zealand this would be weeks to years.
+Determination of a threshold is subjective.
+At large thresholds the large slow slip events should be clear.
+At smaller thresholds there is the possibility of identifying smaller events, but at the risk of false detections.
 
@@
 In this paper, we develop and test a new approach for detecting transient events in GPS time series, such as slow slip events.
-We used wavelet methods to analyze GNSS time series and tremor recordings of slow slip events in Cascadia.
+We used wavelet methods to analyze GNSS time series and tremor recordings of slow slip events in Cascadia, and GNSS time series in New Zealand.
 We used detrended GNSS data, applied the MODWT transform, and stacked the wavelet details over several nearby GNSS stations.
@@
 We assumed that there is a slow slip event whenever the wavelet signal gets above a threshold.
-We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
+We verified that there is a good agreement between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
 The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
 ~\citep{MIC_2019}).
-We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between true detections and false detections.
+We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between robust detections and false detections.
+We then applied the method to GNSS data in New Zealand and detected slow slip events consistent with the events previously detected by ~\citet{TOD_2016}.
 
@@
 
-The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA} \url{http://www.panga.cwu.edu/}.
+The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA} \url{http://www.panga.cwu.edu/} and the Geonet website \url{https://www.geonet.org.nz/}.
 The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/slowslip}.
-Figure 4 was created using GMT ~\citep{WES_1991}.
+Figures 3 and 12 were created using GMT ~\citep{WES_1991}.
 
@@
 
+The authors would like to thank two anonymous reviewers and the Associate Editor Jeanne Hardebeck, whose comments greatly helped improve the manuscript.
 This work was funded by the grant from the National Science Foundation EAR-1358512.
@@
 \begin{table}[hbt!]
-\caption{Episodic Tremor and Slip events with M > 6 identified by MODWT in both the GPS and the tremor data.
+\caption{Episodic Tremor and Slip events with M $>$ 6 identified by MODWT in both the GPS and the tremor data.
 The duration and the number of tremor are from the tremor catalog of the PNSN.
@@
  \begin{table}[hbt!]
- \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
+ \caption{Magnitude 5 to 6 events from ~\citet{MIC_2019}.}
  \centering
@@
 
-\newpage
+ \begin{table}[hbt!]
+ \caption{Number of GPS stations used for the stacking for each location on Figure 3 and number of common stations with the location immediately to the north and the location immediately to the south.}
+ \centering
+ \begin{tabular}{c c c c c}
+ \hline
+ Index & Latitude & Number of stations & Common stations (north) & Common stations (south) \\
+ \hline
+0 & 47.2 & 15 & 14 & \\
+1 & 47.3 & 18 & 17 & 14 \\
+2 & 47.4 & 24 & 20 & 17 \\
+3 & 47.5 & 21 & 20 & 20 \\
+4 & 47.6 & 22 & 14 & 20 \\
+5 & 47.7 & 17 & 12 & 14 \\
+6 & 47.8 & 13 & 8 & 12 \\
+7 & 47.9 & 10 & 9 & 8 \\
+8 & 48.0 & 10 & 7 & 9 \\
+9 & 48.1 & 8 & 7 & 7 \\
+10 & 48.2 & 10 & 8 & 7 \\
+11 & 48.3 & 9 & 9 & 8 \\
+12 & 48.4 & 9 & 5 & 9 \\
+13 & 48.5 & 7 & 5 & 5 \\
+14 & 48.6 & 6 & 5 & 5 \\
+15 & 48.7 & 5 & & 5 \\
+ \hline
+ \end{tabular}
+ \end{table}
+
+ \begin{table}[hbt!]
+ \caption{Cascadia catalog of slow slip events based only on MODWT analysis of GPS time series and inferring that the transition of red followed immediate by blue marks a slow slip event.
+First four columns are the start and end times and start and end latitudes of the green bars in Figure 11.
+The fifth column is 1 for robust detection and 2 if not as robust.
+Column 6 is the time difference in years between the mid times of the GPS catalog and the nearest mid times of the tremor catalog summarized in Table 1.}
+ \centering
+ \begin{tabular}{c c c c c c}
+ \hline
+start time & end time & start latitude & end latitude & &dT tremor catalog \\
+ \hline
+2007.06 & 2007.10 & 47.16 & 48.72 & 1 & 0.02 \\
+2008.30 & 2008.40 & 47.35 & 48.73 & 1 & 0.01 \\
+2009.35 & 2009.44 & 47.92 & 48.73 & 1 & 0.05 \\
+2010.12 & 2010.15 & 47.32 & 48.73 & 1 & 0.50 no match \\
+2010.61 & 2010.64 & 47.17 & 48.72 & 1 & 0.00 \\
+2011.57 & 2011.61 & 47.18 & 48.68 & 1 & 0.01 \\
+2012.65 & 2012.65 & 48.74 & 47.76 & 1 & 0.05 \\
+2013.71 & 2013.75 & 47.47 & 48.73 & 1 & 0.02 \\
+2014.89 & 2014.90 & 48.73 & 47.79 & 1 & 0.01 \\
+2015.98 & 2016.09 & 48.73 & 47.20 & 1 & 0.08 \\
+2017.17 & 2017.24 & 47.38 & 48.72 & 1 & 0.02 \\
+2018.35 & 2018.36 & 47.48 & 47.93 & 1 & 0.13 part of same event? \\
+2018.48 & 2018.50 & 48.72 & 48.09 & 1 & 0.00 \\
+2019.32 & 2019.34 & 47.17 & 47.72 & 2 & 0.10 \\
+2019.90 & 2019.91 & 48.47 & 48.72 & 2 & 0.02 \\
+2020.79 & 2020.83 & 47.18 & 48.13 & 1 & 0.02 \& 0.05 \\
+2021.11 & 2021.12 & 48.75 & 48.48 & 2 & 0.02 \\
+\hline
+ \end{tabular}
+ \end{table}
+
+ \begin{table}[hbt!]
+ \caption{New Zealand catalog of slow slip events for 2010-2016 based only on MODWT analysis of GPS time series and inferring that the transition of red followed immediate by blue marks a slow slip event.
+First four columns are the start and end times and start and end latitudes of the green bars in Figure 13.
+The fifth column is 1 for robust detection and 2 if not as robust.}
+ \centering
+ \begin{tabular}{c c c c c}
+ \hline
+start time & end time & start latitude & end latitude & \\
+ \hline
+2010.05 & 2010.07 & -39.67 & -39.12 & 1 \\
+2010.19 & 2010.22 & -39.12 & -38.07 & 1 \\
+2010.75 & 2010.76 & -39.73 & -39.41 & 1 \\
+2011.36 & 2011.37 & -38.22 & -38.02 & 2 \\
+2011.71 & 2011.74 & -37.97 & -38.41 & 1 \\
+2011.67 & 2011.71 & -39.73 & -38.91 & 1 \\
+2011.92 & 2011.95 & -38.84 & -38.16 & 1 \\
+2012.63 & 2012.63 & -39.42 & -39.62 & 2 \\
+2012.64 & 2012.66 & -38.53 & -38.02 & 1 \\
+2012.95 & 2012.96 & -38.32 & -37.98 & 1 \\
+2013.15 & 2013.16 & -38.87 & -39.72 & 1 \\
+2013.55 & 2013.57 & -38.62 & -38.01 & 1 \\
+2013.74 & 2013.74 & -38.77 & -38.97 & 2 \\
+2013.92 & 2013.93 & -38.17 & -37.98 & 2 \\
+2013.91 & 2013.95 & -39.37 & -39.73 & 1 \\
+2014.78 & 2014.79 & -38.03 & -39.03 & 1 \\
+2014.96 & 2015.00 & -39.07 & -39.72 & 1 \\
+2015.53 & 2015.53 & -39.42 & -39.72 & 1 \\ 
+2015.52 & 2015.55 & -37.97 & -38.43 & 1 \\
+2015.78 & 2015.79 & -38.77 & -39.37 & 1 \\
+\hline
+ \end{tabular}
+ \end{table}
+
+ \begin{table}[hbt!]
+ \caption{New Zealand catalog of slow slip events for 2016-2022 based only on MODWT analysis of GPS time series and inferring that the transition of red followed immediate by blue marks a slow slip event.
+First four columns are the start and end times and start and end latitudes of the green bars in Figure 13.
+The fifth column is 1 for robust detection and 2 if not as robust.}
+ \centering
+ \begin{tabular}{c c c c c}
+ \hline
+start time & end time & start latitude & end latitude & \\
+ \hline
+2016.84 & 2016.90 & -37.96 & -39.72 & 1 \\
+2017.10 & 2017.10 & -38.78 & -39.00 & 2 \\
+2017.73 & 2017.78 & -37.98 & -38.51 & 1 \\
+2018.04 & 2018.06 & -38.58 & -39.07 & 1 \\
+2018.63 & 2018.64 & -38.27 & -37.97 & 2 \\
+2019.26 & 2019.33 & -37.97 & -39.73 & 1 \\
+2020.09 & 2020.12 & -37.97 & -38.23 & 2 \\
+2020.34 & 2020.35 & -37.96 & -39.72 & 1 \\
+2020.33 & 2020.33 & -37.96 & -38.10 & 2 \\
+2020.32 & 2020.32 & -38.62 & -38.79 & 2 \\
+2020.36 & 2020.37 & -39.70 & -39.35 & 2 \\
+2021.11 & 2021.11 & -39.51 & -39.64 & 2 \\
+2021.39 & 2021.47 & -39.72 & -38.08 & 1 \\
+ \hline
+ \end{tabular}
+ \end{table}
+
+\clearpage
 
@@
 \item Figure 2.
-Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).
+Top left: East-west displacement recorded at GPS station PGC5.
+The resulting details and smooth of the wavelet decomposition are shown in increasing level from top to bottom and from left to right.
 
-\item Figure 3. Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.
-
-\item Figure 4.
+\item Figure 3.
 GPS stations used in this study (black triangles).
@@
 The black lines represent the timings of the ETS events from Table 1.
-We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 (for the GPS) or 0.003 (for the tremor).
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 mm (for the GPS) or 0.003 (for the tremor, that is about 17 times the average value of the signal).
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 7. Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).
+\item Figure 6.
+Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 7th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.5 mm (for the GPS) or 0.01 (for the tremor, that is about 56 times the average value of the signal).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 8. Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).
+\item Figure 7.
+Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 6th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.3 mm (for the GPS) or 0.009 (for the tremor, that is about 51 times the average value of the signal).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 9. Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
+\item Figure 8.
+ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
+The black cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
+The values of the threshold are color-coded.
+Reds (bottom curve) correspond to the lowest value of the threshold for the tremor (0.001), while oranges, greens, blues, purples correspond to increasing values of the threshold for the tremor (up to 0.01, top curve).
+The brightest colors (bottom left) correspond to the highest values of the threshold for the GPS (1.5 mm), while the darker colors (top right) correspond to decreasing values of the threshold for the GPS (0.1 mm).
+
+\item Figure 9.
+Top: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: Sum of the 6th, 7th and 8th levels detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.8 mm (for the GPS) or 0.01 (for the tremor, that is about 56 times the average value of the signal).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
 \item Figure 10.
-ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
-Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
-
-\item Figure 11.
 Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
@@
 The blue lines represent the timings of the magnitude 5 events from the catalog of ~\citet{MIC_2019}.
+
+\item Figure 11.
+Same as top panel of Figure 9: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+We mark with a green bar the slow slip events from Table 4 detected with the wavelet method.
+Full lines correspond to robust detections (1 in Table 4) and dotted lines to less robust detections (2 in Table 4).
+
+\item Figure 12.
+GPS stations used for the slow slip detection in New Zealand (black triangles).
+The red triangles are the locations where we stack the GPS data.
+They are located close to the 20 km depth contour of the plate boundary from ~\citet{WIL_2013}.
+
+\item Figure 13.
+Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
+The time period covered is 2010-2016.
+We mark by a red rectangle every time where the amplitude is higher than a threshold equal to 0.8 mm.
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
+Bottom: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition.
+We mark with an orange bar the slow slip events detected by ~\citet{TOD_2016} and with a green bar the slow slip events from Table 5 detected with the wavelet method.
+Full lines correspond to robust detections (1 in Table 5) and dotted lines to less robust detections (2 in Table 5).
+
+\item Figure 14.
+Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
+The time period covered in 2016-2022.
+We mark by a red rectangle every time where the amplitude is higher than a threshold equal to 0.8 mm.
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
+We mark with a green bar the slow slip events from Table 6 detected with the wavelet method.
+Full lines correspond to robust detections (1 in Table 6) and dotted lines to less robust detections (2 in Table 6).
 
@@
 \begin{figure}
-\noindent\includegraphics[width=10cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
+\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
 \caption{Demonstration of a wavelet decomposition for a synthetic dataset.
@@
 \begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
-\caption{Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
+\caption{Top left: East-west displacement recorded at GPS station PGC5.
+The resulting details and smooth of the wavelet decomposition are shown in increasing level from top to bottom and from left to right.}
 \label{pngfiguresample}
@@
 \begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
 \caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).}
@@
 The black lines represent the timings of the ETS events from Table 1.
-We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 (for the GPS) or 0.003 (for the tremor).
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 mm (for the GPS) or 0.003 (for the tremor, that is about 17 times the average value of the signal).
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
-\caption{Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).}
+\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 7th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.5 mm (for the GPS) or 0.01 (for the tremor, that is about 56 times the average value of the signal).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.eps}
-\caption{Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6-7-8.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
-\caption{Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).}
+\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 6th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.3 mm (for the GPS) or 0.009 (for the tremor, that is about 51 times the average value of the signal).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}
@@
 \caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
-Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.}
+Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
+The black cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
+The values of the threshold are color-coded.
+Reds (bottom curve) correspond to the lowest value of the threshold for the tremor (0.001), while oranges, greens, blues, purples correspond to increasing values of the threshold for the tremor (up to 0.01, top curve).
+The brightest colors (bottom left) correspond to the highest values of the threshold for the GPS (1.5 mm), while the darker colors (top right) correspond to decreasing values of the threshold for the GPS (0.1 mm).}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
+\caption{Top: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: Sum of the 6th, 7th and 8th levels detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.8 mm (for the GPS) or 0.01 (for the tremor, that is about 56 times the average value of the signal).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}
@@
 
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8_slowslip.eps}
+\caption{Same as top panel of Figure 9: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+We mark with a green bar the slow slip events from Table 4 detected with the wavelet method.
+Full lines correspond to robust detections (1 in Table 4) and dotted lines to less robust detections (2 in Table 4).}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations_NZ.eps}
+\caption{GPS stations used for the slow slip detection in New Zealand (black triangles).
+The red triangles are the locations where we stack the GPS data.
+They are located close to the 20 km depth contour of the plate boundary from ~\citet{WIL_2013}.}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_events.eps}
+\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
+The time period covered is 2010-2016.
+We mark by a red rectangle every time where the amplitude is higher than a threshold equal to 0.8 mm.
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
+Bottom: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition.
+We mark with an orange bar the slow slip events detected by ~\citet{TOD_2016} and with a green bar the slow slip events from Table 5 detected with the wavelet method.
+Full lines correspond to robust detections (1 in Table 5) and dotted lines to less robust detections (2 in Table 5).}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_2016-2022.eps}
+\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
+The time period covered in 2016-2022.
+We mark by a red rectangle every time where the amplitude is higher than a threshold equal to 0.8 mm.
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
+We mark with a green bar the slow slip events from Table 6 detected with the wavelet method.
+Full lines correspond to robust detections (1 in Table 6) and dotted lines to less robust detections (2 in Table 6).}
+\label{pngfiguresample}
+\end{figure}
+
 \end{document}

### 2022-06-29 paper/paper_bssa_annotated.tex
Ready to submit

--- 2022-03-10 paper/paper_bssa.tex
@@
 \item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
-\item There is a good correlation between detections of slow slip using GNSS data and using tremor data.
-\item The method could be applied in regions where no tremor is detected in conjunction with slow slip events.
+\item In Northern Cascadia, there is good \textbf{agreement} between detections of slow slip using GNSS data and using tremor data.
+\item The method \textbf{can} be applied \textbf{as a starting point to identify slow slip without prior information. New Zealand is shown as on example.}
 \end{itemize}
@@
 We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal.
-We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia.
-The wavelet-based detection method detects well events of magnitude higher than 6 as determined by independent event catalogs (e.g.
+We verify that there is a good \textbf{agreement} between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia.
+The wavelet-based detection method effectively detects events of magnitude higher than 6 as determined by independent event catalogs (e.g.
 ~\citep{MIC_2019}).
+As a demonstration of using the wavelet analysis in a region without significant tremor, we also analyze GNSS data from New Zealand and detect slow slip events that are spatially and temporally close to those detected previously by other studies.
 
@@
 
-In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places ~\citep{HAL_2018}.
+In many places, tectonic tremor is also observed in relation to slow slip, but the spatial \textbf{agreement} between tremor and slow slip may vary along the strike of the plate boundary ~\citep{HAL_2018}.
 Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases.
-Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
-In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\citep{OBA_2002,ROG_2003}.
-Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS).
+Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) whose frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz).
+In subduction zones such as Nankai and Cascadia, tectonic tremor observations \textbf{agree} spatially and temporally with slow slip observations ~\citep{ROG_2003,OBA_2004}.
+Due to this \textbf{agreement}, these paired phenomena have been called Episodic Tremor and Slip (ETS).
 However, this is not always the case.
 For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}.
-In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal correlation between tremor and slow slip occurrence ~\citep{WEC_2016}.
+In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal \textbf{agreement} between tremor and slow slip occurrence ~\citep{WEC_2016}.
 \\
 
-In Cascadia, there are robust signals in both GNSS and tremor ~\citep{HAW_2013}.
+In Cascadia, there are robust signals in both slow slip and tremor ~\citep{HAW_2013}.
 This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS.
@@
 In our study, we use a similar approach to previous studies with a different reasoning.
-We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered.
+We only stack signals at nearby GPS stations, assuming that the east-west displacement due to the slow slip events will then be the same at each of the GPS stations considered.
 We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking.
-Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events.
+Finally, we assume that the noise and the longitudinal displacement due to the slow slip events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the slow slip events.
 We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia.
-Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
+Our objective is to verify that there is a good \textbf{agreement} between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
 We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods.
+Finally, we apply the method to GNSS data in New Zealand and successfully detect several slow slip events without needing to rely on the tremor data.
 \\
@@
 
+\textbf{The boundary conditions at the two edges of the time series will affect the wavelet coefficients.
+For the MODWT, if we denote $L$ the length of the base wavelet filter used for the wavelet decomposition (in our study, we used a Least Asymmetric wavelet filter of length $L = 8$, see ~\citep{PER_2000}, section 4.8, page 107), the length of the wavelet filter at level $j$ used to compute the wavelet detail $D_j$ is:}
+
+\begin{equation*}
+L_j = \left( 2^j - 1 \right) \left( L - 1 \right) + 1
+\end{equation*}
+
+\textbf{The wavelet coefficients of the detail al level $j$ affected by the boundary conditions at the edges would then be the coefficients with indices $t = 0 , \cdots , L_j - 2$ or $t = N - L_j + 1 , \cdots, N - 1$ (see ~\citep{PER_2000}, section 5.11, page 199).
+We get $L_j = 442$ for $j = 6$, $L_j = 890$ for $j = 7$ and $L_j = 1786$ for $j = 8$.
+In practice, the part of the wavelet details affected by the boundary conditions is much shorter than that.
+We compared the wavelet details computed when using only the data between 2008 and 2012 and the wavelet details computed when using the entire time series from 2000 to 2021 (Figure S1 in the Supplementary Material).
+Even at level 8 only about 6 months of data on each side are effected by the boundary conditions.}
+
 \subsection{Application to synthetic data}
@@
 Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform.
-Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5).
+Because of the preprocessing applied to the tremor data before that wavelet transform, \textbf{the measurement unit associated with the corresponding wavelet details is the fraction of tremor in a day divided by the total number of days.
+The average value is 1 divided by the total number of days.} Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5).
 Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.
 
-\section{Results}
+\section{Application to data from Cascadia}
 
 We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.
-The result is shown in the top panel of Figure 6, where each line represents one of the locations along strike.
+The result is shown in the top panel of Figure 5, where each line represents one of the locations along strike.
 To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold.
-To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6.
+To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 5.
 We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks.
@@
 Finally, there is also a small decrease in displacement at about 2020.2 that is difficult to interpret.
+\\
+
+Due to the short distances between the GPS stations and the locations of the red triangles on the map from Figure 3, the same station could be used multiple times for the stacking at different locations.
+When considering two different locations, the stacking is thus made over an overlapping number of stations.
+Table 3 summarizes the number of stations and the number of overlapping stations for each location on Figure 3.
+We hypothesize that the small displacement in the eastern direction seen at about 2007.5 could be due to a misbehaving station common to several locations.
+However, several GPS stations indeed show an increase in the displacement in the eastern direction at about 2007.5.
+There are many missing data around that time, so it is difficult to conclude.
+\\
+
+Another possibility is that common mode signals could stack constructively across GNSS stations and produce peaks in the wavelet details that are actually due to non-tectonic signals.
+We computed common mode signals for different latitude bins (each bin has width equal to half-a-degree of latitude) following the same method as \citet{NUY_2021}.
+We first stacked all the time series for the stations in each latitude bin that are located more than 100 km east of the 40 km depth contour of the plate boundary.
+We assume that these stations are not sensitive to the deformation on the plate interface.
+We then apply a yearly moving average to each common mode signal in order to remove any leftover noise.
+The common mode signal was then removed from the GNSS time series depending on each site’s latitude.
+Figure S3 in the Supplementary Information shows the corresponding sum of the stacks of the 6th, 7th and 8th wavelet details obtained from the resulting time series.
+The common modes seem to have little impact on the results and do not explain the additional four small events that we noted in Figure 9.
+\\
+
+In order to convert our filtered eastward displacement time series into a slow slip event catalog we note that red bars represent displacements exceeding a threshold of 0.8 mm (east), and blue marks displacements less than minus -0.8 mm (west).
+During times with no slow slip GPS stations on the overriding plate are pushed slowly eastward by the locked subducting plate.
+Slow slip events represent GPS motion \textbf{towards} the west.
+Thus, we infer that slow slip events happen when red bars are immediately followed by blue bars in the wavelet details.
+We have identified everywhere that this has happened and mark it with a green line in Figure 11 and as a row in Table 4.
+We find 17 possible SSEs by this method using filtered GPS data only.
+For each of these 17 events we determine the time difference between the mid time of the GPS catalog and the nearest time from the tremor catalog (Table 1).
+These time differences are in column 6 (Table 4).
+Every event in the GPS catalog has a match in the tremor catalog except for the tremor event at 2010.15.
+There is also only one event in the tremor catalog that is not in the GPS catalog.
+It occurs at 2014.65 with a duration of 15 days and 190 hours of tremor.
+It occurs 0.25 years after the nearest GPS event.
+There are also two marginal events in the tremor catalog with time differences of 0.13 and 0.10 years, but those are also among the smaller events with 162 and 193 hours of tremor.
+
+\section{Application to data from New Zealand}
+
+We now apply our wavelet-based method to detect slow slip events in New Zealand, a location where \textbf{the spatial and temporal agreement between tremor and slow slip is not as good as in} other subduction zones.
+The tectonics of the North Island of New Zealand are dominated by the westward subduction of the Pacific Plate under the Australian Plate at the Hikurangi Trench.
+Two types of slow slip events have been observed at the Hikurangi margin.
+Shallow (10-15 km depth), shorter (1-3 weeks), and usually smaller (Mw 6.3-6.8) slow slip events have been observed every 18-24 months in the northern part of the margin.
+Deeper (35-60 km depth), longer (12-18 months), and larger (Mw 7.0) slow slip events have been observed every 5 years in the southern part of the margin ~\citep{WAL_2010,TOD_2016}.
+The detection of tremor has been elusive in northern Hikurangi.
+~\citet{DEL_2009} observed an increase in the rate of microseismicity downdip of the 2004 Gisborne slow slip event.
+More recently, however, ~\citep{KIM_2011} detected a low level of tremor activity that increased during the 2010 Gisborne slow slip event.
+As was the case for the microearthquakes, the source of the tremor was located downdip of the slow slip patch determined from GNSS data.
+~\citep{IDE_2012} detected tremor downdip of the location of two deep slow slip events observed by ~\citet{WAL_2013} in 2006 and 2008.
+However, contrary to ETS events in Cascadia and Nankai, the tremor activity did not seem to increase during the slow slip events.
+~\citet{TOD_2016}  detected tremor associated with most of the shallow slow slip events between 2010 and 2015, and located downdip of the geodetically inferred slip area.
+They also detected deeper tremor between 20 and 50 km depth with unclear origin.
+They hypothesized that these tremor may be related to undetected deep long-term slow slip events.
+\\
+
+To evaluate whether the wavelet analysis is effective in a region without robust tremor, we take all the New Zealand GPS stations located in a 50 km radius of a given \textbf{location}, for the 18 locations indicated in Figure 12, and we stack the 6th level details, the 7th level details or the 8th level details over all the GPS stations.
+We then sum together the 6th, 7th and 8th levels stacked wavelet details (Figure 13, top panel).
+We highlight positive and negative peaks with red and blue colors as was done in Figure 9.
+We cannot use the tremor data to decide what is the appropriate threshold above which we consider that there is a slow slip event.
+\textbf{Slow slip events in New Zealand result in surface displacements that are similar in amplitude to twice as large as those observed in Cascadia.} Therefore, the amplitudes of the peaks in the wavelet details should be similar in New Zealand and in Cascadia and we choose identical thresholds for both regions.
+As a slow slip event in northern New Zealand results in a displacement in the east direction at the earth’s surface, the slow slip events are indicated by a negative peak followed by a positive peak in the stacked wavelet details.
+We compare the results of the timings and locations of the slow slip events to those events detected by ~\citet{TOD_2016}.
+As they only used data from five GPS stations (PUKE, ANAU, GISB, MAHI and CKID), we indicate by a vertical orange bar on the bottom panel of Figure 13 each time a slow slip event was detected for these stations.
+The orange bars are centered on the latitudes of the GPS stations.
+If a slow slip event was detected by more than one station, all the corresponding orange bars are linked together to show the spatial extent of the slow slip.
+~\citet{TOD_2016} indicated by a question mark (on their Figure 2 and their Table 1) additional possible events, and those are indicated by a dotted orange bar on Figure 13.
+To compare with the slow slip events detected with the wavelet method, we also mark by a green bar every time a negative peak lower than the threshold is followed by a positive peak higher than the threshold.
+Table 5 summarizes the slow slip events detected with the wavelet method for 2010-2016.
+\\
+
+We observe that there is a good \textbf{agreement} between the events detected with the wavelet method and the events previously detected by ~\citet{TOD_2016}.
+We clearly see an event propagating from south to north in January-February (event 2 from ~\citet{TOD_2016}), an event in March-April 2010 (event 3), an event in April-May 2011 in the northern part of the region studied (events 6 and 7), an event propagating south-to-north in August-September and September-October 2011 (events 8 and 9), and an event in December 2011 (event 10).
+Although ~\citet{TOD_2016} only detected this last event for GPS station GISB, it seems that this event may have also extended farther to the north and the south.
+We then clearly see an event in the northern part of the region studied in August 2012 (event 12), an event in December 2012-January 2013 (event 13), an event in the southern part of the region studied in February-March 2013 (event 14), an event propagating from south to north in June-July and July-August 2013 (events 15 and 16), an event in September 2014 (events 20 and 21), an event in the southern part of the region studied in December 2014-January 2015 (events 22 and 23), and an event in June-July 2015 in the northern part of the region studied (event 26).
+It is unclear if the event near station ANAU in early 2010 (event 1) is visible in the wavelet details as it is too close to the beginning of the time series.
+The June-July 2010 event (event 4), the August 2010 event (event 5), and the March 2012 event (event 11), are not clearly visible in the wavelet details.
+The events in September-October 2013 (event 17), December 2013 (event 18), May-June 2014 (event 19), January-February (event 24) and February 2015 (events 25) are not clearly seen in the wavelet details, but there could be a small negative peak followed by a small positive peak at these times.
+Additionally, there could be two other events that are not in \citep{TOD_2016} in Fall 2010 (southern part of the region studied) and in Fall 2015.
+\\
+
+Our wavelet-based method thus works well to detect transients in GPS data that could be slow slip events, even in the absence of tremor data.
+The choice of the appropriate threshold to decide that there is a transient and the levels of the wavelet details that we look at for the detection may still not be easily made.
+There is a difference between Cascadia and New Zealand in terms of which wavelet details to stack.
+In particular, as there is more time between two slow slip events in New Zealand than in Cascadia, the biggest slow slip events (early 2010, late 2011, 2013 and late 2014) can also be \textbf{seen} on the 9th level detail for New Zealand, whereas they could not be seen for Cascadia.
+We then use the method to detect slow slip events during the period 2016-2022, which was not covered by ~\citet{TOD_2016} (Figure 14).
+We note four large transients that could be slow slip events in late 2016, late 2017, early 2019 and mid-2021.
+There are also possible smaller events in the northern part of the area in mid-2018 and in most of the area studied in early 2020.
+Table 6 summarizes the slow slip events detected with the wavelet method for 2016-2022.
+\\
+
+\textbf{The method is thus applicable in regions where tremor data are not usable.
+To determine which wavelet levels to stack, we recommend analyzing each level detail.
+Look for spatially coherent patterns, wavelet details with energy at similar times and high signal-to-noise ratios.
+Look for alternating positive and negative peaks that are consistent with the expected direction of slow slip.
+Consider wavelet details with time scales ranging from the expected duration of slow slip events to the expected recurrence times between slow slip events.
+For Cascadia and New Zealand this would be weeks to years.
+Determination of a threshold is subjective.
+At large thresholds the large slow slip events should be clear.
+At smaller thresholds there is the possibility of identifying smaller events, but at the risk of false detections.}
 
@@
 We assumed that there is a slow slip event whenever the wavelet signal gets above a threshold.
-We verified that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
+We verified that there is a good \textbf{agreement} between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data.
 The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g.
 ~\citep{MIC_2019}).
-We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between true detections and false detections.
+We detected signals in the GPS data that could be magnitude 5 events, but it is not easy to differentiate between robust detections and false detections.
+We then applied the method to GNSS data in New Zealand and detected slow slip events consistent with the events previously detected by ~\citet{TOD_2016}.
 
@@
  \begin{table}[hbt!]
- \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
+ \caption{Magnitude \textbf{5 to 6} events from ~\citet{MIC_2019}.}
  \centering
@@
 \item Figure 2.
-Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).
+Top left: \textbf{East-west} displacement recorded at GPS station PGC5.
+The resulting details and smooth of the wavelet decomposition are shown in increasing level from top to bottom and from left to right.
 
-\item Figure 3. Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.
-
-\item Figure 4.
+\item Figure 3.
 GPS stations used in this study (black triangles).
@@
 The black lines represent the timings of the ETS events from Table 1.
-We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 (for the GPS) or 0.003 (for the tremor).
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 mm (for the GPS) or 0.003 (for the tremor, \textbf{that is about 17 times the average value of the signal}).
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 7. Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).
+\item Figure 6.
+Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 7th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.5 mm (for the GPS) or 0.01 (for the tremor, \textbf{that is about 56 times the average value of the signal}).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 8. Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).
+\item Figure 7.
+Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 6th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.3 mm (for the GPS) or 0.009 (for the tremor, \textbf{that is about 51 times the average value of the signal}).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
-\item Figure 9. Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).
+\item Figure 8.
+ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
+The black cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
+The values of the threshold are color-coded.
+Reds (bottom curve) correspond to the lowest value of the threshold for the tremor (0.001), while oranges, greens, blues, purples correspond to increasing values of the threshold for the tremor (up to 0.01, top curve).
+The brightest colors (bottom left) correspond to the highest values of the threshold for the GPS (1.5 mm), while the darker colors (top right) correspond to decreasing values of the threshold for the GPS (0.1 mm).
+
+\item Figure 9.
+Top: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: Sum of the 6th, 7th and 8th levels detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.8 mm (for the GPS) or 0.01 (for the tremor, \textbf{that is about 56 times the average value of the signal}).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.
 
 \item Figure 10.
-ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
-Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
-
-\item Figure 11.
 Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
@@
 \begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Top: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the missing values replaced by linear interpolation plus Gaussian noise (red).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
-\caption{Top: Longitudinal displacement recorded at GPS station PGC5. The resulting details and smooth of the wavelet decomposition are shown in increasing level.}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
+\caption{Top left: \textbf{East-west} displacement recorded at GPS station PGC5.
+The resulting details and smooth of the wavelet decomposition are shown in increasing level from top to bottom and from left to right.}
 \label{pngfiguresample}
@@
 The black lines represent the timings of the ETS events from Table 1.
-We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 (for the GPS) or 0.003 (for the tremor).
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.4 mm (for the GPS) or 0.003 (for the tremor, \textbf{that is about 17 times the average value of the signal}).
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
-\caption{Same as Figure 6 but for the 7th level detail. The thresholds are 0.5 (for the GPS) and 0.01 (for the tremor).}
+\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 7th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.5 mm (for the GPS) or 0.01 (for the tremor, \textbf{that is about 56 times the average value of the signal}).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.eps}
-\caption{Same as Figure 6 but for the 6th level detail. The thresholds are 0.3 (for the GPS) and 0.009 (for the tremor).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6-7-8.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
-\caption{Same as Figure 6 but for the sum of the 6th, 7th and 8th level details. The thresholds are 0.8 (for the GPS) and 0.01 (for the tremor).}
+\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: 6th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.3 mm (for the GPS) or 0.009 (for the tremor, \textbf{that is about 51 times the average value of the signal}).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}
@@
 \caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
-Each black dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
-The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.}
+Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
+The black cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.
+The values of the threshold are color-coded.
+Reds (bottom curve) correspond to the lowest value of the threshold for the tremor (0.001), while oranges, greens, blues, purples correspond to increasing values of the threshold for the tremor (up to 0.01, top curve).
+The brightest colors (bottom left) correspond to the highest values of the threshold for the GPS (1.5 mm), while the darker colors (top right) correspond to decreasing values of the threshold for the GPS (0.1 mm).}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8.eps}
+
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
+\caption{Top: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+Bottom: Sum of the 6th, 7th and 8th levels detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
+The black lines represent the timings of the ETS events from Table 1.
+We mark by a red rectangle every time where the amplitude is higher than a threshold of 0.8 mm (for the GPS) or 0.01 (for the tremor, \textbf{that is about 56 times the average value of the signal}).
+We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
 \label{pngfiguresample}

### 2022-06-29 paper/supplement.tex
Ready to submit

@@
 \begin{figure}
-\noindent\includegraphics[width=4.5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{Top: \textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 on the left and 2011.48 on the right) and replaced by the sum of a straight line and a Gaussian noise component (red)}.
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the \textbf{data with displacement values removed and replaced} by linear interpolation plus Gaussian noise (red).}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/boundaries.eps}
+\caption{\textbf{Top left: East-west displacement recorded at GPS station PGC5.
+Top to bottom and left to right: Wavelet details and smooth computed using only the data from 2008 to 2012 (red) and using the entire time series from 2000 to 2021 (black).}}
+\label{pngfiguresample}
+\end{figure}
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
+\caption{\textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 for the first four rows and 2011.48 for the last four rows) and replaced by the sum of a straight line and a Gaussian noise component (red).
+The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the data with displacement values removed and replaced by linear interpolation plus Gaussian noise (red).}}
 \label{pngfiguresample}
@@
 
+\begin{figure}
+\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8_events.eps}
+\caption{Same as top panel of Figure 9 from the main text. We highlight by four red rectangles the four possible small transients discussed in the text.}
+\label{pngfiguresample}
+\end{figure}
+
 \end{document}

### 2022-06-29 src/boundaries.py

@@ -1,0 +1,138 @@
+""" Script to look at MODWT of GPS data
+and check the effect of boundaries on the MODWT """
+
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
+import numpy as np
+
+from math import log, sqrt
+from matplotlib import ticker
+
+import DWT, MODWT
+from MODWT import get_DS, get_scaling, inv_pyramid, pyramid
+
+# MODWT parameters
+wavelet = 'LA8'
+J = 10
+
+# Choose the station
+station = 'PGC5'
+direction = 'lon'
+dataset = 'cleaned'
+filename = '../data/PANGA/' + dataset + '/' + station + '.' + direction
+
+# Load the data
+data = np.loadtxt(filename, skiprows=26)
+time = data[:, 0]
+disp = data[:, 1]
+error = data[:, 2]
+sigma = np.std(disp)
+
+# Correct for the repeated values
+dt = np.diff(time)
+gap = np.where(dt < 1.0 / 365.0 - 0.0001)[0]
+for i in range(0, len(gap)):
+    if (gap[i] + 2 < len(time)):
+        if ((time[gap[i] + 2] - time[gap[i] + 1] > 2.0 / 365.0 - 0.0001) \
+        and (time[gap[i] + 2] - time[gap[i] + 1] < 2.0 / 365.0 + 0.0001)):
+            time[gap[i] + 1] = 0.5 * (time[gap[i] + 2] + time[gap[i]])
+        elif (gap[i] + 3 < len(time)):
+            if ((time[gap[i] + 2] - time[gap[i] + 1] > 1.0 / 365.0 - 0.0001) \
+            and (time[gap[i] + 2] - time[gap[i] + 1] < 1.0 / 365.0 + 0.0001) \
+            and (time[gap[i] + 3] - time[gap[i] + 2] > 2.0 / 365.0 - 0.0001) \
+            and (time[gap[i] + 3] - time[gap[i] + 2] < 2.0 / 365.0 + 0.0001)):
+                time[gap[i] + 1] = time[gap[i] + 2]
+                time[gap[i] + 2] = 0.5 * (time[gap[i] + 2] + time[gap[i] + 3])
+
+# Look for gaps greater than 1 day
+days = 2
+dt = np.diff(time)
+gap = np.where(dt > days / 365.0 - 0.0001)[0]
[…]

### 2022-06-29 src/compare_tremor_GPS.py

@@ -420,7 +420,8 @@ def plot_correlations():
 
 
 
-def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, threshold, events, small_events):
+def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, \
+    threshold, events, small_events, possible_events, MODWT_events):
     """
     """
     # Read station file
@@ -436,8 +437,8 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
               'ytick.labelsize':24}
     pylab.rcParams.update(params)
 
-    for event in events:
-        plt.axvline(event, color='black', linewidth=2)
+#    for event in events:
+#        plt.axvline(event, color='black', linewidth=2)
 #    for event in small_events:
 #        plt.axvline(event, color='blue', linewidth=1)
 
@@ -466,7 +467,7 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
 
         # Read output files from wavelet transform
         for (station, lon_sta, lat_sta) in zip(sub_stations['name'], sub_stations['longitude'], sub_stations['latitude']):
-            filename = 'MODWT_GPS_mode/' + dataset + '_' + station + '_' + direction + '.pkl'
+            filename = 'MODWT_GPS/' + dataset + '_' + station + '_' + direction + '.pkl'
             (time, disp, W, V, D, S) = pickle.load(open(filename, 'rb'))
             if ((np.min(time) <= 2021.5) and (np.max(time) >= 2006.0)):
                 times.append(time)
@@ -534,21 +535,34 @@ def plot_GPS(station_file, lats, lons, dataset, direction, radius_GPS, J, thresh
             difference = np.diff(times_slowslip)
             jumps = np.where(difference > 1.5 / 365.25)[0]
             
-            if len(jumps > 0):
-                begin_jumps = np.insert(jumps + 1, 0, 0)
-                end_jumps = np.append(jumps, len(times_slowslip) - 1)
-                begin_times = times_slowslip[begin_jumps]
-                end_times = times_slowslip[end_jumps]
+#            if len(jumps > 0):
[…]

### 2022-06-29 src/compare_tremor_GPS_NZ.py

@@ -20,7 +20,7 @@ import DWT
 from MODWT import get_DS, get_scaling, pyramid
 
 
-def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, events, possible_events):
+def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, events, possible_events, MODWT_events):
     """
     """
     # Read station file
@@ -30,8 +30,8 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, even
     # Time limits
 #    tmin = 2000 + 909 / 365.25
 #    tmax = 2000 + 8127 / 365.25
-    tmin = 2010.0
-    tmax = 2016.0
+    tmin = 2016.0
+    tmax = 2022.0
 
     a = 6378.136
     e = 0.006694470
@@ -144,21 +144,32 @@ def plot_GPS(station_file, lats, lons, direction, radius_GPS, J, threshold, even
                 for i in range(0, len(jumps) + 1):
                     x0 = begin_times[i]
                     dx = end_times[i] - begin_times[i]
-#                    if disps_slowslip[begin_jumps[i]] > 0:
-#                        ax.add_patch(Rectangle((x0, lat + 0.01), dx, 0.03, facecolor='red'))
-#                    else:
-#                        ax.add_patch(Rectangle((x0, lat - 0.04), dx, 0.03, facecolor='blue'))
+                    if disps_slowslip[begin_jumps[i]] > 0:
+                        ax.add_patch(Rectangle((x0, lat + 0.01), dx, 0.03, facecolor='red'))
+                    else:
+                        ax.add_patch(Rectangle((x0, lat - 0.04), dx, 0.03, facecolor='blue'))
 
             plt.plot(times_stacked, lat + 0.1 * disps_stacked, color='black')
 
+#            np.savetxt('files_NZ/GPS_NZ_detail_' + str(J + 1) + '_loc_' + \
+#                str(index) + '.txt', np.transpose(np.vstack([times_stacked, disps_stacked])))
+
     # Add slow slip events from Todd and Schwartz (2016)
-    for event in events:
-        plt.plot([event['time'], event['time']], \
[…]

### 2022-06-29 src/missing_values.py

@@ -1,6 +1,7 @@
 """ Script to look at MODWT of GPS data
 with different methods to fill in missing values """
 
+import matplotlib.pylab as pylab
 import matplotlib.pyplot as plt
 import numpy as np
 
@@ -54,10 +55,15 @@ np.random.seed(0)
 # Loop on missing values
 gaps = [605, 790]
 
+params = {'xtick.labelsize':24,
+          'ytick.labelsize':24}
+pylab.rcParams.update(params) 
+
 # Loop on length of gap
 for n in range(1, nmax, nstep):
 
-    plt.figure(1, figsize=(len(gaps) * 3, (J + 2) * 2))
+#    plt.figure(1, figsize=(len(gaps) * 3, (J + 2) * 2))
+    plt.figure(1, figsize=(30, 3 * 4 * len(gaps)))
 
     for count, gap in enumerate(gaps):
         disp_interp = np.copy(disp)
@@ -81,29 +87,51 @@ for n in range(1, nmax, nstep):
         [W, V] = pyramid(disp_interp, name, J)
         (D, S) = get_DS(disp_interp, W, name, J)
 
+        maxD = max([np.max(Dj) for Dj in D])
+        minD = min([np.min(Dj) for Dj in D])
+
         # Plot data
-        ax = plt.subplot2grid((J + 2, len(gaps)), (0, count))
+#        ax = plt.subplot2grid((J + 2, len(gaps)), (0, count))
+        ax = plt.subplot2grid((4 * len(gaps), 3), (4 * count + 0, 0))
         ax.xaxis.set_major_formatter(ticker.FormatStrFormatter('%.1f'))
         plt.plot(time, disp_interp, 'r')
         plt.plot(time, disp, 'k', label='Data')
         plt.xlim(np.min(time[gap - 100]), np.max(time[gap + n + 99]))
-        plt.legend(loc=1)        
+        plt.ylim([np.min(disp), np.max(disp)])
+        plt.legend(loc=3, fontsize=20)
+        ax.axes.xaxis.set_ticklabels([])       
         # Plot details at each level
         for j in range(0, J):
-            ax = plt.subplot2grid((J + 2, len(gaps)), (j + 1, count))
+#            ax = plt.subplot2grid((J + 2, len(gaps)), (j + 1, count))
+            if j < 3:
+                ax = plt.subplot2grid((4 * len(gaps), 3), (4 * count + j + 1, 0))
+            elif j < 7:
[…]

### 2022-06-29 src/synthetics.py

@@ -12,8 +12,8 @@ import MODWT
 #timesteps = [100, 200, 500, 1000]
 timesteps = [500]
 # Durations of slow slip events
-#durations = [5, 10, 20, 40]
-durations = [10, 20]
+durations = [5, 10, 20, 40]
+#durations = [10, 20]
 # MODWT wavelet filter
 name = 'LA8'
 # Duration of recording
@@ -34,8 +34,8 @@ for timestep in timesteps:
     params = {'xtick.labelsize':20,
               'ytick.labelsize':20}
     pylab.rcParams.update(params)   
-#    fig = plt.figure(1, figsize=(5 * len(durations), 3 * (J + 2)))
-    fig = plt.figure(1, figsize=(5 * len(durations), 3 * 2))
+    fig = plt.figure(1, figsize=(5 * len(durations), 3 * (J + 2)))
+#    fig = plt.figure(1, figsize=(5 * len(durations), 3 * 2))
 
     # Loop on duration
     for i, duration in enumerate(durations):
@@ -50,40 +50,42 @@ for timestep in timesteps:
         maxD = max([np.max(Dj) for Dj in D])
         minD = min([np.min(Dj) for Dj in D])
         # Plot data
-#        ax = plt.subplot2grid((J + 2, len(durations)), (0, i))
+        ax = plt.subplot2grid((J + 2, len(durations)), (0, i))
 #        ax = plt.subplot2grid((2, len(durations)), (0, i))
-#        plt.plot(time, disp, 'k', label='Data')
-#        plt.ylim(-2.0, 2.0)
-#        plt.legend(loc=3, fontsize=20)
-#        if i != 0:
-#            ax.axes.yaxis.set_ticks([])
-#        title = 'Duration of event = ' + str(duration) + ' days'
-#        plt.title(title, fontsize=20)
+        plt.plot(time, disp, 'k', label='Data')
+        plt.ylim(-2.0, 2.0)
+        plt.legend(loc=3, fontsize=20)
+        ax.axes.xaxis.set_ticklabels([])
+        if i != 0:
+            ax.axes.yaxis.set_ticklabels([])
+        title = 'Duration of event = ' + str(duration) + ' days'
+        plt.title(title, fontsize=20)
         # Plot details
-        for j in range(9, 10):
-#            ax = plt.subplot2grid((J + 2, len(durations)), (j + 1, i))
-            ax = plt.subplot2grid((2, len(durations)), (j - 9, i))
+        for j in range(0, J):
[…]

### 2022-06-29 src/tremor_MODWT_longer.py

@@ -134,31 +134,48 @@ def compute_wavelets_tremor(station_file, lats, lons, dataset, direction, \
         params = {'xtick.labelsize':24,
                   'ytick.labelsize':24}
         pylab.rcParams.update(params)   
-        fig = plt.figure(1, figsize=(10, 3 * (J + 3)))
+#        fig = plt.figure(1, figsize=(10, 3 * (J + 3)))
+        fig = plt.figure(1, figsize=(30, 3 * 4))
 
         maxD = max([np.max(Dj) for Dj in D])
         minD = min([np.min(Dj) for Dj in D])
 
         # Plot data
-        plt.subplot2grid((J + 2, 1), (0, 0))
+#        plt.subplot2grid((J + 2, 1), (0, 0))
+        ax = plt.subplot2grid((4, 3), (0, 0))
         plt.plot(times_stacked, tremor_detrend, 'k', label='Data')
-        plt.xlabel('Time (years)', fontsize=24)
         plt.xlim([2006.0, 2021.5])
         plt.ylim([np.min(tremor_detrend), np.max(tremor_detrend)])
         plt.legend(loc=3, fontsize=20)
+        ax.axes.xaxis.set_ticklabels([])
         # Plot details
         for j in range(0, J):
-            plt.subplot2grid((J + 2, 1), (j + 1, 0))
+#            plt.subplot2grid((J + 2, 1), (j + 1, 0))
+            if j < 3:
+                ax = plt.subplot2grid((4, 3), (j + 1, 0))
+            elif j < 7:
+                ax = plt.subplot2grid((4, 3), (j - 3, 1))
+            else:
+                ax = plt.subplot2grid((4, 3), (j - 7, 2))
             plt.plot(times_stacked, D[j], 'k', label='D' + str(j + 1))
             plt.xlim([2006.0, 2021.5])
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
+            if ((j == 2) or (j == 6)):
+                plt.xlabel('Time (years)', fontsize=24)
+            else:
+                ax.axes.xaxis.set_ticklabels([])
+            if j >= 3:
+                ax.axes.yaxis.set_ticklabels([])
         # Plot smooth
-        plt.subplot2grid((J + 2, 1), (J + 1, 0))
+#        plt.subplot2grid((J + 2, 1), (J + 1, 0))
+        ax = plt.subplot2grid((4, 3), (3, 2))
[…]

### 2022-06-29 src/vespagram.py

@@ -120,41 +120,52 @@ def compute_wavelets(station_file, lats, lons, radius, direction, dataset, \
                   'ytick.labelsize':24}
         pylab.rcParams.update(params)   
  #       fig = plt.figure(1, figsize=(10, 3 * (J + 3)))
-        fig = plt.figure(1, figsize=(10, 3 * 4))
+        fig = plt.figure(1, figsize=(30, 3 * 4))
 
         maxD = max([np.max(Dj) for Dj in D])
         minD = min([np.min(Dj) for Dj in D])
 
         # Plot data
 #        plt.subplot2grid((J + 2, 1), (0, 0))
-#        plt.subplot2grid((4, 1), (0, 0))
-#        plt.plot(time, disp, 'k', label='Data')
-#        plt.xlim([2006.0, 2021.5])
-#        plt.ylim([np.min(disp), np.max(disp)])
-#        plt.legend(loc=3, fontsize=20)
+        ax = plt.subplot2grid((4, 3), (0, 0))
+        plt.plot(time, disp, 'k', label='Data')
+        plt.xlim([2006.0, 2021.5])
+        plt.ylim([np.min(disp), np.max(disp)])
+        plt.legend(loc=3, fontsize=20)
+        ax.axes.xaxis.set_ticklabels([])
         # Plot details
-        for j in range(7, J):
+        for j in range(0, J):
 #            plt.subplot2grid((J + 2, 1), (j + 1, 0))
-            plt.subplot2grid((4, 1), (j - 7, 0))
+            if j < 3:
+                ax = plt.subplot2grid((4, 3), (j + 1, 0))
+            elif j < 7:
+                ax = plt.subplot2grid((4, 3), (j - 3, 1))
+            else:
+                ax = plt.subplot2grid((4, 3), (j - 7, 2))
             plt.plot(time, D[j], 'k', label='D' + str(j + 1))
             plt.xlim([2006.0, 2021.5])
             plt.ylim(minD, maxD)
             plt.legend(loc=3, fontsize=20)
-#            if j == 6:
-#                plt.xlabel('Time (years)', fontsize=24)
+            if ((j == 2) or (j == 6)):
+                plt.xlabel('Time (years)', fontsize=24)
+            else:
+                ax.axes.xaxis.set_ticklabels([])
+            if j >= 3:
+                ax.axes.yaxis.set_ticklabels([])
         # Plot smooth
 #        plt.subplot2grid((J + 2, 1), (J + 1, 0))
[…]

### 2022-07-06 paper/paper_bssa.tex
Updated submission

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 \usepackage{lineno}
@@
 \author[2]{David A. Schmidt}
-\affil[1]{Corresponding author. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
+\affil[1]{Corresponding author.
+University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310, \href{mailto:ariane.ducellier.pro@gmail.com}{ariane.ducellier.pro@gmail.com}}
 \affil[2]{University of Washington, Department of Earth and Space Sciences}
@@
 Figures 3 and 12 were created using GMT ~\citep{WES_1991}.
+The Supplementary Material contains three figures showing the effects of boundary conditions, missing data and common modes, and a figure showing four additional small displacements detected in the GPS data.
 
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{plainnat_modif}
 \bibliography{bibliography}

### 2022-07-06 paper/paper_bssa_annotated.tex
Updated submission

@@
 Figures 3 and 12 were created using GMT ~\citep{WES_1991}.
+\textbf{The Supplementary Material contains three figures showing the effects of boundary conditions, missing data and common modes, and a figure showing four additional small displacements detected in the GPS data.}
 

### 2022-07-12 paper/paper_bssa.tex
REvised for publication

@@
 
-\section*{Key points}
-
-\begin{itemize}
-\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
-\item In Northern Cascadia, there is good agreement between detections of slow slip using GNSS data and using tremor data.
-\item The method can be applied as a starting point to identify slow slip without prior information. New Zealand is shown as on example.
-\end{itemize}
-
-\newpage
+%\section*{Key points}
+%
+%\begin{itemize}
+%\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
+%\item In Northern Cascadia, there is good agreement between detections of slow slip using GNSS data and using tremor data.
+%\item The method can be applied as a starting point to identify slow slip without prior information. New Zealand is shown as an example.
+%\end{itemize}
+%
+%\newpage
 
@@
 Figures 3 and 12 were created using GMT ~\citep{WES_1991}.
-The Supplementary Material contains three figures showing the effects of boundary conditions, missing data and common modes, and a figure showing four additional small displacements detected in the GPS data.
+Supplemental Material for this article includes three figures showing the effects of boundary conditions, missing data and common modes, and a figure showing four additional small displacements detected in the GPS data.
 
@@
 
-The authors would like to thank two anonymous reviewers and the Associate Editor Jeanne Hardebeck, whose comments greatly helped improve the manuscript.
+The authors would like to thank two anonymous reviewers, the Associate Editor Jeanne Hardebeck and the Editor-in-Chief P.
+Martin Mai, whose comments greatly helped improve the manuscript.
 This work was funded by the grant from the National Science Foundation EAR-1358512.
@@
 
-Ariane Ducellier. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310. \\
+Ariane Ducellier.
+University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310.
+\href{mailto:ariane.ducellier.pro@gmail.com}{ariane.ducellier.pro@gmail.com}\\
 
@@
 
-\newpage
-
-\section*{Figures}
-
-\begin{figure}
-\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
-\caption{Demonstration of a wavelet decomposition for a synthetic dataset.
+%\newpage
+%
+%\section*{Figures}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=11cm, trim={0cm 0cm 0cm 0cm},clip]{figures/500_DS.eps}
+%\caption{Demonstration of a wavelet decomposition for a synthetic dataset.
 A synthetic time series is created (top row) with steps of period 500 days, and transient durations of 2 days (left), 5 days, 10 days, and 20 days (right).
@@
 The amplitude of the synthetic time series is normalized to 1, and the details and smooths show the relative amplitude.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
-\caption{Top left: East-west displacement recorded at GPS station PGC5.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/cleaned_PGC5_lon.eps}
+%\caption{Top left: East-west displacement recorded at GPS station PGC5.
 The resulting details and smooth of the wavelet decomposition are shown in increasing level from top to bottom and from left to right.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations.eps}
-\caption{GPS stations used in this study (black triangles).
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations.eps}
+%\caption{GPS stations used in this study (black triangles).
 The black line represents the 40 km depth contour of the plate boundary model by ~\citet{PRE_2003}.
@@
 The small grey dots are all the tremor locations from the PNSN catalog.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
-\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.eps}
-\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_13.eps}
+%\caption{Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost red triangles on Figure 3 (latitude 48.5).}
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.eps}
+%
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.eps}
+%\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
 Bottom: 8th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
@@
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
-\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.eps}
+%
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
+%\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
 Bottom: 7th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
@@
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.eps}
-\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.eps}
+%
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6.eps}
+%\caption{Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
 Bottom: 6th level detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
@@
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/ROC_6-7-8.eps}
-\caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/ROC_6-7-8.eps}
+%\caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition.
 Each dot represents the true positive rate of event detections and the false positive rate of event detections for a given pair of thresholds (for the GPS and for the tremor).
@@
 The brightest colors (bottom left) correspond to the highest values of the threshold for the GPS (1.5 mm), while the darker colors (top right) correspond to decreasing values of the threshold for the GPS (0.1 mm).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
-\caption{Top: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8.eps}
+%
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
+%\caption{Top: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
 Bottom: Sum of the 6th, 7th and 8th levels detail multiplied by -1 of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.
@@
 We mark by a blue rectangle every time where the amplitude is lower than minus the threshold.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.eps}
-\caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_michel_detail_5.eps}
+%\caption{Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
 The red lines represent the timings of the ETS events from Table 1.
 The blue lines represent the timings of the magnitude 5 events from the catalog of ~\citet{MIC_2019}.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8_slowslip.eps}
-\caption{Same as top panel of Figure 9: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_details_6-7-8_slowslip.eps}
+%\caption{Same as top panel of Figure 9: Stacked sum of the 6th, 7th and 8th levels details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 red triangles indicated in Figure 3.
 We mark with a green bar the slow slip events from Table 4 detected with the wavelet method.
 Full lines correspond to robust detections (1 in Table 4) and dotted lines to less robust detections (2 in Table 4).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations_NZ.eps}
-\caption{GPS stations used for the slow slip detection in New Zealand (black triangles).
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/map_GPS_stations_NZ.eps}
+%\caption{GPS stations used for the slow slip detection in New Zealand (black triangles).
 The red triangles are the locations where we stack the GPS data.
 They are located close to the 20 km depth contour of the plate boundary from ~\citet{WIL_2013}.}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8.eps}
-
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_events.eps}
-\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8.eps}
+%
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_events.eps}
+%\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
 The time period covered is 2010-2016.
@@
 Full lines correspond to robust detections (1 in Table 5) and dotted lines to less robust detections (2 in Table 5).}
-\label{pngfiguresample}
-\end{figure}
-
-\begin{figure}
-\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_2016-2022.eps}
-\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
+%\label{pngfiguresample}
+%\end{figure}
+%
+%\begin{figure}
+%\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_NZ_details_6-7-8_2016-2022.eps}
+%\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 18 red triangles indicated in Figure 12.
 The time period covered in 2016-2022.
@@
 Full lines correspond to robust detections (1 in Table 6) and dotted lines to less robust detections (2 in Table 6).}
-\label{pngfiguresample}
-\end{figure}
+%\label{pngfiguresample}
+%\end{figure}
 

### 2022-08-09 paper/supplement.tex
Edit proof

@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/boundaries.eps}
-\caption{\textbf{Top left: East-west displacement recorded at GPS station PGC5.
-Top to bottom and left to right: Wavelet details and smooth computed using only the data from 2008 to 2012 (red) and using the entire time series from 2000 to 2021 (black).}}
+\caption{Top left: East-west displacement recorded at GPS station PGC5.
+Top to bottom and left to right: Wavelet details and smooth computed using only the data from 2008 to 2012 (red) and using the entire time series from 2000 to 2021 (black).}
 \label{pngfiguresample}
@@
 \noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
-\caption{\textbf{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 for the first four rows and 2011.48 for the last four rows) and replaced by the sum of a straight line and a Gaussian noise component (red).
-The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the data with displacement values removed and replaced by linear interpolation plus Gaussian noise (red).}}
+\caption{Original data from GPS station PGC5 (black) and same data where displacement values have been artificially removed at two different times (2010.97 for the first four rows and 2011.48 for the last four rows) and replaced by the sum of a straight line and a Gaussian noise component (red).
+The corresponding ten details and smooths of the wavelet composition are shown in increasing levels for the original data (black) and for the data with displacement values removed and replaced by linear interpolation plus Gaussian noise (red).}
 \label{pngfiguresample}
